### 2024-10-10 Theory recap/notes.tex
Add Theory recap

\documentclass{article}
\input{math_commands}
\input{custom_imports}
\usepackage[margin=54pt]{geometry}
\usepackage{biblatex}
\addbibresource{biblio.bib}

\begin{document}

\date{Theory recap}
\author{Marco Sterlini}

\title{Title}
\maketitle

\section*{First works}
My contribution started by exploring the works \cite{css-paper} and \cite{css-extended}. The problem addressed is the design of an event-triggered control (ETC) for discrete-time linear-time-invariant (LTI) systems.
The aim was to decrease the computational load that follows the application of non-linear activation functions inside the neural network (NN) controller. Also thanks to previous works \cite{sophie-book-saturation}, \cite{arcak-first}, \cite{iqc-intro} several tools to treat these non-linearities have been studied and applied allowing the use of classical stability analysis methods to highly non-linear NN controllers. Lyapunov theory will be used to guarantee the stability of the NN controlled plant.

\subsection*{System under analysis}

The feedback system under examination is the following:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.6\textwidth]{img/simple-etm}
    \label{}
\end{figure}

The NN controller is called $\pi_{ETM}$, the plant to stabilize is $F$, and it is described by the dynamics:

\begin{align*}
    &x(k + 1) = A_{F} x(k) + B_{F} u(k)\\
    \\
    &\hat{\omega}^{0}(k) = x(k),\\
    &\nu^{i}(k) = W^{i} \hat{\omega}^{i-1}(k) + b^{i},\ \ i \in \left\{ 1, \dots, l \right\},\\
    &\omega^{i}(k) = \text{sat}(\nu^{i}(k)),\\
    &u(k) = W^{l+1} \hat{\omega}^{l} (k) + b^{l+1}\\ 
\end{align*}

Where the first equation describes the plant $F$ as a discrete-time linear time-invariant system, the other equations describe the behavior of the NN controller: $\hat{\omega}^{i}$ and $\omega^{i}$ are respectively the last and current output of the $i^{th}$ layer, $\nu^{i}$ is the current input to the $i^{th}$ layer. The control input $u(k)$ is hence the output of the last layer of the NN, also the weights and the biases of the $i^{th}$ layer are indicated by $W^{i}$ and $b^{i}$. As per the activation function the saturation function was used and is defined by:
$$
   \text{sat}(\nu^{i}_{j}(k)) = \text{sign}(\nu^{i}_{j}(k)) min(|\nu^{i}_{j}(k)|, \bar{\nu}^{i}_{j})
$$  
For the $j^{th}$ neuron of the $i^{th}$ layer where $\bar{\nu}^{i}_{j}$ is the relative maximum limit value for the input signal. The activation function is applied element wise but a notation to isolate the non linearities is used:
$$
    \nu_{\phi} = \bmx{\nu^{1\tr}, \dots \nu^{l\tr}}\tr, \omega_{\phi} = \bmx{\omega^{1\tr}, \dots \omega^{l\tr}}\tr, \hat{\omega}_{\phi} = \bmx{\hat{\omega}^{1\tr}, \dots \hat{\omega}^{l\tr}}\tr \in \RR^{n_{\phi}}
$$
With $n_{\phi}$ the total number of neurons involved in the controller. It is then possible to express:
$$
    \text{sat}(v_{\phi}) = \bmx{\text{sat}(\nu^{1})\tr, \dots, \text{sat}(\nu^{l})\tr}\tr \in \RR^{n_{\phi}}
$$
The control input $u(k)$ and the input layers' vector $\nu_{\phi}$ can be expressed as a function of the state $x(k)$ and the last layers' output vector $\hat{\omega}_{\phi}(k)$

\begin{equation}
    \bmx{u(k)\\ \nu_{\phi}(k)} = \bmx{\begin{array}{c|cccc|c} 
        \0 & \0 & \dots & \0 & W^{l+1} & b^{l+1}\\
        \hline
        W^{1} & \0 & \dots & \0 & \0 & b^{1}\\
        \0 & W^{2} & \dots & \0 & \0 & b^{2}\\
        \vdots & \vdots & \ddots & \vdots & \vdots & \vdots\\
        \0 & \0 & \dots & W^{l} & \0 & b^{l}
    \end{array}}  \bmx{x(k) \\ \hat{\omega}_{\phi}(k) \\ 1}
\end{equation}

$$
    \bmx{u(k)\\ \nu_{\phi}(k)} = N  \bmx{x(k) \\ \hat{\omega}_{\phi}(k) \\ 1}\text{ and } N = \bmx{N_{ux} & N_{u \omega} & N_{ub} \\ N_{vx} & N_{v \omega} & N_{vb}}
$$

Then the equilibrium point $\left( x_{*}, u_{*}, \nu_{*}, \omega_{*} \right)$ is computed,

\pagebreak
\printbibliography

\end{document}

### 2024-10-10 Theory recap/notes.tex
test

@@
 
-Then the equilibrium point $\left( x_{*}, u_{*}, \nu_{*}, \omega_{*} \right)$ is computed,
+Then the equilibrium point $\left( x_{*}, u_{*}, \nu_{*}, \omega_{*} \right)$ is computed.
+
+\subsection*{ETM definition}
+Different approaches were used in \cite{css-paper} and \cite{css-extended}, the first work takes into account the difference between the input and output vectors of the network at different time instants, the latter instead directly uses the sector conditions as the triggering ones.
 

### 2024-10-21 Idea to solve 1lin case/ideas.tex
Starting of discussion about 1D sat case

\documentclass{article}
\input{math_commands}
\input{custom_imports}
\usepackage[margin=54pt]{geometry}
\usepackage{biblatex}
\addbibresource{biblio.bib}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Solve LMI with saturation 1 dimensional case}
\maketitle

To solve the LMI for the plant:

\begin{equation}
  \bmx{
  \theta \\
  \dot{\theta}
  }^{+} = \bmx{
  1 & \delta t\\
  \frac{g}{l} dt & 1 - \frac{\mu}{m l^2} dt
  } \bmx{
  \theta \\
  \dot{\theta}
  } + \bmx{
    0 \\
    \frac{dt}{m l^2}
  } u = A x + B u
\end{equation}

And a controller of the form:

\begin{equation}
  u = \text{sat}(-K x)
\end{equation}

With $K$ the result of a LQR optimization, we need to solve a LMI. Defining the incremental state $\tilde{x} = x - x_{*}$, and the Lyapunov function $V(x) = x \tr P x$ I can write the increment of the Lyapunov function as:

\begin{multline}
\Delta V = (\tilde{x}^{+})\tr P \tilde{x}^{+} - \tilde{x}\tr P \tilde{x} = 
\left( A \tilde{x} + B \tilde{u} \right) \tr P ( \bullet )- \tilde{x}\tr P \tilde{x} = \\
\left( A \tilde{x} + B \left( N_{u \omega} \omega \right)\right) \tr P \left( \bullet \right) - \tilde{x}\tr P \tilde{x} = \\
\left( A \tilde{x} + B \left( \tilde{\nu} - \tilde{\psi} \right)\right) \tr P \left( \bullet \right) - \tilde{x}\tr P \tilde{x} = \\
\left( (A - B K) \tilde{x} - B \tilde{\psi} \right) \tr P \left( \bullet \right) - \tilde{x}\tr P \tilde{x} \\
\end{multline}

In matrix form denoting $A_{k} = A - B K$:

\begin{equation}
  \Delta V = \bmx{\tilde{x} \\ \tilde{\psi}} \tr \bmx{
    A_k\tr P A_k - P & -A_k\tr P B \\
    -B\tr P A_k & B\tr P B
  } \bmx{\tilde{x} \\ \tilde{\psi}}
\end{equation}

Using the same notation of \cite{css-extended} and \cite{css-paper} I can express the saturation involved in the controller as:

\begin{equation}
    \bmx{u\\ \nu} = \bmx{\begin{array}{c|c|c} 
        0 & \Id & 0 \\
        \hline
        -K & 0 & 0
    \end{array}}  \bmx{x \\ \omega \\ 1}
\end{equation}

And hence express the sector conditions of the saturation as:

\begin{equation*}
  \tilde{\psi}\tr T \left[ G \tilde{x} - \tilde{\nu} + \tilde{\psi} \right] \leq 0
\end{equation*}

\begin{equation}\label{sector}
  \bmx{\tilde{x} \\ \tilde{\psi}}\tr \bmx{
    0 & 0 & 0 \\
    TG & -T & T
  } \bmx{\tilde{x} \\ \tilde{\nu} \\ \tilde{\psi}} \leq 0 \to
  \bmx{\tilde{x} \\ \tilde{\psi}}\tr \bmx{
    0 & 0 & 0 \\
    TG & -T & T
  } \bmx{\Id & 0\\
  -K & 0\\
  0 & \Id} \bmx{\tilde{x} \\ \tilde{\psi}} \leq 0 \to \bmx{\tilde{x} \\ \tilde{\psi}}\tr M_{1} R_{\phi} \bmx{\tilde{x} \\ \tilde{\psi}}
\end{equation}

Given the sign definiteness of this quantity, I can bound $\Delta V$ as: 
\begin{equation}\label{matricione}
  \Delta V = \bmx{\tilde{x} \\ \tilde{\psi}}\tr \bmx{
    A_k\tr P A_k - P & -A_k\tr P B \\
    -B\tr P A_k & B\tr P B
  } -\text{He}\left\{ M_1 R_{\phi} \right\} \bmx{\tilde{x}\\ \tilde{\psi}} \leq 0 \to \bmx{\tilde{x} \\ \tilde{\psi}}\tr M \bmx{\tilde{x} \\ \tilde{\psi}} \leq 0
\end{equation}

Since the sector condition is not global I have to add the inclusion inside an ellipsoid to assure the local stability of the solution. I want $\tilde{x}\tr \alpha P \tilde{x} \leq 1$, from the sector condition definition, see set $\cS$ in \cite{css-extended} I have:

$$
  \cS = \left\{ x(k) \in \RR^{n_{p}} : -\bar{\nu} - \nu_* \leq G (x(k) - x_*) \leq \bar{\nu} - \nu_* \right\}
$$

Since $\nu_* = 0$ the condition directly translates into $|G \tilde{x}| \leq \bar{\nu} \to (G \tilde{x})\tr (G \tilde{x}) \leq \bar{\nu}^2 \to \tilde{x}\tr \frac{G\tr G}{\bar{\nu}^{2}} \tilde{x} \leq 1$ I want this condition to be included in the ellipsoid defined by $\cE(P, x_*) = \left\{ x \in \RR^{n_{F}}: \tilde{x}\tr \gamma P \tilde{x} \leq 1 \right\}$

\begin{multline*}
  \tilde{x}\tr \frac{G\tr G}{\bar{\nu}^{2}} \tilde{x} \leq \tilde{x}\tr \gamma P \tilde{x} \leq 1 \to \frac{G\tr G}{\bar{\nu}^{2}} \leq \gamma P \leq 1 \to \gamma P - \frac{G\tr G}{\bar{\nu}^{2}} \geq 0\\
  0
\end{multline*}

It is now possible to apply Schur's complement, dividing everything by $\gamma$:

\begin{equation}
  P - \frac{G\tr G}{\gamma \bar{\nu}^{2}} \geq 0 \to \bmx{P & G\tr\\ G & \gamma \bar{\nu}^{2}} \geq 0
\end{equation}

Now by pre and post multiplying by $\bmx{\Id & 0\\ 0 & T}$:

\begin{equation}\label{ellipsoid}
  \bmx{P & G\tr T\\
  TG & \gamma \bar{\nu}^{2} T^{2}} \geq 0
\end{equation}

Equation \ref{ellipsoid} is the last LMI condition to add. The problem arises due to the bilinear term $TG$ that appear also in equation \ref{sector}. The approach followed in \cite{css-extended} does not seem to be directly appliable to the one-dimensional case scenario even though it results to be compliant with a multi-layer network controller. New tricks have to be explored to solve this problem:

\subsection*{Iterative approach}
Set $G$ to nullify the cross term in matrix $M$ of equation \ref{matricione}. Solve for $P, T, \gamma$. Then set the found value of $T$ and solve for $P, G, \gamma$. Iterate until convergence to be established.

From equation \ref{matricione} the hermitian of the sector conditions translate as:

\begin{equation}
  M_1 R_{\phi} + R_{\phi}\tr M_1\tr = \bmx{
    0 & T(G\tr + K\tr)\\
    T(G + K) & 2T
  }  \to M = \bmx{
    A_k\tr P A_k - P & -A_k\tr P B + T(G\tr + K\tr)\\
    -B\tr P A_k + T(G + K) & B\tr P B + 2T
  }
\end{equation}

I set $G = -K - T^{-1} B\tr P A_k$ to nullify the cross term and start the iteration.

\subsection*{Scalar approach}
Follow the approach used in paper \cite{isabelle-paper} to find out how they dealt with $\gamma \bar{\nu}^{2} = \tau$ with \texttt{fminsearch}

\pagebreak
\printbibliography

\end{document}

### 2024-10-22 Automatica paper discussion/notes.tex
Folder order

\documentclass[12pt]{article}
\usepackage{float}
\usepackage{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% \setlength{\parindent}{0pt}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Notes on Event-triggered neural network control using quadratic constraints for perturbed systems paper}

\maketitle

From \textbf{Event-triggered neural network control using quadratic constraints for perturbed systems} paper it is useful to extract the application of the ETM mechanism directly inside the NN controller to decide whether to propagate an input. 
With respect to CSS paper there's a generalization of the activation function, no more restricted to saturation only and the structure of the ETM, error based in CSS and QC based here. Also, the system here is perturbed.

The plant to control is represented with $\mathbb{G}(G, \Delta)$ meaning the linear part $G$ and the non-linear perturbation part $\Delta$. It is described by:

\begin{equation}
  \begin{aligned}
    x(k+1) &= A_G x(k) + B_{G_u} u(k) + B_{G_q} q(k), \\
    p(k) &= C_G x(K) + D_{G_u} u(k) + B_{G_q} q(k),
  \end{aligned}
\end{equation}

With $p$ and $q$ being the input and output of the perturbation block $\Delta$: $q(\cdot) = \Delta(p(\cdot))$

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{img/scheme.png}
    \caption{Overall scheme of the controlled plant}
    \label{}
\end{figure}

\textbf{Notation on controller $\pi_{ETM}$}

The controller is a Feed Forward Neural Network of $l$ layers. With $\hat{w}$ it is intended the last transmitted output of a layer, with $\omega$ the current output of the layer and with $\nu$ the input to the activation function.

\begin{equation}
  \begin{aligned}
    \hat{\omega}^0(k) &= x(k), \\    
    \nu^i(k) &= W^i \hat{\omega}^{i-1}(k) + b^i, \\
    \omega^i(k) &= \phi^i( \nu^i(k)), \\
    u(k) &= W^{l+1} \hat{\omega}^l(k) + b^{l+1},
  \end{aligned}
\end{equation}

Where the input signal $u(k)$ is the output of the last layer and $b^i$ represents the bias vector of the $i^{th}$ layer.
Everything is vectorized so to isolate the non-linearities:

\begin{equation}
  \begin{bmatrix}
    u(k)\\
    \nu_{\phi}(k)
  \end{bmatrix} = N \begin{bmatrix}
    x(k)\\
    \hat{\omega}(k)\\
    1
  \end{bmatrix}
\end{equation}

Where $N$ is the matrix like in equation 8 of the paper.

Then follows the modelling of the perturbations, I won't treat it for now. Just to sum it up the perturbation block is modelled as a filter (equation. 9 and 10 in the paper) and the output of this filter $r(k)$ respects an Integral (a sum in DT) Quadratic Constraint, this means that the perturbation $\Delta$ must satisfy a specific energy-like inequality over time.

Then it is introduced the \textit{Local Sector Condition} used to deal with non-linear functions in a given interval. This will be used to treat the activation functions close the equilibrium. The non-linear function will be encapsulated in linear bounds that will then be directly injected in the stability LMI conditions.

Lemma 1 introduces a way to express the local sector condition for non-linearity $\phi$ in a compact form using the matrix $\Pi_{ds}$ called \textit{Diagonally structured multiplier} that belongs to a class of polytopic bounding multipliers, then another lemma is introduced for the latter.

After the applications of all these lemmas, equation 19 is obtained.

It is then defined the new ETM policy, that doesn't rely on the error like in the previous paper, the quadratic constraint instead measures the deviation of the pair $(\nu, \omega)$ from the equilibrium one, it represents if the system is behaving as expected or not and hence if it requires an update.

\begin{equation}
  \hat{\omega}^i(k) = \begin{cases}
    \hat{\omega}(k-1), & \text{if } [\bullet ]^{\top} \Pi_i \begin{bmatrix}
      \nu^i(k) - \nu_*^i \\
      \hat{\omega}^i(k-1) - \omega_*^i
    \end{bmatrix} \geq 0.\\
    \omega^i(k), & \text{otherwise}
  \end{cases}
\end{equation}

For all $i=1, \dots, l$.

Like in the previous paper it holds that if a layer is not updated all the following ones won't be also.

In the rest of the paper it was treated the problem of designing the triggering parameters that characterize $\Pi$ while assuring the stable behavior of the system. An augmented state vector that takes into account the perturbation has been used and in order to design the ETM matrix Theorem 1 has been stated and proved. Everything so that equation 26, the LMI condition for Lyapunov stability, holds. Adding another condition the LMI solution would give as a result also an estimate of the region of attraction, the problem is that this system is composed of infinite equations, to address this problem adding a constraint in $\Pi$ the number becomes finite. An optimization problem than has been developed to maximize the trade-off between size of the region of attraction and the saving in computation of the neural network.

Simulations were run for different values for both $\alpha$ and $\beta$ to better show how the performance change along the trade-off. It has been shown that for 4 layers architectures the computational saving can add up to 50\%.
\end{document}

% \begin{figure}[H]
%     \centering
%     \begin{subfigure}{0.4\textwidth}
%     \includegraphics[width=\textwidth]{}
%     \caption{}
%     \label{}
%     \end{subfigure}
%     \hfill
%     \begin{subfigure}{0.55\textwidth}
%     \includegraphics[width=\textwidth]{}
%     \caption{}
%     \label{}
%     \end{subfigure}
%     \caption{}
%     \label{}
% \end{figure}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=0.65\textwidth]{}
%     \caption{}
%     \label{}
% \end{figure}

### 2024-10-22 Dynamic ETM intro discussion/notes.tex
Folder order

\documentclass[12pt]{article}
\usepackage{float}
\usepackage{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{parskip}
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% \setlength{\parindent}{0pt}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Notes on Dynamic Triggering Mechanisms for Event-Triggered Control paper}

\maketitle

From \textbf{Dynamic Triggering Mechanisms for Event-Triggered Control} paper it is useful to extract the  dynamic ETM applied to the linear system case scenario:

$$
  \dot{x} = Ax + Bu, x \in \mathbb{R}^{n}, u \in \mathbb{R}^{m} 
$$

Assuming linear feedback controller $u = K x$ we obtain the ideal closed-loop system, we add the error considered in the paper e

$$
  \dot{x} = Ax + BK(x + e)
$$

That being asymptotically stable presents a Lyapunov function $V(x) = x^{\top} P x$ with $P$ p.d. and symmetric such That

$$
  (A + BK)^{\top}P + P(A + BK) = -Q
$$

We have then

\begin{multline*}
  \frac{d}{dt} V(x(t)) = \dot{x}^{\top} P x + x P \dot{x} = \\ 
  = (Ax + BKx + BKe)^{\top} P x + x^{\top} P (Ax + BKx + BKe) = \\
  = x^{\top} \left[ (A + BK)^{\top} P + P (A + BK) \right] x + x^{\top} P BKe + e^{\top} K^{\top} B^{\top} P x = \\
  \\
  \text{since the result is a scalar I can sum up the last two terms} \\
  \\
  = x^{\top} \left[ (A + BK)^{\top} P + P (A + BK) \right] x + 2 x^{\top} P B K e = - x^{\top} Q x + 2 x^{\top} P B K e
\end{multline*}

As a first approach the static ETM that follows can be used:

\begin{align} \label{static}
  t_0 &= 0, \nonumber \\
  t_{i+1} &= \text{inf} \left\{
  t \in \mathbb{R} | t > t_i \wedge \sigma x(t)^{\top} Q x(t) - 2 x(t)^{\top} P B K e(t^{-} ) \leq 0
  \right\}
\end{align}

This ensures that the next event will occur whenever the derivative of the Lyapunov function will cease to be negative. Ensuring a forced definite negativeness.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{img/static}
    \caption{It is clear how the Lyapunov function $V(x)$ is strictly decreasing}
    \label{static-plot}
\end{figure}

Now a dynamic ETM is proposed based on the dynamic variable $\eta$ governed by the differential equation

\begin{equation}
  \dot{\eta} = - \lambda \eta + \sigma x^{\top} Q x - 2 x^{\top} P B K e, \eta(0) = \eta_0
\end{equation}

With $\sigma \in \left( 0, 1 \right), \lambda > 0, \eta_0 \in \mathbb{R}^{+}_0$ the design parameters

It is important noticing how this addition of dynamic changes the overall behavior of the ETM mechanism. In the static case (equation. \ref{static}) the value is updated every time the condition is violated. By adding a dynamic term the update will not be immediate, $\eta(t)$ integrates the previous triggering signal over time rather than responding to the current value of the signal. The term $-\lambda \eta(t)$ makes $\eta(t)$ decay over time to that it doesn't accumulate the signal history indefinitely, it smoothly adjusts based on the previous value of the signal. Overall it's a filter of the previous signal since it gives a "smoothed" version of the signal used for the static ETM.

It is then defined the new dynamic ETM:

\begin{align} \label{dynamic}
  t_0 &= 0, \nonumber \\
  t_{i+1} &= \text{inf} \left\{
  t \in \mathbb{R} | t > t_i \wedge \eta(t) + \theta \left(  \sigma x(t)^{\top} Q x(t) - 2 x(t)^{\top} P B K e(t^{-} ) \right) \leq 0
  \right\}
\end{align}

It has been added the parameter $\theta \in \mathbb{R}^{+}_0$. The static ETM (equation. \ref{static}) can now be seen as the limit case scenario of the dynamic one (equation. \ref{dynamic}) where $\theta$ tends to $+ \infty$ and hence the $\eta(t)$ term can be ignored.

It is then proved how the inter-execution times with the dynamic ETM are larger with respect to the static ETM use. A new candidate Lyapunov function $W : \mathbb{R}^n \times \mathbb{R}_0^+ \to \mathbb{R}_0^+$ designed for the augmented system that sees the introduction of $e$ and $\eta$ variables:

\begin{equation}
  W(x, \eta) = V(x) + \eta 
\end{equation}

\begin{equation}
  \frac{d}{dt} W(x(t), \eta(t)) = \left( \sigma - 1 \right) x(t)^{\top} Q x(t) - \lambda \eta(t)
\end{equation}

That with the previous assumptions (not declared here but in the paper) that $\sigma \in \left( 0,1 \right)$ and $\eta(t) \geq 0$ with $\lambda > 0$ Assures that $\dot{W}$ stays definite negative and both $x(t)$ and $\eta(t)$ converge asymptotically to the origin.

It's then treated how each of the design parameters influence the overall system behavior. 3 different cases were portrayed where an explicit expression of the inter-execution time $\tau$ is drawn out.

It follows the treatment of the Lyapunov function decay rate, it is strongly influenced by $\sigma$, not influenced by $\theta$ and to understand the influence of this last term it has been used the \textit{Quadratic Integral Performance Index}.

To choose the parameters it was first tuned the Lyapunov function decay rate by imposing $\eta_0 = 0$ and $\lambda = (1 - \sigma) \kappa$ (what this $\kappa$ means it is explained in the paper), here $\sigma$ represents the degradation of Lyapunov function decay with respect to the ideal close-loop system. It was then chosen $\theta$ taking into account the considerations on $\tau$ formulation.

\textbf{Simulation results}

In comparison to the results of figure (fig. \ref{static-plot}) the dynamic ETM has been applied in the following:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{img/dynamic}
    \caption{}
    \label{dynamic-plot}
\end{figure}

It was also done the comparison with a dynamic ETM taken from one of the references that worked as follows:

\begin{align}
  t_0 &= 0 \nonumber \\
  t_{i+1} &= \text{inf} \left\{ t \in \mathbb{R} | t > t_i \wedge V(x(t)) \geq e^{\left( \sigma - 1 \right) \kappa \left( t - t_i \right)} V(x(t_i)) \right\}
\end{align}

In this case the update will occur every time the value of the Lyapunov function will exceed the one of the exponential as shown in the following:

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{img/reference}
    \caption{}
    \label{reference-plot}
\end{figure}

In the final comment it's shown how the approach of this paper puts itself as a good compromise between the good monotonic behavior of the static ETM and the large inter-execution times of the reference approach (fig. \ref{reference-plot}) that, however, exhibit large variations in $V$ value.

Also, it is shown that using a dynamic ETM the coefficient of variability of the inter-execution times is reduced by a lot, meaning that the controller acts in a more predictable way.

\end{document}

% \begin{figure}[H]
%     \centering
%     \begin{subfigure}{0.4\textwidth}
%     \includegraphics[width=\textwidth]{}
%     \caption{}
%     \label{}
%     \end{subfigure}
%     \hfill
%     \begin{subfigure}{0.55\textwidth}
%     \includegraphics[width=\textwidth]{}
%     \caption{}
%     \label{}
%     \end{subfigure}
%     \caption{}
%     \label{}
% \end{figure}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=0.65\textwidth]{}
%     \caption{}
%     \label{}
% \end{figure}

### 2024-10-22 Journal/notes.tex
Folder order

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage[margin=54pt]{geometry}
\usepackage{biblatex}
\addbibresource{../biblio.bib}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Journal}
\maketitle

Ordered list of problems and steps taken to solve them, or at least to understand them better.

\section*{Initial works}

The first weeks were spent understanding the papers and the Matlab code given by Sophie (\cite{automatica}, \cite{css-paper}, \cite{data-driven}) regarding the application of ETM to the NN controller of a linearized and non linearized pendulum. The aim is to save as much computational load as possible by putting ETM in between the layers. Doing so it is possible to decide whether propagating the current layer output to the activation function. 
The main computational save will regard directly the application of the activation function that, being non-linear, is the most expensive operation in the network.

\subsection*{Faulty application of global sector conditions}

\section*{Reproducing the results of the papers}

\section*{Adding the conditions for dynamic ETM}

\section*{Problem in the extension to custom trained NN controllers}

\section*{Idea to use Samuele's results to address robustness of a NN controlled system with integrator in DT}

\section*{Development of new controllers}

\subsection*{Reinforcement learning approach}

\subsection*{Deep learning approach of LQR controller}

\section*{Problems in the addition of system non-linearities}

\subsection*{Progressive simplification of the problem to understand the issue}

\subsection*{1 layer saturation controller to linear pendulum}

\subsection*{1 layer saturation controller to linear pendulum with integrator in the state}

\pagebreak
\printbibliography

\end{document}

### 2024-10-22 Linear integrator 1 dimension case/ideas.tex
Folder order

--- 2024-10-21 Idea to solve 1lin case/ideas.tex
@@
 \documentclass{article}
-\input{math_commands}
-\input{custom_imports}
+\input{../math_commands}
+\input{../custom_imports}
 \usepackage[margin=54pt]{geometry}
 \usepackage{biblatex}
-\addbibresource{biblio.bib}
+\addbibresource{../biblio.bib}
 
@@
 
+The ellipsoid condition without bilinear terms is:
+
+\begin{equation}
+  \bmx{
+    P & G\tr \\
+    G & \gamma \bar{\nu}^{2}
+  } \geq 0
+\end{equation}
+
+The problem that arises now is that imposing $G = -K - T^{-1} B\tr P A_k$ results in a bilinear condition since we have both $T^{-1}$ and $P$
+
 \subsection*{Scalar approach}
+
 Follow the approach used in paper \cite{isabelle-paper} to find out how they dealt with $\gamma \bar{\nu}^{2} = \tau$ with \texttt{fminsearch}
+
+\textbf{Paper analysis}: \dots Interrupted immediately since I found a value for $\alpha$ with respect to equation \ref{ellipsoid} and paper \cite{css-extended}: $\alpha = 1$\dots
+
+\section*{Pendulum with integrator}
+Since I found the solution for the LMI for the case without integrator, I can now try to solve the case with integrator. The plant is now:
+
+\begin{equation}
+  \bmx{
+  \theta \\
+  \dot{\theta} \\
+  \eta
+  }^{+} = \bmx{
+  1 & \delta t & 0\\
+  \frac{g}{l} dt & 1 - \frac{\mu}{m l^2} dt & 0\\
+  1 & 0 & 1
+  } \bmx{
+  \theta \\
+  \dot{\theta} \\
+  \eta
+  } + \bmx{
+    0 \\
+    \frac{dt}{m l^2} \\
+    0
+  } u = A x + B u
+\end{equation}
+
+Following analogous considerations I end up with a condition formally identical as in equation.
+\ref{matricione}.
+The only difference is that the dimensions of the matrices that now are scaled accordingly and the gain matrix $K$ that is now the proper LQR gain for the new system with integrator.
+I set up the ellipsoid condition as in equation \ref{ellipsoid} with $\gamma = 1$.
+Introducing the variable $Z = TG$ and exploiting the result from:
+
+$$
+  \left[ \bar{\nu}^{-2} - T \right] \bar{\nu}^{2} \left[ \bar{\nu}^{-2} - T \right] \geq 0 \to \bar{\nu}^{-2} - T - T + \bar{\nu}^{2} T^{2} \geq 0 \to \bar{\nu}^{2} T^{2} \geq 2 T - \bar{\nu}^{-2}
+$$
+
+Equation. \ref{ellipsoid} becomes:
+
+\begin{equation}
+  \bmx{
+    P & Z\tr \\
+    Z & 2 T - \bar{\nu}^{-2} 
+  } \geq 0
+\end{equation}
+
+That guarantees the validity in a local region of the state space.
+Specifically the region defined by the ellipsoid $\cE(P, x_*) = \left\{ x \in \RR^{n_{F}} : (x - x_*)\tr P (x - x_*) \leq 1 \right\}$ that is represented in the following plots.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.48\textwidth}
+    \includegraphics[width=\textwidth]{img/red_state_out}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
+    \includegraphics[width=\textwidth]{img/lyap_out}
+    \end{subfigure}
+    \caption{Initial state out of ROA, the system is still able to reach the origin but the Lyapunov function is not strictly decreasing}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.48\textwidth}
+    \includegraphics[width=\textwidth]{img/red_state_inside}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
+    \includegraphics[width=\textwidth]{img/lyap_inside}
+    \end{subfigure}
+    \caption{Initial state inside the ROA, the system is able to reach the origin and the Lyapunov function is strictly decreasing}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{img/detail_lyap}
+    \caption{Detail of the Lyapunov function in the case the initial state is inside the ROA}
+\end{figure}
+
+\subsection*{Next steps}
+Now the following steps regard the extraction of the "region of linearity" from this trivial 1 layer case and extend it iteratively to craft a multi-layer network controller by hand.
 

### 2024-10-22 Sector conditions intro discussion/notes.tex
Folder order

\documentclass[12pt]{article}
\usepackage{float}
\usepackage{graphicx}
\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{parskip}
\usepackage[a4paper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% \setlength{\parindent}{0pt}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Notes on Sector condition study}
\maketitle

Sector conditions are used to handle non-linearity in linear systems (activation functions in my case). From the book of Sophie I will take the basic concepts, firstly referred to the Dead-zone non-linearity and then proceeding to the general case.

Defining the Dead-zone non-linearity:

\begin{equation} \label{non-l}
  \phi(v(t)) = sat(v(t)) - v(t) = \begin{cases}
    u_{max(i)} - v_{(i)} & \text{if } v_{(i)} > u_{max(i)}\\
    0 & \text{if } -u_{min(i)} \leq v_{(i)} \leq u_{max(i)}\\
    -u_{min(i)} - v_{(i)} & \text{if }  v_{(i)} < -u_{min(i)}
  \end{cases}
\end{equation}

With this non-linearity it's possible to explicit the global sector conditions:

\textbf{Lemma 1} \textit{For all $v \in \mathbb{R}^n$, the non-linearity $\phi(v)$ satisfies the following inequality:}
\begin{equation} \label{global-quad}
  \phi(v)^{\top} T \left( \phi(v) + v \right) \leq 0
\end{equation}
\textit{for any diagonal positive matrix $T \in \mathbb{R}^{n \times n}$}

\textit{Proof} the sign of \ref{global-quad} will be discussed
\begin{itemize}
  \item If $v > u_{max}$ we have $\phi(v) < 0$ and $\phi(v) + v = u_{max} > 0$ hence we have the overall product $< 0$
  \item if $v < - u_{min}$ we have $\phi(v) > 0$ and $\phi(v) + v = -u_{min} < 0$ hence we have the overall product $< 0$
  \item if $-u_{min} < v < u_{max}$ we have $\phi(v) = 0$ hence the product will always be $0$
\end{itemize}
This proves that the product is always $\leq 0$

The general form refers to the elements inside the set:
\begin{equation} \label{s-cond}
  S(v - \omega, u_{min}, u_{max}) = \left\{ v \in \mathbb{R}^n ; \omega \in \mathbb{R}^n; - u_{min} \leq v - \omega \leq u_{max} \right\}
\end{equation}
Then the following inequality is satisfied $\forall T \in \mathbb{R}^{n \times n}$ diagonal positive definite
\begin{equation} \label{sec-cond-0-1j}
  \phi(v)^{\top} T \left( \phi(v) + \omega \right) \leq 0
\end{equation}

\textit{Proof} By exploiting condition \ref{s-cond} the proof is analogous.

It is a powerful expression since we can substitute $v$ with other expressions like for example the feedback law $v(t) = K x(t)$ with a dynamic like $\dot x = A x + B \text{ sat}(K x) = (A + BK) x + B \phi(K x(t))$ with $\phi(K x) = \text{sat} (K x) - Kx$ obtaining something like
$$
  \phi(Kx)^{\top} T \left( \phi(Kx) + Gx \right) \leq 0
$$
For every $x$ in the polyhedral set:
$$
S(K - G, u_{min}, u_{max}) = \left\{ x \in \mathbb{R}^n; -u_{min} \leq \left( K - G \right) x \leq u_{max}\right\}
$$

Like already seen for the global sector condition with $S, T, R$ these conditions are useful to inject into the discussion of the Lyapunov function incremental sign since it's a term whose sign is always defined. Doing so we take into account the non-linearity of the system, and we can proceed with the LMI formulation.
$$
  V(x) = x^{\top} P x, P = P^{\top} > 0
$$
\begin{equation} \label{lyap-cond}
  \dot V (x) \leq \dot V (x) - 2 \phi(K x)^{\top} T (\phi(K x) + G x) \forall x \in \mathcal{E}(P, 1)
\end{equation}
Note that $\mathcal{E}(P, 1)$ refers to an ellipsoid that is the current region of asymptotic stability (RAS). It is included in the LMI problem via a condition similar to this:
\begin{equation}
  \begin{bmatrix}
    W & W K^{\top} - Z^{\top}\\
    \star & u_{0}^2
  \end{bmatrix} \geq 0
\end{equation}
With $ P = W^{-1}, Z = GW, S = T^{-1}$. Usually we pre multiply and post multiply equation \ref{lyap-cond} by $\begin{bmatrix}
  P^{-1} & 0\\
  0 & T^{-1}
\end{bmatrix}$, and we change the variables we obtain something like this
$$
  \left[ x^{\top} \phi(K x)^{\top} \right] \begin{bmatrix}
    W(A + BK)^{\top} + (A + BK) W & BS-WK^{\top} - Z^{\top}\\
    \star & - 2S
  \end{bmatrix} \begin{bmatrix}
    x\\
    \phi(Kx)
  \end{bmatrix} < 0
$$

\section*{Arcak paper discussion}
The local sector constraints are considered with the offset to the equilibrium points $(\nu_*, \phi(\nu_*))$ for each single activation function:

Let $\alpha_{\phi}, \beta_{\phi}, \underline{\nu}, \overline{\nu}, \nu_* \in \mathbb{R}^{n_{\phi}}$ be given with $\alpha_{\phi} \leq \beta_{\phi}$, $\underline{\nu} \leq \nu_* \leq \overline{\nu}$, $\omega_* := \phi(\nu_*)$. Assuming $\phi$ satisfies the offset local sector $\left[ \alpha_{\phi}, \beta_{\phi} \right]$ around $(\nu_*, \omega_*)$ element wise for all $\nu_{\phi} \in [\underline{\nu}, \overline{\nu}]$. If $\lambda \in \mathbb{R}^{n_{\phi}}$ with $\lambda \geq 0$ then:
\begin{equation}
  \begin{bmatrix}
    \nu_{\phi} - \nu_*\\
    \omega_{\phi} - \omega_*
  \end{bmatrix}^{\top} \Psi_{\phi}^{\top} M_{\phi}(\lambda) \Psi_{\phi} \begin{bmatrix}
    \nu_{\phi} - \nu_*\\
    \omega_{\phi} - \omega_*
  \end{bmatrix} \geq 0\ \ \forall \nu_{\phi} \in [\underline{\nu}, \overline{\nu}],\ \omega_{\phi} = \phi(\nu_{\phi})
\end{equation}

where
$$
\Psi_{\phi} := \begin{bmatrix}
  \textit{diag}(\beta_{\phi}) & - I_{\phi}\\
  - \textit{diag}(\alpha_{\phi}) & I_{\phi}
\end{bmatrix}
$$
and
$$
M_{\phi} (\lambda) := \begin{bmatrix}
  0_{n_{\phi}} & \textit{diag} (\lambda)\\
  \textit{diag} (\lambda) &  0_{n_{\phi}}
\end{bmatrix}
$$

By putting into explicit form this product we obtain:
$$
  \sum_{i=1}^{n_{\phi}} \lambda_i \left( \Delta \omega_i - \alpha_i \Delta \nu_i \right) \cdot \left( \beta_i \Delta \nu_i - \Delta \omega_i \right)
$$
Which is the offset local sector condition applied for each activation function in the NN.

Another interesting thing is the computation of vectors $\underline{\nu}, \overline{\nu}$. Essentially we choose $\underline{\nu}^1, \overline{\nu}^1$ with $\nu_*$ in between. Then we compute $[\underline{\omega}^1, \overline{\omega}^1]$ from the output of $\omega^1 = \phi^1(\nu^1)$ that can be used to compute the bounds $\underline{\nu}^2, \overline{\nu}^2$ and so on. To sum up everything is dependent on the initial choice of $\underline{\nu}^1, \overline{\nu}^1$. This is important for the later estimation of the ROA: decreasing $(\overline{\nu}^1 - \nu_*^1)$ if beneficial for the ROA estimation but restricts the region where ROA inner approximations lie. The opposite if we increase $\left( \overline{\nu}^1 - \nu_*^1 \right)$, it has been chosen to parametrize as $\delta$ this quantity and choose $\delta$ that leads to the larges inner approximation.

\section*{Static ETM paper discussion}

\section*{Appendix}
\begin{itemize}
  \item Lipschitz constant of NN: specifies how much the output of the network can change with respect to changes in the input. It is Lipschitz continuous if $\exists L \geq 0$ such that $\forall x_1, x_2$:
  $$
    || f(x_1) - f(x_2)|| \leq L || x_1 - x_2||
  $$
  \item General global sector condition: Let $\alpha \leq \beta$, the function $\phi : \mathbb{R} \to \mathbb{R}$ lies in the global sector $[\alpha, \beta]$ if:
  $$
    (\phi(\nu) - \alpha \nu) \cdot (\beta \nu - \phi(\nu)) \geq 0\ \ \forall \nu \in \mathbb{R}
  $$
  Note how this condition can be brought back to  eq\ref{sec-cond-0-1j} that is the sector condition $[0, -1]$ by imposing $\alpha = 0, \beta = -1$:
  $$
    \phi(\nu) \cdot (-\nu - \phi(\nu)) \geq 0 \to \phi(\nu) \cdot (\nu + \phi(\nu)) \leq 0
  $$
  For the one dimensional case, brought to multidimensional case with $T$ diagonal and positive definite
  \item Offset local sector: Let $\alpha, \beta, \underline{\nu}, \overline{\nu}, \nu_*$ be given with $\alpha \leq \beta$ and $\underline{\nu} \leq \nu_* \leq \overline{\nu}$. The function $\phi : \mathbb{R} \to \mathbb{R}$ satisfies the offset local sector $[\alpha, \beta]$ around the point $\left( \nu_*, \phi(\nu_*) \right)$ if 
  $$
    \left( \Delta \phi(\nu) - \alpha \Delta \nu \right) \cdot \left( \beta \Delta \nu - \Delta \phi(\nu) \right) \geq 0\ \ \forall \nu \in [\underline{\nu}, \overline{\nu}]
  $$
  where $\Delta \phi(\nu) := \phi(\nu) - \phi(\nu_*)$ and $\Delta \nu := \nu - \nu_*$
\end{itemize}

\end{document}

% \begin{figure}[H]
%     \centering
%     \begin{subfigure}{0.4\textwidth}
%     \includegraphics[width=\textwidth]{}
%     \caption{}
%     \label{}
%     \end{subfigure}
%     \hfill
%     \begin{subfigure}{0.55\textwidth}
%     \includegraphics[width=\textwidth]{}
%     \caption{}
%     \label{}
%     \end{subfigure}
%     \caption{}
%     \label{}
% \end{figure}

% \begin{figure}[H]
%     \centering
%     \includegraphics[width=0.65\textwidth]{}
%     \caption{}
%     \label{}
% \end{figure}

### 2024-10-22 Journal/notes.tex
Added last resource

@@
 The main computational save will regard directly the application of the activation function that, being non-linear, is the most expensive operation in the network.
+Some sum ups can be found in the folders \textit{"Automatica paper discussion", "Dynamic ETM intro discussion", "Sector conditions intro discussion"}.
 
@@
 
+I tried to apply directly the proposition of Sophie's notes to the system portrayed in \cite{css-paper}.
+Everything is documented in the notes \textit{"Faulty implementation of dynamic ETM report"} in \textit{"extra material"} folder.
+
 \section*{Reproducing the results of the papers}
+At first, I tried t
 

### 2024-10-22 Journal/notes.tex
Further update of Journal

@@
 \section*{Reproducing the results of the papers}
-At first, I tried t
+
+A first attempt into the implementation has been portrayed in Matlab and immediately after in Python.
+The most used modules were \texttt{cvxpy} as a framework for LMI resolution and \texttt{NumPy} for generic mathematical manipulation.
+
+\subsection*{Linear Inverted Pendulum}
+The system used in \cite{css-paper} has been implemented in Python.
+The same weights used in the paper have been used to build up the NN controller.
+The main module exploited is \texttt{PyTorch}, the activation functions used are saturations.
+Tweaking the LMI conditions I was able to obtain the same results of the paper with the same basins of attraction.
+The main concern about the approach followed in the paper regards the ETM mechanism design:
+$$
+  \hat{\omega}^{i}(k) = \begin{cases}
+    \omega^{i}(k) & \text{if } f^{i}(\omega^{i}(k), \hat{\omega}^{i}(k-1), \hat{\omega}^{i-1}(k)) \geq 0\\
+    \hat{\omega}^{i}(k-1) & \text{otherwise}  
+  \end{cases}
+$$
+Where 
+$$ 
+f^{i}(\omega^{i}(k), \hat{\omega}^{i}(k-1), \hat{\omega}^{i-1}(k)) = ||\hat{\omega}^{i}(k-1) - \omega^{i}(k)||^{2}_{q^{i}_{\Delta}} - ||\omega^{i}(k) - \omega^{i}_*||^{2}_{q^{i}_{\omega}} - ||\hat{\omega}^{i-1}(k) - \omega^{i-1}_*||^{2}_{q^{i-1}_{\hat{\omega}}}
+$$
+The good thing about this function is the capability to stop early the computation and avoid computing the remaining layer if one doesn't trigger.
+However, this capability looses meaning when applied to a 2 layer NN controller.
+The bad point lies instead in the term $\omega^{i}(k) = sat(\nu^{i}(k))$ that is required at each layer verification.
+The need to compute this term, hence computing the activation function, makes the ETM useless if not for the cascade effect that remains good but not enough to be worth the computational load.
+
+Nevertheless, I was able to reproduce the results for the static ETM, and I was able to extend the problem to a dynamic ETM inspired by the work of \cite{data-driven}.
 
 \section*{Adding the conditions for dynamic ETM}
+The main logic was to update the data used to compute the new input to the system only when necessary:
+$$
+  \pmx{\chi \\ s} = \begin{cases}
+    \pmx{\omega\\ k} & \text{if the memory is updated}\\ 
+    \pmx{\chi_{k-1} \\ s_{k-1}} & \text{otherwise}
+  \end{cases}
+$$
+
+With the ETM mechanism described as follow:
+
+$$
+s^{+} = \text{min}_{m \in \NN} \left\{ m \geq s + 1 | \psi(\omega_m, \chi) \geq \rho \eta_m \right\}
+$$
 
@@
 \pagebreak
+\emergencystretch=2em % To account for hbox overflow warning
 \printbibliography

### 2024-10-23 Journal/notes.tex
Progress on Journal

@@
 
-With the ETM mechanism described as follow:
+With the ETM mechanism described as follows:
 
@@
 $$
+
+And $\eta$ dynamic is defined as:
+
+$$
+  \eta^{+} = (\rho + \lambda) \eta - \psi(\omega_m, \chi)
+$$
+
+This formulation ensures $\eta$ to be always positive and made it possible to consider it as a positive system like in \cite{positive-systems}. 
+As a quick proof I can say that if the triggering condition $\psi$ is negative then the positiveness of $\eta$ is trivial. If instead $\psi$ is positive and $\psi \geq \rho \eta $
 
@@
 
+It works by putting $\alpha = 1$
+
 \subsection*{1 layer saturation controller to linear pendulum with integrator in the state}
+
+It works by putting $\alpha = 1$
+
+\subsection*{Attempt on handcrafting NN}
+Considerations on it can be found in the last subsection of \textit{"Linear integrator 1 dimension case"}.
+I am now trying to train a 3-layer NN with deep learning method on a K matrix found with LQR method and not pole placement. It will then be attempted to:
+\begin{itemize}
+  \item Find a LMI solution for the system with integrator and non-linearity for the 3 layer network. If it doesn't work simplify the problem and add complexity bit by bit
+  \item Test the K gain from LQR method to non-linear system with and without saturation
+  \item Find the LMI solution for the non-linear system with input saturation: $u = sat(K_{LQR} \cdot x)$
+\end{itemize}
+
+\subsection*{1 layer saturation controller to non-linear pendulum with integrator in the state}
+The system under analysis is represented by the dynamic:
+
+$$
+m l^{2} \ddot{\theta} = m g l\ sin(\theta) - \mu \dot{\theta} + u
+$$
+
+That discretized becomes:
+
+$$
+  \begin{cases}
+    \theta^{+} = \theta + \dot{\theta} \delta t\\
+    \dot{\theta}^{+} = \dot{\theta} + \ddot{\theta} \delta t = \dot{\theta} + (\frac{g}{l} sin(\theta) - \frac{\mu}{m l^{2}} \dot{\theta}) \delta t + \frac{dt}{m l^{2} u} \\
+    \eta^{+} = \eta + y = \eta + (\theta - r)
+  \end{cases}
+$$
+
+I want to deal with the non-linearity in the form of $sin(\theta) - \theta$.
+In order to do so, I add and subtract the term $\frac{g}{l} \cdot dt \cdot \theta$ to $\dot{\theta}$ dynamics.
+Then by expressing $\phi(\theta) = sin(\theta) - \theta$ I can rewrite the overall dynamics as:
+
+\begin{equation}
+  \bmx{
+    \theta^{+} \\
+    \dot{\theta}^{+} \\
+    \eta^{+}
+  } = \bmx{
+    1 & \delta t & 0 \\
+    \frac{g \cdot dt}{l} & 1 - \frac{\mu \cdot dt}{m l^{2}} & 0 \\
+    1 & 0 & 1
+  } \bmx{
+    \theta \\
+    \dot{\theta} \\
+    \eta
+  } + \bmx{
+    0 \\
+    \frac{dt}{m l^{2}} \\
+    0
+  } u + \bmx{
+    0 \\
+    \frac{g \cdot dt}{l} \\
+    0 
+  } \phi 
+\end{equation}
+
+With more condensed notation:
+$$
+  x^{+} = A x + B u + C \phi
+$$
+
+The sector conditions on the saturation are the same of \cite{css-extended} adapted for the single saturation case:
+$$
+N = \bmx{
+\begin{array}{c|c|c}
+  0 & 1 & 0 \\
+  \hline
+  K & 0 & 0
+\end{array}
+}
+$$
+
+Each element is null except for $N_{u \omega} = 1$ and $N_{\nu x} = K$ so to have:
+\begin{align*}
+  \nu &= N_{\nu x} \cdot x + N_{\nu \omega} \cdot \omega + N_{\nu b} = K x\\
+  u &= N_{ux} \cdot x + N_{u \omega} \cdot \omega + N_{u b} = \omega = sat(\nu)
+\end{align*}
+
+In the end the control input is $u = sat(\nu) = sat(K x)$
+
+I consider a quadratic Lyapunov function $V = \tilde{x}\tr P \tilde{x}$ and I want to bound $\Delta V$ to be strictly negative.
 

### 2024-10-23 Linear integrator 1 dimension case/ideas.tex
Progress on Journal

@@
 
+We have explored some way to craft it. The problem lies in the dimensionality of the matrices, given the feasibility of the 1 dimensional case:
+
+$$
+  u = sat(Kx)
+$$
+
+We wanted to extend the problem to an $l$ layer controller by designing something like:
+
+$$
+  u = sat_1 \left( W_1 \left( sat_0 \left( W_0 x \right) \right) \right)
+$$
+
+Such that, staying in the linear zone of the saturation, we have $u = W_1 W_0 x = K x$ and to design something like $W_1 = k W_0^{-1}$ with $W_0$ and $W_1$ to be defined.
+Furthermore, we have to ensure that after the first layer we keep staying inside the linearity region.
+To do so we have to force something like this:
+
+$$
+  -\bar{\nu}_i \leq W_i \nu \leq \bar{\nu}_i
+$$
+
+For each layer, this directly translates to the inclusion of vector $\nu$ to the following set: \\ $ S_i = \left\{ \nu \in \RR^{n_{\phi}} : \nu\tr W_i\tr W_i \nu < \bar{\nu}^{2} \right\}$
+Then we can express the ellipsoid conditions:
+
+$$
+  \nu\tr \frac{W_i\tr W_i}{\bar{\nu}^{2}} \nu \leq 1
+$$
+
+Since we want the inclusion of one linearity zone into the following we impose for every layer:
+
+$$
+  \nu\tr \frac{W_{i-1}\tr W_{i-1}}{\bar{\nu}_{i-1}^{2}} \nu \geq \nu\tr \frac{W_{i}\tr W_{i}}{\bar{\nu}_{i}^{2}} \to W_{i-1}\tr W_{i-1} - W_{i}\tr W_{i} \frac{\bar{\nu}_{i-1}^{2}}{\bar{\nu}_{i}^{2}} \geq 0
+$$
+
+Applying Schur's complement:
+
+$$
+  \bmx{
+    W_{i-1}\tr W_{i-1} & W_{i}\tr \\
+    W_{i} & \frac{\bar{\nu}_{i}^{2}}{\bar{\nu}_{i-1}^{2}} \Id
+  } \geq 0
+$$
+
+Everything seems to be fine by fixing $\bar{\nu}_0$ equal to the actuator saturation level, but the problem lies in the dimensionality of the matrices $W_i$ since we need their inverse to compute the matrix for the following layer and to be invertible they need to be square.
+This enforces a strong limit on the architecture of the network since it forces to have the dimension of the input, the state, equal to the number of each hidden layer and also to the dimension of the output.
+Hence, this method is not applicable to the general case of a multi-layer network controller.
+
 \pagebreak

### 2024-10-24 Journal/notes.tex
Journal up to date

@@
 I consider a quadratic Lyapunov function $V = \tilde{x}\tr P \tilde{x}$ and I want to bound $\Delta V$ to be strictly negative.
+Expressing $\tilde{x}_*^{+}$ in a compact way:
+
+$$
+  \tilde{x}^{+} = A x + B u + C \phi - A x_* - B u_* - C \phi_* = A \tilde{x} + B \tilde{u} + C \tilde{\phi}
+$$
+
+In this expression $\tilde{u}$ can be expressed with respect to $\omega$ and $x$:
+$$
+  \tilde{u} = N_{ux} x + N_{u \omega} + N_{ub} - N_{ux} x_* - N_{u \omega} \omega_* - N_{ub} = N_{ux} \tilde{x} + N_{u \omega} \tilde{\omega}
+$$
+
+I consider now the change of variables $\psi = \nu - \omega$ to make the problem compliant with the sector conditions used in \cite{css-extended}:
+
+\begin{align*}
+  \tilde{\nu} &= R N_{\nu x} \tilde{x} \\
+  \tilde{u} &= N_{ux} \tilde{x} + N_{u \omega} (\tilde{\nu} - \tilde{\psi}) = N_{ux} \tilde{x} + N_{u \omega} R N_{\nu x} \tilde{x} - N_{u \omega} \tilde{\psi} = (N_{ux} + N_{u \omega} R N_{\nu x}) \tilde{x} - N_{u \omega} \tilde{\psi} = R_{\omega} \tilde{x} - N_{u \omega} \tilde{\psi} \\
+  \tilde{x}^{+} &= (A + B R_{\omega}) \tilde{x} - B N_{u \omega} \tilde{\psi} + C \tilde{\phi} = \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi}
+\end{align*}
+
+With \begin{itemize}
+  \item $R = \left( \Id - N_{\nu \omega} \right)^{-1}$
+  \item $R_{\omega} = N_{ux} + N_{u \omega} R N_{\nu x}$
+  \item $\bar{A} = A + B R_{\omega}$
+  \item $\bar{B} = - B N_{u \omega}$
+\end{itemize}
+
+I can express then:
+
+\begin{align*}
+  \Delta V &= (\tilde{x}^{+})\tr P \tilde{x}^{+} - \tilde{x}\tr P \tilde{x} = \left( \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi} \right)\tr P \left( \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi} \right) - \tilde{x}\tr P \tilde{x} = \\
+  &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr \bmx{
+    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C \\
+    \bar{B}\tr P \bar{A} & \bar{B}\tr P \bar{B} & \bar{B}\tr P C \\
+    C\tr P \bar{A} & C\tr P \bar{B} & C\tr P C
+  } \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }
+\end{align*}
+
+I have to include now the sector conditions to deal with the non-linearity of the activation functions of the NN controller.
+In this particular case we deal with only one saturation function.
+Following the approach in \cite{css-extended} I can express the sector condition as:
+
+$$
+  \tilde{\psi}\tr T \left[ G \tilde{x} - \tilde{\nu} + \tilde{\psi} \right] \leq 0 \to \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr \bmx{
+    0 & 0 & 0 & 0 \\
+    TG & -T & T & 0 \\
+    0 & 0 & 0 & 0
+  } \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\nu} \\
+    \tilde{\phi}
+  } \to \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr \bmx{
+    0 & 0 & 0 & 0 \\
+    TG & -T & T & 0 \\
+    0 & 0 & 0 & 0
+  } \bmx{
+    \Id & 0 & 0 \\
+    R N_{\nu x} & \Id - R & 0 \\
+    0 & \Id & 0  \\
+    0 & 0 & \Id
+  } \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }
+$$
+
+Having written $\tilde{\nu} = N_{\nu x} \tilde{x} + N_{\nu \omega} \tilde{\omega} = N_{\nu x} \tilde{x} + N_{\nu \omega} (\tilde{\nu} - \tilde{\psi}) \to (\Id - N_{\nu \omega}) \tilde{\nu} = N_{\nu x} \tilde{x} - N_{\nu \pi} \tilde{\psi} \to \tilde{\nu} = R N_{\nu x} \tilde{x} + (\Id - R) \tilde{\psi}$
+Exploiting the condition $N_{\nu \omega} = \Id - R^{-1} \to -R N_{\nu \omega} = \Id - R$
+
+I can rewrite the sector conditions as:
+
+\begin{align*}
+  \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr M_{1} R_{\phi} \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } \leq 0
+\end{align*}
+
+Given their sign definiteness I can subtract the term to $\Delta V$ in order to impose a negative bound to the Lyapunov function derivative:
+
+$$
+  \Delta V \leq \Delta V - He \left\{\bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } M_{1} R_{\phi} \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\right\} < 0
+$$
+
+Finally, I include the sector condition for the non-linearity $\phi = sin(\theta) - \theta$. The sector used is the following that is valid for $\theta \in \left[ -\pi, \pi \right]$:
+
+\begin{align*}
+  & \bmx{
+    \tilde{\theta} \\
+    sin(\tilde{\theta}) - \tilde{\theta}
+  }\tr \bmx{
+    0 & -1 \\
+    -1 & -2
+  } \bmx{
+    \tilde{\theta} \\
+    sin(\tilde{\theta}) - \tilde{\theta}
+  } \geq 0 \to \\
+  & \to \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr \bmx{
+    1 & 0 & 0 & 0_{1 \times n_{\phi}} & 0 \\
+    & 0_{1 \times n_{x}} & & 0_{1 \times n_{\phi}} & 1
+  }\tr \bmx{
+    0 & -1 \\
+    -1 & -2
+  } \bmx{
+    1 & 0 & 0 & 0_{1 \times n_{\phi}} & 0 \\
+    & 0_{1 \times n_{x}} & & 0_{1 \times n_{\phi}} & 1
+  } \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } > 0 \to \\
+  & \to \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr R_{q}\tr \bmx{
+    0 & -1 \\
+    -1 & -2
+  } R_{q} \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } > 0
+\end{align*}
+
+I add this sector condition to $\Delta V$:
+
+\begin{align*}
+  \Delta V &\leq \Delta V - He \left\{\bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } M_{1} R_{\phi} \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\right\} + \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } R_{q}\tr \bmx{
+    0 & -1 \\
+    -1 & -2
+  } R_{q} \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } < 0 \to \\
+  & \to \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr
+  \left( \bmx{
+    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C \\
+    \bar{B}\tr P \bar{A} & \bar{B}\tr P \bar{B} & \bar{B}\tr P C \\
+    C\tr P \bar{A} & C\tr P \bar{B} & C\tr P C
+  } - R_{\phi} \tr M_{1}\tr - M_{1} R_{\phi} + R_{q}\tr \bmx{
+    0 & -1 \\
+    -1 & -2
+  } R_{q}\right)
+  \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } < 0 \to \\
+  & \to \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr
+  M
+  \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  } < 0 \to M < 0 
+\end{align*}
+
+As regards the ellipsoid condition I use the same as in \textit{"Linear integrator 1 dimension case"}.
+
+\subsection*{3 layer saturation controller to non-linear pendulum with integrator in the state}
+
+The same formulation was extended to the 3 layer network where the only things that change are the matrix $N$ and the ellipsoid conditions that now are larger: $n_{\phi}$ instead of just $1$.
+
+After solving the LMI problem I compared the sizes of the regions of attraction (ROA) of the linear system controlled with the saturated input, the non-system controlled with the saturated input and the non-system controlled with the 3-layer NN controller.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.65\textwidth]{ROA_comparison}
+    \caption{It is intuitive to think that the ROA decreases linearly with the number of constraints set to the LMI}
+\end{figure}
 

### 2024-10-25 Journal/notes.tex
Refactor math_commands.tex to include dsfont package for set-style numbers

@@
 
-$$
+\begin{equation} \label{ETM-mechanism}
 s^{+} = \text{min}_{m \in \NN} \left\{ m \geq s + 1 | \psi(\omega_m, \chi) \geq \rho \eta_m \right\}
-$$
+\end{equation}
 
@@
 
-$$
+\begin{equation} \label{eta-dyn}
   \eta^{+} = (\rho + \lambda) \eta - \psi(\omega_m, \chi)
-$$
+\end{equation}
 
@@
 
+\subsection*{Extension to dynamic ETM}
+To further extend the problem the LMI formulation has been changed.
+A dynamic threshold on the ETM mechanism as been added and some matrices have been changed to account for the possible reference to include in the integrator dynamics.
+The same mechanism introduced in (\ref{ETM-mechanism}) and (\ref{eta-dyn}) has been used.
+
+Now each layer will have its own dynamical threshold:
+
+\begin{align*}
+  \eta_i^{+} &= (\rho_{i} + \lambda_{i}) \eta_i - \psi \\
+  \eta^{+} &= \bmx{
+    \eta_1^{+} \\
+    \vdots \\
+    \eta_l^{+}
+  } = \left( \bmx{
+    \rho_1 & 0 & \dots & 0\\
+    0 & \rho_2 & \dots & 0\\
+    \vdots & \vdots & \ddots & \vdots\\
+    0 & 0 & \dots & \rho_l
+  } + \bmx{
+    \lambda_1 & 0 & \dots & 0\\
+    0 & \lambda_2 & \dots & 0\\
+    \vdots & \vdots & \ddots & \vdots\\
+    0 & 0 & \dots & \lambda_l
+  }\right) \bmx{
+    \eta_1 \\
+    \vdots \\
+    \eta_l
+  } - \bmx{
+    \psi_1 \\
+    \vdots \\
+    \psi_l
+  } = \\
+  &= \left( R + L \right) \eta - \psi
+\end{align*}
+
+I define the vector $\one \in \RR^{n_{x}} = \bmx{1 & \dots & 1}\tr$
+
+And the Lyapunov function is now defined as $V = \tilde{x}\tr P \tilde{x} + He \left\{ \one\tr \eta \right\}$
 \pagebreak

### 2024-10-29 Journal/notes.tex
Added commentary up to ETM solution that includes the integrator term. Also a draft on the golden-section search has been has been draft out

@@
 
-\begin{align*}
+\begin{align} \label{matricione}
   \Delta V &= (\tilde{x}^{+})\tr P \tilde{x}^{+} - \tilde{x}\tr P \tilde{x} = \left( \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi} \right)\tr P \left( \bar{A} \tilde{x} + \bar{B} \tilde{\psi} + C \tilde{\phi} \right) - \tilde{x}\tr P \tilde{x} = \\
@@
   }
-\end{align*}
+\end{align}
 
@@
 
-\begin{align*}
+\begin{align} \label{V-condition}
   \Delta V &\leq \Delta V - He \left\{\bmx{
@@
   } < 0 \to M < 0 
-\end{align*}
+\end{align}
 
@@
 
-And the Lyapunov function is now defined as $V = \tilde{x}\tr P \tilde{x} + He \left\{ \one\tr \eta \right\}$
+And the Lyapunov function is now defined as $V = \tilde{x}\tr P \tilde{x} + He \left\{ \one\tr \eta \right\}$.
+The ETM mechanism for the $i$-th layer now includes the sector conditions for the activation functions and a contribution relative to the integrator state:
+
+$$
+  \tilde{\psi}\tr T \left[ G \tilde{x} - \tilde{\nu} + \tilde{\psi} \right] - k\tr \Omega k \geq R \eta
+$$
+
+With $k$ being the integrator $C x - r$ and $\Omega$ being a positive definite diagonal matrix.
+The dynamic of the integrator can now be possibly expressed as $z^{+} = z + k = z + C x - r$.
+There is a particular property in this configuration being the first element of $C \tilde{x} = C x - r$.
+This is useful because it doesn't require us to add another term $r$ in this case to the extended vector and still regroup everything with respect to $\left[ \tilde{x}, \tilde{\psi}, \tilde{\phi} \right]\tr$.
+The dynamic of the system can be expressed by:
+
+$$
+  x^{+} = A x + B u + C \phi + D r \to \tilde{x}^{+} = A \tilde{x} + B \tilde{\psi} + C \tilde{\phi} + D r - D r = A \tilde{x} + B \tilde{\psi} + C \tilde{\phi}
+$$
+
+I rewrite the sector conditions in a more compact way expressing:
+
+\begin{align*}
+  \Phi &= \bmx{
+    \tilde{\psi}_{1}\tr T_{1} [G_{1} \tilde{x} - \tilde{\nu}_{1} + \tilde{\psi}_{1}] - k\tr \Omega_{1} k  & \cdots & 0 \\
+    \vdots & \ddots & \vdots \\
+    0 & \cdots & \tilde{\psi}_{l}\tr T_{l} [G_{l} \tilde{x} - \tilde{\nu}_{l} + \tilde{\psi}_{l}] - k\tr \Omega_{l} k
+  } = \\
+  &= \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr \bmx{
+    -C_{k}\tr \Omega C_{k} & 0 & 0 & 0 & 0 \\
+    TG & -T & T & 0 & 0 \\
+    0 & 0 & 0 & 0 & 0 
+  } \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\nu} \\
+    \tilde{\phi} \\
+  } = \\
+  &= \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }\tr \bmx{
+    -C_{k}\tr \Omega C_{k} & 0 & 0 & 0 & 0 \\
+    TG & -T & T & 0 & 0 \\
+    0 & 0 & 0 & 0 & 0 
+  } \bmx{
+    \Id & 0 & 0 \\
+    R N_{\nu x} & \Id - R & 0 \\
+    0 & \Id & 0 \\
+    0 & 0 & \Id 
+  } \bmx{
+    \tilde{x} \\
+    \tilde{\psi} \\
+    \tilde{\phi}
+  }
+\end{align*}
+
+With $C_{k} \in \RR^{n_{\phi} \times n_{x}} = \bmx{
+  1 & 0 & 0 \\
+  \vdots & \vdots & \vdots \\
+  1 & 0 & 0 
+}$ selected to extract just the $\theta$ parameter from the state $x$.
+$\eta^{+}$ can be rewritten as: $\eta^{+} = (R + L) \eta - \Phi \one$
+With this condensed notation I proceed to discuss the sign of the increment of the Lyapunov function $\Delta V$ with respect to equation. \ref{matricione}:
+
+\begin{align*}
+  \Delta V &= (\tilde{x}^{+})\tr P (\tilde{x}^{+}) + He\left\{ \one\tr \eta^{+} \right\} - \tilde{x}\tr P \tilde{x} - He\left\{ \one\tr \eta \right\} =\\
+  &= \left( A \tilde{x} + B \tilde{u} + C \tilde{\phi} \right)\tr P \left( A \tilde{x} + B \tilde{u} + C \tilde{\phi} \right) - \tilde{x}\tr P \tilde{x} + He\left\{ \one\tr \left(   \left( R + L \right)\eta - \Phi \one - \eta  \right) \right\} = \\
+  &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr \bmx{
+    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C \\
+    \bar{B}\tr P \bar{A} & \bar{B}\tr P \bar{B} & \bar{B}\tr P C \\
+    C\tr P \bar{A} & C\tr P \bar{B} & C\tr P C
+  } \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } + He\left\{ \one\tr \left( \left( R + L - I \right) \eta - \Phi \one \right) \right\} =
+\end{align*}
+
+I can synthesize the ETM condition rewriting something like: $\Phi \one \geq R \eta$ that holds element-wise.
+I can then exploit $\eta \leq R^{-1} \Phi \one$ and plug it in $\Delta V$ to get rid of $\eta$:
+
+\begin{align*}
+  \Delta V &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr M \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } + He\left\{ \one\tr \left( \left( R + L - I \right) R^{-1} \Phi \one - \Phi \one \right) \right\} = \\
+  &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr M \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } + He\left\{ \one\tr \left( L - I \right) R^{-1} \Phi \one \right\} < 0
+\end{align*}
+
+Calling now $ \Gamma = (L - I) R^{-1} $ and regrouping everything with respect to the augmented vector I can rewrite:
+
+\begin{align*}
+  \Delta V &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr M \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } - \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr \bmx{
+    -C_{k}\tr \Gamma \Omega C_{k} & 0 & 0 & 0 & 0 \\
+    \Gamma T G & - \Gamma T & \Gamma T & 0 & 0 \\
+    0 & 0 & 0 & 0 & 0
+  } \bmx{
+    Id & 0 & 0 \\
+    R N_{\nu x} & Id - R & 0 \\
+    0 & Id & 0\\
+    0 & 0 & Id
+  } \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } = \\
+  &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr \left( M - M_{\pi} R_{\pi} - R_{\pi}\tr M_{\pi}\tr \right) \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }
+\end{align*}
+
+Finally, adding the sector conditions for the non-linearity $\phi$ like in equation. \ref{V-condition} I can impose the negative definiteness of the Lyapunov function derivative:
+
+\begin{equation}
+  \Delta V \leq \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr \left( 
+    M - M_{\pi} R_{\pi} - R_{\pi}\tr M_{\pi}\tr - R_{q}\tr \bmx{
+      0 & -1 \\
+      -1 & -2
+    } R_{q}
+   \right) \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+   }
+\end{equation}
+
+Given the local validity of the sector conditions for the activation functions it's necessary to add some ellipsoid conditions to the Lyapunov function.
+I want $\tilde{x}\tr \alpha P \tilde{x} \leq 1$, from the sector condition definition, see set $\cS$ in \cite{css-extended} I have:
+
+$$
+  \cS = \left\{ x(k) \in \RR^{n_{p}} : -\bar{\nu} - \nu_* \leq G (x(k) - x_*) \leq \bar{\nu} - \nu_* \right\}
+$$
+
+Since $\nu_* = 0$ the condition directly translates into $|G \tilde{x}| \leq \bar{\nu} \to (G \tilde{x})\tr (G \tilde{x}) \leq \bar{\nu}^2 \to \tilde{x}\tr \frac{G\tr G}{\bar{\nu}^{2}} \tilde{x} \leq 1$ I want this condition to be included in the ellipsoid defined by $\cE(P, x_*) = \left\{ x \in \RR^{n_{F}}: \tilde{x}\tr \gamma P \tilde{x} \leq 1 \right\}$
+
+\begin{multline*}
+  \tilde{x}\tr \frac{G\tr G}{\bar{\nu}^{2}} \tilde{x} \leq \tilde{x}\tr \gamma P \tilde{x} \leq 1 \to \frac{G\tr G}{\bar{\nu}^{2}} \leq \gamma P \leq 1 \to \gamma P - \frac{G\tr G}{\bar{\nu}^{2}} \geq 0\\
+  0
+\end{multline*}
+
+It is now possible to apply Schur's complement, dividing everything by $\gamma$:
+
+\begin{equation}
+  P - \frac{G\tr G}{\gamma \bar{\nu}^{2}} \geq 0 \to \bmx{P & G\tr\\ G & \gamma \bar{\nu}^{2}} \geq 0
+\end{equation}
+
+Now by pre and post multiplying by $\bmx{\Id & 0\\ 0 & T}$:
+
+\begin{equation}\label{ellipsoid}
+  \bmx{P & G\tr T\\
+  TG & \gamma \bar{\nu}^{2} T^{2}} \geq 0
+\end{equation}
+
+Equation \ref{ellipsoid} is the last LMI condition to add.
+The problem arises due to the bilinear term $TG$.
+A new change of variable is introduced: $Z \triangleq TG$.
+Doing so we deal with the bilinear term that appears here and in $M_{\pi}$.
+We still have to deal with the bilinear term $T^{2}$.
+Following the same approach as in \cite{css-extended} I exploit the inequality:
+
+$$
+  \left[ \alpha \bar{\nu}^{-2} - T \right] \bar{\nu}^{2}\left[ \alpha \bar{\nu}^{-2} - T \right] \geq 0 \to \alpha^{2} \bar{\nu}^{-2} - 2 \alpha T + \bar{\nu}^{2} T^{2} \geq 0 \to \bar{\nu}^{2} T^{2} \geq 2 \alpha T - \alpha^{2} \bar{\nu}^{-2}
+$$
+
+I can rewrite the ellipsoid condition as putting $\gamma = 1$
+
+$$
+  \bmx{P & Z\tr\\
+  Z & \bar{\nu}^{2} T^{2}} \geq \bmx{
+    P & Z\tr\\
+    Z & 2 \alpha T - \alpha^{2} \bar{\nu}^{-2}
+  } \geq 0
+$$
+
+More precisely a condition like the previous has to be imposed for each activation function in the network.
+Then we have to discuss the value of $\alpha$: the unitary value works but results to give conservative results.
+To find the optimal value a golden-section search algorithm can be used since the problem to address can be expressed as an unimodal function defined on a given interval, it has the advantage to reduce the search interval by a constant factor at each step.
+
+\begin{enumerate}
+    \item \textbf{Define the Interval and the Golden Ratio:} \\
+    Let $[a = 1, b = 0]$ be the initial interval containing the values for which the LMI is feasible. Defining the golden ratio $\varphi$ as:
+    $$
+    \varphi = \frac{1 + \sqrt{5}}{2}
+    $$
+
+    \item \textbf{Set Initial Interior Points:} \\
+    Calculate two interior points $ \alpha_1 $ and $ \alpha_2 $ within $[a, b]$ such that:
+    $$
+    x_1 = b - \frac{b - a}{\varphi}, \quad x_2 = a + \frac{b - a}{\varphi}.
+    $$
+
+    \item \textbf{Solve the LMI:} \\
+    Solve the LMI and store the value of the maximum eigenvalue of the $P$ matrix: $ \lambda_1, \lambda_2 $.
+
+    \item \textbf{Narrow the Interval:} \\
+    Based on the values $ \lambda_1, \lambda_2$, update the interval $[a, b]$ as follows:
+    \begin{itemize}
+        \item If $ \lambda_1 < \lambda_2 $, then the minimum is in the interval $[a, x_2]$. Set $ b = x_2 $.
+        \item If $ \lambda_2 < \lambda_1 $, then the minimum is in the interval $[x_1, b]$. Set $ a = x_1 $.
+    \end{itemize}
+
+    \item \textbf{Repeat the Process:} \\
+    Recalculate $ x_1 $ and $ x_2 $ within the new interval $[a, b]$ and repeat steps 3 and 4.
+At each iteration, the interval is reduced by a factor of the golden ratio, retaining only the portion containing the minimum.
+
+    \item \textbf{Convergence:} \\
+    Continue until the length of the interval $[a, b]$ is less than a chosen tolerance level, $\epsilon$.
+At this point, any $\alpha$ value within the interval can be considered to give the biggest Region of attraction estimate.
+\end{enumerate}
+
+After solving the LMI, even though the solution is found the ROA is still incredibly small.
+In addition to that the LMI finds a solution such that $\Omega$ has every item close to 0, this means that the "help" we want to give the ETM to trigger less is not helpful to the LMI problem and is therefore ignored.
+
+Since no other tweaks on the LMI formulation seems to be possible the next approach will be using the currently deep learning trained network and re-train it with RL by projecting a solution of the LMI in each training step to enlarge as much as possible the starting ROA.
+
+\section*{Retraining of NN with RL to enlarge the ROA}
+The main problem now is to design an efficient cost function to enlarge the ROA. After a first brainstorm the ideas are:
+\begin{itemize}
+  \item Give as a cost $ln(1/\bar{\lambda})$ with $\bar{\lambda}$ the maximum eigenvalue of last $P$. 
+  \item Use $\Delta V$ when it's negative as a cost. Like a $\texttt{relu}(\Delta V)$.
+\end{itemize}
+
 \pagebreak

### 2024-10-31 Dubbi/notes.tex
dubbi

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage[margin=54pt]{geometry}

\begin{document}

\date{}
\author{}

\title{Set-up del RL per migliorare la rete}
\maketitle

\section*{Environment}
Usando sempre \texttt{stable\_baseline} ho pensato di semplificare il problema e invece di usare il pendolo come env uso direttamente i pesi e i bias. La action quindi sarebbe un vettore lungo quanto tutti i parametri (pesi e bias) che contiene degli incrementi da sommare ai parametri.

\subsection*{Action e observation space}
Sono dei vettori belli lunghi. Per la nuova rete addestrata sul sistema senza integratore ($x = \bmx{\theta, \dot{\theta}}$) ho 3 hidden layer da 32 neuroni:
\begin{itemize}
  \item N pesi: $2 \cdot 32 + 32 \cdot 32 + 32 \cdot 32 + 32 \cdot 1 = 2144$
  \item N bias: $32 + 32 + 32 + 1 = 97$
\end{itemize}
In totale 2241 parametri. La action quindi sarà un vettore di 2241 elementi.

\subsection*{Reward}
A ogni step prendo il nuovo set di pesi e bias e risolvo la LMI corrispondente.
\begin{itemize}
  \item Feasible ma ROA peggiore: reward positivo e proporzionale in qualche modo alla differenza tra la migliore ROA attuale e quella ottenuta con i nuovi parametri.
  \item Feasible e ROA migliore: reward molto positivo e conservo i pesi e bias attuali
  \item Non feasible: reward negativo
\end{itemize}
\textit{Problema:} I bias influenzano solo l'equilibrio finale ma non la dinamica del pendolo e la ROA. Quale potrebbe essere un buon criterio per capire come cambiarli?

\subsection*{Reset}
Quando viene chiamato il reset rimetto i pesi e bias che danno la ROA migliore trovata finora.\\

\subsection*{Durata episodi}
Dato che ora l'env non è un vero e proprio sistema come regolo la durata di un episodio? Pensavo a fare continuare un episodio finché trova risultati feasible anche se non migliori.

\section*{Training}
Data la dimensione della action, hai qualche consiglio sulla scelta della architettura della rete? Ho provato con reti relativamente semplici ($32, 128, 128, 32$) e fanno abbastanza schifo, trova al massimo 3-4 soluzioni migliori (di pochissimo) all'inizio e poi parte per la tangente

\subsection*{Tempi di addestramento}
Dato che deve risolvere una LMI a ogni step ovviamente si parla di tempi biblici ma almeno il problema è incrementale e se si trova anche solo una soluzione migliore in una run si può ripartire da quella.

\end{document}

### 2024-11-04 Journal/notes.tex
Add detailed explanation of non-linear pendulum approach and training strategies

@@
 
+The problem under examination is now the simpler non-linear pendulum without integrator in the state.
+A new starting network has been trained with deep neural network based on a linear LQR controller.
+Now the two approaches listed before will be followed:
+
+\begin{itemize}
+  \item The first approach sees the treatment of the environment as the NN itself.
+The action will be a list of increments to add to each element of the weights and the biases.
+The starting point will be the previously trained network.
+A step is the creation of the new weights and biases and solve the LMI for them.
+The cost will be continuous and proportional to the area of the ROA found.
+If infeasible the cost will be $0$.
+The length of one episode will be decided a priori.
+If we have a better solution that gives a bigger ROA then the weights will be stored but not updated.
+They will be updated only on reset and if we have an increment in the ROA's area that is big enough.
+Another aspect to take into account is the size of the training NN.
+The action will be a vector of $\simeq 2100$ elements.
+  \item This SEEMS easier to implement.
+A rollout of the system will be done and the $\Delta V$ of the rollout will be computed with respect to the last $P$ found, it will then be used as a cost.
+The architecture of the NN should be the same as the one trained with deep learning and should accept as a starting point the previously trained weights and biases.
+At each reset a LMI will be solved, if it leads to a bigger ROA then the $P$ matrix to which we refer to improve will be updated.
+\end{itemize}
+
 \pagebreak

### 2024-11-15 Journal/notes.tex
ROA enlargement final notes New direction for research

@@
 The action will be a vector of $\simeq 2100$ elements.
+
+  \textbf{This method was discarded because of the big dimension of the action. Several attempts were executed but with no good results.}
+
   \item This SEEMS easier to implement.
@@
 
+Last method resulted to be the efficient one.
+It was hard to set up correctly the environment to make it compliant with \texttt{stable\_baselines3}, in particular the handling of the LMI mid-training and the following update of the current P matrix stored by the environment.
+Rather than a linear method a sequence of different tests has been done to find the best hyperparameters for the training.
+Initially the results from the \textit{LQR} solution were used as a "warm start" to try and find solutions that led to bigger Region of Attractions.
+To accomplish so two different reward functions design were alternatively used:
+
+\begin{itemize}
+  \item Classical RL reward: $J = -w_{\theta} \theta^2 - w_{v} v^2 - w_{\eta} \eta^2 - w_{u} u^2 + \delta$\\
+  Where $\delta$ is a small constant that acts as a constant reward for each step the policy manages to keep the environment "alive".
+Even if not directly demonstrated using a reward function shaped like this should give results similar in quality to the LQR since both approaches try to minimize the quadratic weight of the state and the control input.
+  \item Custom defined reward: $J = - \Delta V + \frac{\Delta(ROA)}{n_{episode}}$
+
+  Here directly passing $\Delta V$ serves to tell the algorithm to penalize every action that leads to a Lyapunov function that is not strictly decreasing.
+The second term is a bonus that is given to the agent when it manages to enlarge the ROA, since the update of the policy is done at the end of each rollout the bonus is divided by the number of episodes to make the reward small enough to not overshadow the $\Delta V$ term.
+If the current policy update gives a bigger ROA then the weights are stored and the $\Delta(ROA)$ is instead the overall area of the ROA found, hence giving a bit more reward than cost.
+Also, the $P$ matrix that is used to compute the reward is updated with the last one found.
+
+  An important aspect that had to be taken into account is the nature of the action give by the RL algorithm being normalized between $-1$ and $1$.
+To address this peculiarity the LQR solution was found again to be compliant.
+Given the normalized nature of the action that then get multiplied by the maximum torque appliable by the system I thought to consider it as a saturation applied to the output of the NN and the input of the system.
+I modified accordingly the LMI by adding an element to the matrices that manage the activation functions sector conditions and another constraint for the ellipsoid condition that now takes the maximum torque as a bound and not the saturation functions value used in the activation function.
+
+  By empirical observations it has been noticed that after a certain amount of rollouts the algorithm stopped to find feasible policy updates.
+To overcome this problem a new logic has been added: after a certain amount of rollouts without improvements or in the occurrence of an infeasible update the policy would be restored to the best one found so far, the optimizer is also reset and hence the training starts again from the best policy found so far.
+The good thing about this approach is the fact that the rollout buffer is not emptied hence, even if very slowly, the algorithm can still find better policies in the long term.
+After many iterations and several runs I ended up with a ROA that is more than 50 times larger in volume than the one found after the deep learning training.
+\end{itemize}
+
+\section*{New direction of the research}
+Once the ROA has been enlarged the next step is to try and find a direction to present the results found so far.
+A first idea is to make the paper centered around the ETM implementation.
+Using the new policy already it shows that it is feasible with dynamic ETM.
+The behavior is good characterized by a very high update rate of the first layer ($\simeq 100\%$) but the update rate of the last layer, the one responsible for the actual input update to the system, is more than halved stabilizing around $40\%$.
+Obviously the addition of the ETM mechanism makes the system more conservative since it reduces the Region of Attraction.
+For this reason we want now to find a way to make the reduction of the ROA preserve some particular directions that are now guaranteed to have more stable states due to the addition of the integrator in the dynamics.
+
+To address this problem Samuele advised to look at some work regarding multi-objective LMIs and to look at previous works of Sophie.
+Most probably It will be translated into the application of another constraint to the $\Delta V$ sign in the LMI and some change in the objective of the problem definition.
+
 \pagebreak

### 2024-11-15 Journal/notes.tex
Update in direction of research

@@
 
+Ideas to address:
+\begin{itemize}
+  \item Since the current static ETM matches the minimum feasible ROA we could think of some other ETM mechanism.
+To be clearer the quadratic constraints of the non-linearities of the activation functions are mandatory and cannot be neglected.
+Since the static ETM uses the exact same conditions It is clear that the ROA will be the same.
+  \item One possible criteria to try and obtain with an ETM is something regarding the inter-event period and the preferred direction to preserve in ROA shrinking.
+In particular the direction around $\theta$ since the addition of the integrator in the dynamics makes the system more stable in that direction.
+Some work of Sophie should have addressed something similar.
+As regards the inter-event period look at some work of Antoine Girard.
+  \item Another possible improvement is changing the quadratic constraints whenever we have some fixed matrices like the one used to address the $sin(\theta) - \theta$ non-linearity.
+It sees the addition of new variables and constraints to the LMI like seen in the two \textit{k-contraction} papers of Samuele.
+  \item The last idea is to try and find a way to implement multiple objectives in the LMI. There are some papers to take idea from like in \cite{multiobjective-1} and \cite{multiobjective-2}.
+\end{itemize}
+
 \pagebreak

### 2024-11-18 Journal/notes.tex
Update on Monday 18 nov

@@
 
+After fixing some error on $x_{*}$ computation solving the following implicit equation:
+
+\begin{align*}
+  x^{+} &= A x + B u + C \phi + D r\\
+  x_{*} &= A x_{*} + B u_{*} + C \phi(x_{*}) + D r \\
+  u_{*} &= N_{ux} x_{*} + N_{u \omega} \omega_{*} + N_{ub} = \\
+  &= N_{ux} x_{*} + N_{u \omega} \nu_{*} + N_{ub} = \\
+  &= N_{ux} x_{*} + N_{u \omega} \left( \Id - N_{\nu \omega} \right) ^{-1} \left(  N_{\nu x} x_{*} + N_{\nu b}\right) + N_{\nu b} = \\
+  &= \left( N_{ux} + N_{u \omega} R N_{\nu x} \right) x_{*} + N_{u \omega} R N_{\nu b} + N_{ub} = R_{\omega} x_{*} + R_{b}\\
+  x_{*} &= A x_{*} + B R_{\omega} x_{*} + B R_{b} + C \left( sin(K x_{*}) - K x \right) + D r = \\
+  x_{*} &= \left( A + B R_{\omega} - C K \right) x_{*} + B R_{b} + C sin(K x_{*}) + D r
+\end{align*}
+
+Hence, the implicit equation to solve is:
+
+\begin{equation}
+  \textit{func}(x) = (-I + A + B R_{\omega} - C K) x + B R_{b} + C sin(K x) + D r 
+\end{equation}
+
+With $K = \bmx{1 & 0 & 0}$
+
+The issue I wanted to solve was the continuous update and event triggering even after the system has reached the equilibrium.
+The equilibrium reach was confirmed by inspecting the behavior of the system, it has been shown a Lyapunov function that tends to 0 (order of $1e-30$) and a norm of $\tilde{x} = x - x_{*}$ that tends to 0 (order of $1e-6$) after just 300 steps.
+The control input also tends to $u_{*}$ faster than the state but slower than the Lyapunov.
+The intuition was to add some threshold on the magnitude of the actors of the ETM, since the ETM mechanism defined by $
+  \Psi \geq \rho \eta $ triggers around equilibrium for very small values of $\Psi$ (around $1e-30$) by adding a random threshold small enough, even if it was not yet implemented in the LMI, an improvement on the update rate has been noticed.
+However, it didn't eliminate the behavior we wanted to avoid being the non-necessary updates once the equilibrium has been reached.
+An increment on the update rate has been observed until the threshold was not big enough to mess up all the system and ETM mechanism.
+
+  To continue addressing the problem I will look into the \textit{"Stubborn State Observers"} topic first introduced by Luca. See \cite{stubborn-1}, \cite{stubborn-2} and \cite{stubborn-3}.
+
 \pagebreak

### 2024-11-22 Journal/notes.tex
Progress on finsler approach

@@
 
+  \subsection*{Finsler's Lemma workaround}
+
+  A way to try and solve the problem was to use a different matrix for the ETM. For the $i^{th}$ layer it was used the following matrix so far:
+
+  $$    
+    \bmx{
+      \tilde{x}\\
+      \tilde{\psi_{i}}\\
+      \tilde{\nu_{i}}
+    }\tr
+    \Omega_{i}
+    \bmx{
+      \tilde{x}\\
+      \tilde{\psi_{i}}\\
+      \tilde{\nu_{i}}
+    } = \bmx{
+      \tilde{x}\\
+      \tilde{\psi_{i}}\\
+      \tilde{\nu_{i}}
+    }\tr
+    \bmx{
+      0 & 0 & 0\\
+      T_{i} G_{i} & T_{i} & -T_{i}\\
+      0 & 0 & 0 
+    }
+    \bmx{
+      \tilde{x}\\
+      \tilde{\psi_{i}}\\
+      \tilde{\nu_{i}}
+    }
+  $$
+
+  With $TG = Z$ to address the bilinear term.
+Due to the sparseness of $\Omega$ we want to find a denser matrix $X$ that approximates $\Omega$ and is compliant with the Finsler's Lemma.
+It can be done by adding the following constraint in the LMI:
+
+  $$
+  \Omega = \bmx{
+    \begin{array}{ccc|ccc|ccc}
+      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      Z_{1} & T_{1} & -T_{1} & 0 & 0 & 0 & 0 & 0 & 0\\
+      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      \hline
+      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      0 & 0 & 0 & Z_{2} & T_{2} & -T_{2} & 0 & 0 & 0\\
+      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      \hline
+      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      0 & 0 & 0 & 0 & 0 & 0 & Z_{3} & T_{3} & -T_{3}\\
+      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\
+    \end{array}
+  }
+$$
+
+$$
+  X = \bmx{
+    \begin{array}{c|c|c}
+      X_{1} & 0 & 0\\
+      \hline
+      0 & X_{2} & 0\\
+      \hline
+      0 & 0 & X_{3}
+    \end{array}
+  }
+$$
+
+$$
+R_{\xi}\tr He\left\{ X - \Omega \right\} R_{\xi} + He \left\{  \pmx{N_{1} \\ N_{2} \\ N_{3}} \pmx{R N_{\nu x} & \Id - R & -\Id} \right\} \leq 0
+$$
+
+The matrix $R_{\xi}$ orders the vector that should be multiplied by $X$ and $\Omega$:
+
+$$
+     \bmx{
+      \tilde{x}\\
+      \tilde{\psi}\\
+      \tilde{\nu}
+     } = \bmx{
+      \tilde{x}\\
+      \tilde{\psi_{1}}\\
+      \tilde{\psi_{2}}\\
+      \tilde{\psi_{3}}\\
+      \tilde{\nu_{1}}\\
+      \tilde{\nu_{2}}\\
+      \tilde{\nu_{3}}
+    } = \bmx{
+      \tilde{x}\\
+      \tilde{\psi_{1}}\\
+      \tilde{\nu_{1}}\\
+      \tilde{x}\\
+      \tilde{\psi_{2}}\\
+      \tilde{\nu_{2}}\\
+      \tilde{x}\\
+      \tilde{\psi_{3}}\\
+      \tilde{\nu_{3}}
+    } R_{\xi} = \bmx{
+      \tilde{x}\\
+      \tilde{\psi_{1}}\\
+      \tilde{\nu_{1}}\\
+      \tilde{x}\\
+      \tilde{\psi_{2}}\\
+      \tilde{\nu_{2}}\\
+      \tilde{x}\\
+      \tilde{\psi_{3}}\\
+      \tilde{\nu_{3}}
+    } \bmx{
+      \Id & 0 & 0 & 0 & 0 & 0 & 0\\
+      0 & \Id & 0 & 0 & 0 & 0 & 0\\
+      0 & 0 & 0 & 0 & \Id & 0 & 0\\
+      0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      0 & 0 & \Id & 0 & 0 & 0 & 0\\
+      0 & 0 & 0 & 0 & 0 & \Id & 0\\
+      0 & 0 & 0 & 0 & 0 & 0 & 0\\
+      0 & 0 & 0 & \Id & 0 & 0 & 0\\
+      0 & 0 & 0 & 0 & 0 & 0 & \Id
+    } 
+$$
+
+With this condensed notation I proceed to discuss the sign of the increment of the Lyapunov function $\Delta V$.
+I start by plugging the ETM condition, further considerations on the proper sector conditions will be done later.
+Now the matrix $X$ is used instead of $\Omega$:
+
+\begin{align*}
+  \Delta V &= (\tilde{x}^{+})\tr P (\tilde{x}^{+}) + He\left\{ \one\tr \eta^{+} \right\} - \tilde{x}\tr P \tilde{x} - He\left\{ \one\tr \eta \right\} =\\
+  &= \left( A \tilde{x} + B \tilde{u} + C \tilde{\phi} \right)\tr P \left( A \tilde{x} + B \tilde{u} + C \tilde{\phi} \right) - \tilde{x}\tr P \tilde{x} + He\left\{ \one\tr \left(   \left( R + L \right)\eta - X \one - \eta  \right) \right\} = \\
+  &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr \bmx{
+    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C \\
+    \bar{B}\tr P \bar{A} & \bar{B}\tr P \bar{B} & \bar{B}\tr P C \\
+    C\tr P \bar{A} & C\tr P \bar{B} & C\tr P C
+  } \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } + He\left\{ \one\tr \left( \left( R + L - I \right) \eta - X \one \right) \right\} =
+\end{align*}
+
+I can synthesize the ETM condition rewriting something like: $X \one \geq R \eta$ that holds element-wise. I can then exploit $\eta \leq R^{-1} X \one$ and plug it in $\Delta V$ to get rid of $\eta$:
+
+\begin{align*}
+  \Delta V &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr M \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } + He\left\{ \one\tr \left( \left( R + L - I \right) R^{-1} X \one - X \one \right) \right\} = \\
+  &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr M \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } + He\left\{ \one\tr \left( L - I \right) R^{-1} X \one \right\} < 0
+\end{align*}
+
+Calling now $ \Gamma = (L - I) R^{-1} $ and regrouping everything with respect to the augmented vector I can rewrite:
+
+\begin{align*}
+  \Delta V &= \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr M \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  } + \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }\tr R_{\nu}\tr \bar{R_{\xi}}\tr 
+  He \left\{ \Gamma \bar{X}  \right\} \bar{R_{\xi}} R_{\nu} \bmx{
+    \tilde{x}\\
+    \tilde{\psi}\\
+    \tilde{\phi}
+  }
+\end{align*}
+
+The matrix $\bar{R_{\xi}}$ now takes into account the non-linearity $\tilde{\phi}$, the matrix $R_{\nu}$ takes care of $\tilde{\nu}$
+
+$$
+  \bar{R_{\xi}} = \bmx{
+  \begin{array}{c | c}
+    R_{\xi} & 0\\
+    \hline
+    0 & \Id
+  \end{array}
+  }; R_{\nu} = \bmx{
+    \Id & 0 & 0\\
+    0 & \Id & 0\\
+    R N_{\nu x} & \Id - R & 0\\
+    0 & 0 & \Id
+  }
+$$
+
+Now the previously seen sector conditions for the sin non linearities can be added in the same way as before.
+
+\textbf{SHOULD I ADD SECTOR CONDITIONS FOR ETM OR ARE THEY ALREADY TAKEN INTO ACCOUNT INSIDE THE FINSLER CONDITIONS?}
+
 \pagebreak

### 2024-12-05 paper structure/notes.tex
Paper structure

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage[margin=54pt]{geometry}
\usepackage{biblatex}
\addbibresource{../biblio.bib}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Ideas on paper structure}
\maketitle

\section{\textbf{Abstract}}
Things to cite:
\begin{itemize}
  \item Design of \textbf{dynamic} ETM for Discrete Time \textbf{non-linear} systems stabilized by neural network controller
  \item Events' triggering logic based on the use of local sector conditions for the non-linear activation functions, saturation
  \item Development of sufficient LMI conditions for ETM parameters' design together with the computation of an inner approximation of the region of attraction
  \item Convex optimization procedures to achieve different objectives in the final system behavior. (e.g. size of ROA, number of events, etc.)
  \item \textbf{Sara dice che l'Abstract si fa per ulitmo e ha ragione}
\end{itemize}

\section{Introduction}

\section{Problem statement}
\subsection{System description}
\subsection{Controller description}
\subsection{Event-triggering Mechanism description}

\section{}

% \pagebreak
% \printbibliography

\end{document}

### 2024-12-06 paper structure/theory.tex
Schemes created

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage[margin=54pt]{geometry}
\usepackage{biblatex}
\addbibresource{../biblio.bib}

\begin{document}

\date{}
\author{Marco Sterlini}

\title{Theory draft}
\maketitle

\subsection*{\textbf{Notation}}
$\NN, \RR^{n}, \RR^{n \times m}$ denote the set of natural non-negative integers, the set of real vectors of dimension $n$, and the set of real matrices of dimension $n \times m$, respectively. For any matrix $A$, $A\tr$ is its transpose. For any square matrix $A$ the operator $He\left\{ A \right\} = A + A\tr$ is defined. $\texttt{diag}(A_1, A_2)$ is a block-diagonal matrix with block diagonal matrices $A_1$ and $A_2$. For a partitioned matrix the symbol $\star$ stands for symmetric blocks.

\section*{Problem statement}

\section*{\textbf{Lemma}}
Considering Finsler's lemma in the context of our ETM, with respect to system (\dots) and layer $i \in \left[ 0, \dots, l \right]$, and assuming that (\dots) holds $\forall \xi \in \left[ \underbar{\xi},  \bar{\xi} \right]$, if $\exists X_i \in \RR^{\left( n_x + 2 \cdot n_{\phi_i} \right) \times\left( n_x + 2 \cdot n_{\phi_i} \right)}, N_{1}^{i} \in \RR^{n_{\phi} \times n_x}, N_2^i \in \RR^{n_{\phi} \times n_{\phi}}, N_3^i \in \RR^{n_{\phi} \times n_{\phi}}$

\end{document}

### 2024-12-06 Paper draft/draft.tex
Progress in paper

\documentclass{article}
\input{../math_commands}
\input{../custom_imports}
\usepackage{biblatex}
\addbibresource{../biblio.bib}

\begin{document}

\title{Draft of paper}

\author{Marco Sterlini}

\section{Introduction}

\section{Problem Formulation}
We consider the non-linear discrete-time system stabilized by a Neural Network controller represented in figure \ref{fig:first_scheme} described by:
\begin{equation}
  x^{+} = A x + B \text{sat}(\bar{u}) + C \Phi(E x) + D d 
\end{equation}
with $x \in \RR^{n_x}$ the state vector, $\bar{u}$ the raw output of the controller, $\Phi$ the representation of the non-linearity of the system and $d$ a constant disturbance input.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.65\textwidth]{img/first_scheme}
    \caption{Feedback system}
    \label{fig:first_scheme}
\end{figure}

The controller is implemented as a Multi Layer Perceptron with $l$ layers, each with $n_{\phi_i}$ neurons, all the activation functions are saturation functions, represented by $\phi_i$. The objective is to design an event-triggering mechanism that allows to ease the computational burden associated with the computation of many non-linear activation functions providing at the same time sufficient sector conditions to handle the non-linearity of the system and the controller. The controller $\pi$ is defined by:
\begin{equation}\label{eq:nn-equations}
  \begin{aligned}
  \omega^{0}(k) &= x(k) \\
  \nu^{i}(k) &= W^{i} \omega^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
  \omega^{i}(k) &= \phi_i(\nu^i(k))\\
  \bar{u}(k) &= \phi_l(\nu^l(k))
  \end{aligned} 
\end{equation}
with $\nu^i(k) \in \RR^{n_{\phi_i}}$ the input to the $i$-th activation function $\phi_i: \RR^{n_{\phi_i}} \to \RR^{n_{\phi_i}}$, $\omega^i(k) \in \RR^{n_{\phi_i}}$ the output. The weights $W^i \in \RR^{n_{\phi_i} \times n_{\phi_{i-1}}}$ and biases $b^i \in \RR^{n_{\phi_i}}$ define the affine operation of each layer. The saturation function application is to be intended element-wise and will be denoted as $\phi_i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$ and $\varphi: \RR \to \RR$ is the scalar activation function assumed to be symmetric and identical for every neuron. Adopting the same notation as in \cite{css-paper} we define the controller policy in a condensed form. Denoting the augmented vectors:
\begin{equation*}
  \nu_{\phi} = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega_{\phi} = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi_{\nu_{\phi}} = \bmx{\phi_1(\nu^1) \\ \vdots \\ \phi_l(\nu^l)} 
\end{equation*}
with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega_\phi = \phi(\nu_\phi)$. Finally, conditions \ref{eq:nn-equations} can be rewritten as:
\begin{equation*}
  \bmx{u(k)\\ \nu_\phi(k)} = N \bmx{x(k) \\ \omega_\phi(k) \\ 1} 
\end{equation*}
where
\begin{equation}\label{eq:first-N}
  \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & W^l & b^l \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^{l-1} & \0 & b^{l-1}
    \end{array}
  }
\end{equation}

Similarly as in \cite{css-extended} we implement an event-triggering-mechanism (ETM) in every layer of the controller, we add also a final ETM to the output that will be used to further reduce the total number of events and call to the non-linear functions. The system and the controller are now described by figure \ref{fig:second_scheme}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.45\textwidth]{img/second_scheme}
    \caption{Feedback system subject to ETM in the controller}
    \label{fig:second_scheme}
\end{figure}

The dynamics of the system are now described by:
\begin{equation}\label{eq:system-dynamics}
  x^{+} = A x + B u + C \Phi(E x) + D d
\end{equation}
with $u$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \ref{eq:first-N} as:
\begin{equation}\label{eq:last-N}
  N = \bmx{
    \begin{array}{c | c c c c | c}
      \0 & \0 & \dots & \0 & \Id & 0 \\ 
      \hline
      W^1 & \0 & \dots & \0 & \0 & b^1 \\
      \0 & W^2 & \dots & \0 & \0 & b^2 \\
      \vdots & \vdots & \ddots & \vdots & \vdots & \vdots \\
      \0 & \0 & \dots & W^l & \0 & b^l
    \end{array}
  } = \bmx{
    \begin{array}{c | c | c}
    N_{u x} & N_{u \omega} & N_{u b} \\
    \hline
    N_{\nu x} & N_{\nu \omega} & N_{\nu b}
    \end{array} 
  }
\end{equation}

\section{Main Results}

\subsection{Event-triggering mechanism}

\subsection{Lyapunov conditions}

\subsection{Optimization procedure}

\section{Simulations}

\section{Conclusions}

\printbibliography

\end{document}

### 2024-12-07 paper structure/notes.tex
Theory nodes

@@
 \maketitle
+
+\section{Results and news I want to show}
+\begin{itemize}
+  \item Theory structure with Finsler application portrayed as a Lemma in the most general way possible so to be able to cite it whenever I need.
+Condition on stability via LMI conditions portrayed as a Theorem like in Sophie's paper.
+All the results in the flexibility of $\lambda$ to obtain different results on the same LMI problem portrayed as a proposition.
+  \item New complexity of the system being now non-linear and with the integrator in the state
+  \item Very brief explanation of the training procedure followed (RL) to obtain the new controller with focus on the solution of the LMI embedded at the end of each rollout of the algorithm and used as cost or reward for the following episodes discarding non-feasible results ensuring a better or equal solution.
+  \item Change in the ETM: now dynamic and improved by finsler lemma
+  \item Plots:
+  \begin{itemize}
+    \item Plot of performances regarding size of ROA and density of updates with respect to $\lambda$ values, seen as interpolation of results or as double bar histograms
+    \item Plot of $\eta_i$ for $\lambda = 0$ to see the small decay rate
+    \item Plot of a very ugly trajectory inside the 3D ellipsoid with the projections on the 3 planes
+    \item Plot of a Lyapunov and $\Delta V$ maybe for $\lambda = 0$ and $\lambda = 1$
+    \item Maybe a stacked plot of the control inputs for a fixed time window that highlights the change in value after and event.
+  \end{itemize}
+  \item Tables with the update rates of each layer for different $\lambda$ values with respect to the case without finsler where the first layer sees an update rate close to $100\%$
+  \item Discussion about the best $\eta_0$ value to get with respect to $x_0$ to minimize the number of events.
+\end{itemize}
+
+\section{Elements I want to import from previous works}
+\begin{itemize}
+  \item Block representation of the system with the controller and the ETM, see Arcak and Carla
+  \item NN controller with weights and bias formulation with extraction of equilibrium.
+  \item Theoretical citation to Sophie's book as regards the sector conditions for saturation and their implementation as ETM conditions.
+  \item Paragraph focused on the optimization procedures implemented, maybe shorted but preceded by a section where it is explained in full detail stressing the flexibility of the problem.
+Concerning: \begin{itemize}
+    \item $R$ matrix minimization or maximization to $\Id$
+    \item Search algorithm for the best $\alpha$ used in the inclusion condition to deal with the bilinear term
+    \item Search algorithm 
+  \end{itemize} 
+  \item Simulations results previously discussed
 

### 2024-12-08 Paper draft/draft.tex
End of problem formulation

@@
 
+The system is compliant with the assumptions of \textbf{Lemma 2} in \cite{css-extended} that allow us to define the following expressions regarding the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller given $x_*$, defining the matrices:
+
+\begin{equation}
+    \begin{aligned}
+         R &= (\Id - N_{\nu \omega})^{-1}\\
+         R_{\omega} &= N_{u x} + N_{u \omega} R N_{\nu x}\\
+         R_b &= N_{u \omega} R N_{\nu b} + N_{u b}
+    \end{aligned}
+\end{equation}
+
+With $N_{\nu \omega}$ always invertible due to its lower triangular structure. We express:
+
+\begin{equation}
+  \begin{aligned}
+    u_* &= R_{\omega} x_* + R_b\\
+    \nu_* &= R N_{\nu x} x_* + R N_{\nu b}\\
+    \omega_* &= \nu_*
+  \end{aligned}
+\end{equation}
+
+The equilibrium state $x^*$ needs some further discussion since the expression to retrieve it is no longer explicit due to the non-linearity $\Phi$.
+We can retrieve the vector using a numerical solver that minimizes the following expression for a given disturbance $\bar{d}$:
+
+\begin{equation}
+  (A + B R_{\omega} - \Id) x_* + C \Phi(E x_*) + D \bar{d} + B R_b = 0
+\end{equation}
+
 \section{Main Results}

### 2024-12-09 Paper draft/draft.tex
Update up to ETM desing. TODO Finsler lemma

@@
 \begin{equation}
+  \bmx{u_* \\ \nu_*} = N \bmx{x_* \\ \omega_* \\ 1}; \ \  
   \begin{aligned}
-    u_* &= R_{\omega} x_* + R_b\\
+    \omega_* &= \nu_*\\
     \nu_* &= R N_{\nu x} x_* + R N_{\nu b}\\
-    \omega_* &= \nu_*
+    u_* &= R_{\omega} x_* + R_b
   \end{aligned}
@@
 \section{Main Results}
+Taking inspiration from \textbf{Proposition 1} and \textbf{Lemma 3} in \cite{css-extended} we design the ETM for the output of each layer of the controller and the final output.
+The idea is to exploit the sector conditions of the activation functions as ETM conditions.
 
 \subsection{Event-triggering mechanism}
+An ETM works by triggering an event that will update the current layer and propagate its value across the network.
+The triggering function will take as inputs the incremental variables of the state of the system
+$\tilde{x}$, the current state of the layer $\tilde{\nu}$ and the dead-zone vector $\tilde{\psi} = \tilde{\nu} - \tilde{\omega}$. The ETM was defined as:
+\begin{equation}
+  \omega^{i}(k) = \begin{cases}
+    \omega^{i}(k-1) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > 0\\
+    \phi_i(\nu^i(k)) & \text{otherwise}
+  \end{cases}
+\end{equation}
+
+In the precedent work \cite{css-extended} the function $f$ is defined on the basis of the saturation sector conditions in the following quadratic form:
+
+\begin{equation}\label{eqn:base_omega}
+  f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) = 
+  \tilde{\psi}\tr T^i \left[ G^i \tilde{x} + \tilde{\psi} - \tilde{\nu} \right] = 
+  \bmx{\tilde{x}\\ \tilde{\psi}^i \\ \tilde{\nu}^i}\tr
+  \bmx{
+    0 & 0 & 0\\
+    T^i G^i & T^i & -T^i\\
+    0 & 0 & 0
+  } 
+  \bmx{\tilde{x}\\ \tilde{\psi} \\ \tilde{\nu}} > 0
+\end{equation}
+
+This approach acts as a static ETM triggering an event as soon as the triggering function $f$ is positive.
+Inspired by the work of \cite{data-driven} we define a dynamic threshold to be implemented in the ETM.
+The idea is to lower the conservatism of the controller by introducing a time-varying threshold that will adapt to the current state of the system reducing the overall number of events and computational burden.
+The new triggering function is defined as:
+
+\begin{equation}
+  \omega^{i}(k) = \begin{cases}
+    \omega^{i}(k-1) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta_i\\
+    \phi_i(\nu^i(k)) & \text{otherwise}
+  \end{cases};
+\end{equation}
+\begin{equation}
+  \eta^+_i = \rho^i \eta_i - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu})\  \forall i \in \left( 1, \dots, l \right) 
+\end{equation}
+with $0 < \rho^i < 1 \ \forall i \in \left( 1, \dots, l \right)$. Since the previous conditions hold for each layer $i$ the notation can be simplified by denoting with:
+
+\begin{equation}
+  R = \text{diag} \left\{ \rho^1, \dots, \rho^l \right\}; \ \Psi = \text{diag}\left\{ f^1, \dots, f^l \right\}; \ \eta = \text{diag} \left\{ \eta^1, \dots, \eta^l \right\}
+\end{equation}
+
+The ETM triggering condition can be rewritten as: $\Psi > R \eta$ in an element-wise fashion.
 

### 2024-12-10 Paper draft/draft.tex
Draft so far

--- 2024-12-08 Paper draft/draft.tex
@@
 \section{Problem Formulation}
-We consider the non-linear discrete-time system stabilized by a Neural Network controller represented in figure \ref{fig:first_scheme} described by:
+We consider the non-linear discrete-time system stabilized by a Neural Network (NN) controller represented in figure \ref{fig:first_scheme} described by:
 \begin{equation}
-  x^{+} = A x + B \text{sat}(\bar{u}) + C \Phi(E x) + D d 
+  x^{+} = A x + B \texttt{sat}(\bar{u}) + C \Phi(E x) + D d 
 \end{equation}
-with $x \in \RR^{n_x}$ the state vector, $\bar{u}$ the raw output of the controller, $\Phi$ the representation of the non-linearity of the system and $d$ a constant disturbance input.
+where the saturation function \texttt{sat} is defined as:
+$$
+    \texttt{sat}(\bar{u}) = \texttt{sign}(\bar{u})\text{min}(|\bar{u}|, \hat{u})
+$$
+with $\hat{u}$ the lmit of the input signal, $x \in \RR^{n_x}$ the state vector, $\bar{u} \in \RR^{n_u}$ the raw output of the controller, $\Phi \in \RR^{n_q}$ the representation of the non-linearity of the system \textbf{ADD MORE DETAILS ABOUT THE NON-LINEARITY SINCE ITS NOLCOS} and $d \in \RR^{n_d}$ a constant disturbance input.
 
@@
     \centering
-    \includegraphics[width=0.65\textwidth]{img/first_scheme}
+    \includegraphics[width=0.45\textwidth]{img/first_scheme}
     \caption{Feedback system}
@@
 
-The controller is implemented as a Multi Layer Perceptron with $l$ layers, each with $n_{\phi_i}$ neurons, all the activation functions are saturation functions, represented by $\phi_i$.
-The objective is to design an event-triggering mechanism that allows to ease the computational burden associated with the computation of many non-linear activation functions providing at the same time sufficient sector conditions to handle the non-linearity of the system and the controller.
+The controller is implemented as a Multi Layer Perceptron (MLP) with $l$ layers, each with $n_{\phi_i}, i \in \left\{ 1, \dots, l\right\}$ neurons, all activation functions are saturation functions, represented by $\phi_i$.
+The objective is to design an Event-Triggering Mechanism (ETM) that allows to ease the computational burden associated with the computation of many non-linear activation functions providing at the same time sufficient sector conditions to handle the non-linearity of the system and the controller.
 The controller $\pi$ is defined by:
@@
 \begin{equation*}
-  \nu_{\phi} = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega_{\phi} = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi_{\nu_{\phi}} = \bmx{\phi_1(\nu^1) \\ \vdots \\ \phi_l(\nu^l)} 
+  \nu = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi = \bmx{\phi_1(\nu^1) \\ \vdots \\ \phi_l(\nu^l)} 
 \end{equation*}
-with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega_\phi = \phi(\nu_\phi)$.
+with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega = \phi(\nu)$.
 Finally, conditions \ref{eq:nn-equations} can be rewritten as:
 \begin{equation*}
-  \bmx{u(k)\\ \nu_\phi(k)} = N \bmx{x(k) \\ \omega_\phi(k) \\ 1} 
+  \bmx{u(k)\\ \nu(k)} = N \bmx{x(k) \\ \omega(k) \\ 1} 
 \end{equation*}
@@
 \end{equation}
-
-Similarly as in \cite{css-extended} we implement an event-triggering-mechanism (ETM) in every layer of the controller, we add also a final ETM to the output that will be used to further reduce the total number of events and call to the non-linear functions.
+Similarly as in \cite{css-extended} we implement an ETM in every layer of the controller, we add also a final one to the output that will be used to further reduce the total number of events and calls to the non-linear functions.
 The system and the controller are now described by figure \ref{fig:second_scheme}.
-
 \begin{figure}[H]
@@
     \includegraphics[width=0.45\textwidth]{img/second_scheme}
+    \centering
     \caption{Feedback system subject to ETM in the controller}
@@
 \end{figure}
-
-The dynamics of the system are now described by:
+The dynamics will be rewritten as:
 \begin{equation}\label{eq:system-dynamics}
@@
 \end{equation}
-with $u$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \ref{eq:first-N} as:
+with $u \in \RR^{n_u}$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \ref{eq:first-N} as:
 \begin{equation}\label{eq:last-N}
@@
 \end{equation}
+All the non-linearity of the system and the controller are selected as saturation functions.
+With respect to \cite{css-extended}, Lemma 3, we declare the local sector conditions that hold for this function:
+
+\emph{For $i \in \left\{ 1, \dots, l \right\}$ given a matrix $G^i \in \RR^{n_{\phi_i} \times n_x}$, if $x(k)$ belongs to the set}
+$$
+S^i = \left\{ x(k) \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}_* \leq G^i (x(k) - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
+$$
+
+\emph{Then the following quadratic constraint holds for any diagonal positive definite matrix $T^i \in \RR^{n_{\phi_i} \times n_{\phi_i}}$:}
+\begin{equation}\label{eqn:first_sec_cond}
+  \left[ \nu^i(k) - \omega^i(k) \right]\tr T^i \left[ G^i (x(k) - x_*) - (\omega^i(k) - \omega^i_*) \right] \leq 0
+\end{equation}
 
 The system is compliant with the assumptions of \textbf{Lemma 2} in \cite{css-extended} that allow us to define the following expressions regarding the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller given $x_*$, defining the matrices:
-
 \begin{equation}
@@
 
-With $N_{\nu \omega}$ always invertible due to its lower triangular structure. We express:
-
+With $N_{\nu \omega}$ always invertible due to its lower triangular structure.
+We find the equilibrium of the state by the use of a numerical solver that minimizes the following expression since it is no longer explicit due to the non-linearity $\Phi$.
+Given a constant disturbance $\bar{d}$:
 \begin{equation}
+  x_* = \min_{x} \left[(A + B R_{\omega} - \Id) x + C \Phi(E x) + D \bar{d} + B R_b \right]
+\end{equation}
+\begin{equation}\label{eqn:equilibrium}
+  \bmx{u_* \\ \nu_*} = N \bmx{x_* \\ \omega_* \\ 1}; \ \  
   \begin{aligned}
-    u_* &= R_{\omega} x_* + R_b\\
+    \omega_* &= \nu_*\\
     \nu_* &= R N_{\nu x} x_* + R N_{\nu b}\\
-    \omega_* &= \nu_*
+    u_* &= R_{\omega} x_* + R_b
+  \end{aligned}
+\end{equation}
+In order to have a lighter notation the dead-zone vector  will be defined as $\psi = \nu - \omega$.
+Furthermore, all following discussions will take into account incremental variables with respect to their equilibrium, denoted with a tilde, i.e.
+$\tilde{x} = x - x_*$.
+Considering \ref{eqn:equilibrium} we have $\psi_* = \nu_* - \omega_* = 0$ hence $\tilde{\psi} = \psi$.
+The sector condition \ref{eqn:first_sec_cond} can be rewritten with no loss of generality as:
+\begin{equation}\label{eqn:sec_cond}
+  \tilde{\psi}\tr T^i \left[ G^i \tilde{x} + \tilde{\psi} - \tilde{\nu} \right] \leq 0
+\end{equation}
+We can express now the new inter-dependency between the variables of interest:
+\begin{equation}\label{eqn:incremental}
+  \begin{aligned}
+    \tilde{u} &= R_{\omega} \tilde{x} - N_{u \omega} R \tilde{\psi}\\
+    \tilde{\nu} &= R N_{\nu x} \tilde{x} + (\Id - R) \tilde{\psi}
   \end{aligned}
@@
 
-The equilibrium state $x^*$ needs some further discussion since the expression to retrieve it is no longer explicit due to the non-linearity $\Phi$.
-We can retrieve the vector using a numerical solver that minimizes the following expression for a given disturbance $\bar{d}$:
+\section{Main Results}
+Taking inspiration from \textbf{Proposition 1} and \textbf{Lemma 3} in \cite{css-extended} we design the ETM for the output of each layer of the controller and the final output.
+The original idea was to trigger an event whenever the sector conditions were violated.
 
+\subsection{Event-triggering mechanism}
+An ETM works by triggering an event that will update the current layer and propagate its value across the network.
+The triggering function will take as inputs the incremental variables of the state of the system
+$\tilde{x}$, the current state of the layer $\tilde{\nu}$ and the dead-zone vector $\tilde{\psi}$. The ETM was defined as:
 \begin{equation}
-  (A + B R_{\omega} - \Id) x_* + C \Phi(E x_*) + D \bar{d} + B R_b = 0
+  \omega^{i}(k) = \begin{cases}
+    \phi_i(\nu^i(k)) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > 0\\
+    \omega^{i}(k-1) & \text{otherwise}
+  \end{cases}
+\end{equation}
+In the precedent work \cite{css-extended} the function $f$ is defined on the basis of the saturation sector conditions in the following quadratic form with respect to \ref{eqn:sec_cond}:
+\begin{equation}\label{eqn:base_omega}
+  f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) =
+  \bmx{\tilde{x}\\ \tilde{\psi} \\ \tilde{\nu}}\tr
+  \Omega^i
+  \bmx{\tilde{x}\\ \tilde{\psi} \\ \tilde{\nu}} > 0
+\end{equation}
+With $$
+\Omega^i =  \bmx{
+    0 & 0 & 0\\
+    T^i G^i & T^i & -T^i\\
+    0 & 0 & 0
+  } 
+$$
+
+This approach acts as a static ETM triggering an event immediately after the sector conditions are violated.
+Inspired by the work of \cite{data-driven} we define a dynamic quantity to be implemented in the ETM triggering condition.
+The idea is to lower the conservatism of the controller by introducing a time-varying threshold that will adapt to the current state of the system reducing the overall number of events and computational burden.
+The new triggering function for the new \emph{dynamic-ETM} is defined as:
+\begin{equation}\label{eqn:dynamic_trig}
+  \omega^{i}(k) = \begin{cases}
+    \phi_i(\nu^i(k)) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta_i\\
+    \omega^{i}(k-1) & \text{otherwise}
+  \end{cases};
 \end{equation}
 
-\section{Main Results}
+Where $\eta^i$ characterized by the following dynamics:
+\begin{equation}\label{eqn:eta_dynamics}
+  \eta^+_i = \rho^i \eta_i - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu})\  \forall i \in \left( 1, \dots, l \right) 
+\end{equation}
 
-\subsection{Event-triggering mechanism}
+\textbf{Remark} \emph{Similarly as in Lemma 3 of \cite{data-driven} we will demonstrate how $\eta^i_k$ being the dynamic threshold of layer $i$ at time instant $k$ is non-negative for any $\rho^i \in [0, 1)$, $\eta^i_0 \geq\ \forall k \in \NN$}
+
+\textbf{Proof} We consider the dynamics \ref{eqn:eta_dynamics} and the ETM triggering rule \ref{eqn:dynamic_trig} we prove the remark by recursion:\\
+\emph{Initial condition}: Assuming $\eta^i_0 \geq 0$ we may have two possible outcomes:
+if $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) < \rho^i \eta^i_0$ we have that 
+\begin{equation}
+  f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) < \rho^i \eta^i_0 \to \rho^i \eta^i_0 - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) = \eta^i_1 \geq 0
+\end{equation}
+
+If instead $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i_0$ then the event is triggered and the values of the current layer state $(\tilde{\psi}, \tilde{\nu})$ are updated to $(\bar{\psi}, \bar{\nu})$, leading again to the first case scenario where $f^i(\tilde{x}, \bar{\psi}, \bar{\nu}) < \rho^i \eta^i_0 \to \eta^i_1 \geq 0$.\\
+\emph{Recursion step}: Assuming $\eta^i_k \geq 0$ similarly as the initial condition we may have:
+$$
+f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) < \rho^i \eta^i_k \to \rho^i \eta^i_k - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) = \eta^i_{k+1} \geq 0
+$$
+If $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i_k$ once again the event is triggered, we update $(\tilde{\psi}, \tilde{\nu}) \to (\bar{\psi}, \bar{\nu})$, the triggering function switches sign $f^i(\tilde{x}, \bar{\psi}, \bar{\nu}) < \rho^i \eta^i_k \to \eta^i_{k+1} \geq 0$.\\
+Hence, $\eta^i_k \geq 0$ implies $\eta^i_{k+1} \geq 0$ concluding the recursion proof.
+
+We denote a condensed notation by taking into account all conditions for each layer
+\begin{equation}
+\begin{aligned}
+   \bR &= \text{diag} \pmx{\rho^1 & \dots & \rho^l }\\
+   \mathbf{\Psi} &= \text{diag} \pmx{ f^1 & \dots & f^l }\\ 
+   \boldsymbol{\eta} &= \text{diag} \pmx{ \eta^1 & \dots & \eta^l }   
+\end{aligned}
+\end{equation}
+
+The ETM triggering condition and dynamic thresholds' dynamics can be rewritten in an element-wise fashion as:
+\begin{equation}
+\begin{aligned}
+    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \mathbf{\Psi}\\
+    \mathbf{\Psi} &> \bR \boldsymbol{\eta}
+\end{aligned}
+\end{equation}
+
+\subsection{Finsler Lemma application}
+After the implementation of the dynamic threshold a further improvement of the triggering condition has been taken in consideration.
+More precisely by denoting $\xi^i = (\tilde{x}, \tilde{\psi}^i, \tilde{\nu}^i)$ we rewrite the even triggering condition $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i$ by taking in consideration \ref{eqn:base_omega} as
+\begin{equation}
+    {\xi^i}\tr \Omega^i \xi^i > \rho^i \eta^i
+\end{equation}
+Intuitively the lower the left-hand term of the inequality the lower will be the number of triggered events.
+Unfortunately $\Omega^i$ structure is non-flexible due to his strict relation with the sector condition expression.
+The following lemma addresses the issue by designing some triggering matrices that guarantee a lower update rate for each layer by the use of Finsler's lemma:
+
+\textbf{Lemma 1}: \emph{Considering the triggering policy \ref{eqn:dynamic_trig}, $R^i$ such that $\bmx{\tilde{x} & \tilde{\psi}^i & \tilde{\nu}^i}\tr = R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr \to {\xi^i}\tr = R^i \xi\tr$ and assuming \ref{eqn:incremental} holds $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists X^i \in \RR^{nx + 2n_{\phi_i} \times nx + 2n_{\phi_i}}, N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ such that:
+\begin{equation}\label{eqn:finsler}
+     {R^i}\tr He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0
+\end{equation}
+Then we have ${\xi^i}\tr X^i \xi^i \leq {\xi^i}\tr \Omega^i \xi^i$}
+
+\textbf{Proof} By pre and post multiplying for $\xi$ the previous expression we end up with
+$$
+    {\xi^i}\tr He\left\{X^i - \Omega^i\right\} \xi^i + He\left\{ \xi\tr \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id} \xi \right\} \leq 0
+$$
+By expanding the term $\pmx{R N_{\nu x} & \Id - R & -\Id} \xi$ we end up with condition \ref{eqn:incremental} meaning that the expression is reduced to
+$$
+    {\xi^i}\tr He(X^i - \Omega^i) \xi^i \leq 0
+$$
+By further expanding the expression and considering that it is eventually reduced to a scalar for which it holds that the transpose is equivalent to itself we obtain:  
+\begin{equation*}
+\begin{aligned}
+    &{\xi^i}\tr (X^i + {X^i}\tr) \xi^i \leq {\xi^i}\tr (\Omega^i + {\Omega^i}\tr) \xi^i\\
+    &2 {\xi^i}\tr X^i\xi^ \leq 2 {\xi^i}\tr \Omega^i \xi^i\\
+    & {\xi^i}\tr X^i \xi^i \leq {\xi^i}\tr \Omega^i \xi^i
+\end{aligned}
+\end{equation*}
+
+Assuming we have the $X^i$ matrices $i \in \left\{1, \dots, l \right\}$ we rewrite the triggering conditions for each layer:
+\begin{equation}
+  \omega^{i}(k) = \begin{cases}
+    \phi_i(\nu^i(k)) & \text{if } {\xi^i}\tr X^i \xi > \rho^i \eta^i(k)\\
+    \omega^{i}(k-1) & \text{otherwise}
+  \end{cases}
+\end{equation}
+By denoting $\bX = \text{diag} \pmx{{R^1}\tr X^1 R^1 & \dots & {R^l}\tr X^l R^l}$, the ETM triggering condition and dynamic thresholds' dynamics can be rewritten in an element-wise fashion as:
+\begin{equation}
+\begin{aligned}
+    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \bX \\
+    \bX &> \bR \boldsymbol{\eta}\end{aligned}
+\end{equation}
 
 \subsection{Lyapunov conditions}
+In this section we will provide sufficient conditions for the stability of the system and the controller under the form of Linear Matrix Inequalities (LMIs) with the guarantee of local asymptotic stability.
+The solution of the problem will also provide the matrices $X^i$ and the parameters $\rho^i$ that will be used in the ETM triggering conditions along with an estimate of the Region Of Attraction (ROA) of the equilibrium point.
+
+We will first introduce a more compact notation for the system dynamics as long as some auxiliary matrices:
+\begin{equation}
+\begin{aligned}
+    \bar{A} &= A + B R_{\omega}\\
+    \bar{B} &= -B N_{u \omega} R\\
+\end{aligned};\ \ 
+R_{\nu} = \bmx{
+  \Id & \0 & \0 & \0\\
+  \0 & \Id & \0 & \0\\
+  R N_{\nu x} & \Id - R & \0 & \0\\
+}
+\end{equation}
+
+\textbf{Theorem 1} \emph{ Consider the control system in \ref{eq:system-dynamics} and the controller $\pi_{\text{ETM}}$ in \ref{eq:nn-equations}.
+Assuming the existence of the matrices $P = P\tr > 0 \in \RR^{n_x \times n_x}$,} $T = \text{diag} \pmx{T^1 & \dots & T^l} \in \RR^{n_{\phi} \times n_{\phi}} > 0$, \emph{$Z = \pmx{{Z^1}\tr & \dots & {Z^l}\tr}\tr \in \RR^{n_{\phi} \times n_x}$, $\bR > 0$, $\bX$, $N^i_1, N^i_2, N^i_3 \ i \in \left\{ 1, \dots, l \right\}$, \textcolor{red}{Aggiungi condizioni sui settori} and a scalar $\alpha > 0$ such that the following matrix inequalities hold:}
+\begin{equation}\label{eqn:lyapunov}
+  \bmx{
+    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C & 0\\
+    \star & \bar{B}\tr P \bar{B} & \bar{B}\tr P C & 0\\
+    \star & \star & C\tr P C & 0\\
+    \star & \star & \star & 2 (\bR - \Id)
+  } + R_s\tr \bmx{
+    S & Q\\
+    \star & W
+  }  R_s - (\1_l \otimes R_{\nu})\tr He(\bX) (\1_l \otimes R_{\nu}) < 0
+\end{equation}
+\begin{equation}\label{eqn:inclusion}
+  \bmx{
+    P & {Z_j^i}\tr\\
+    \star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
+  } \geq 0 \ \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\}
+\end{equation}
+\begin{equation}\label{eqn:finsler-conditions}
+  {R^i}\tr He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\}
+\end{equation}
+\emph{Where $\hat{\nu}_j^i = \min(|-\bar{\nu}_j^i - \nu^i_{*, j}|,|\bar{\nu}_j^i - \nu^i_{*, j}|)$, $Z^i = T^i G^i,\ i \in \left\{ 1, \dots, l \right\} $}
 

### 2024-12-12 Paper draft/draft.tex
Ora funziona

@@
-\documentclass{article}
-\input{../math_commands}
-\input{../custom_imports}
-\usepackage{biblatex}
-\addbibresource{../biblio.bib}
+\documentclass{ifacconf}
+%%%put this lines in ifacconf before hyperref to use it
+\makeatletter
+\let\old@ssect\@ssect % Store how ifacconf defines \@ssect
+\makeatother
+\input{math_commands}
+\input{custom_imports}
+\usepackage{natbib}
+%\usepackage{booktabs}
+%%%put this lines in ifacconf after hyperref to use it
+\makeatletter
+\def\@ssect#1#2#3#4#5#6{%
+  \NR@gettitle{#6}% Insert key \nameref title grab
+  \old@ssect{#1}{#2}{#3}{#4}{#5}{#6}% Restore ifacconf's \@ssect
+}
+\makeatother
+
+%%theorems
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{proposition}{Proposition}
+\newtheorem{assumption}{Assumption}
+\newtheorem{lemma}{Lemma}
+\newtheorem{remark}{Remark}
+\newtheorem{definition}{Definition}
+\newtheorem{corollary}{Corollary}
+\newtheorem{problem}{Problem}
+\newenvironment{proof}{\paragraph*{Proof:}}{\hfill$\square$}
 
 \begin{document}
+\begin{frontmatter}
 
@@
 
-\author{Marco Sterlini}
+\author[First]{Marco Sterlini}
+\author[Second]{Samuele Zoboli}
+\author[Second]{Sophie Tarbouriech}
+
+\address[First]{Università di Trento, Trento, Italy (e-mail: marco.sterlini@studenti.unitn.it) }
+\address[Second]{LAAS-CNRS, Université de Toulouse, CNRS, Toulouse, France (e-mail:bu) }
+
+\begin{abstract}
+\textbf{INSERT ABSTRACT}
+\end{abstract}
+
+\begin{keyword}
+\textbf{INSERT KEYWORDS}
+\end{keyword}
+
+\end{frontmatter}
 
 \section{Introduction}
+\textbf{INSERT INTRODUCTION}\\
+\emph{Notation:} $\NN, \RR^{n}, \RR^{n \times m}$ denote the set of natural non-negative integers, the set of real vectors of dimension $n$ and the set of real matrices of dimension $n \times m$, respectively.
+For any matrix $A$, $A\tr$ is its transpose.
+For any square matrix $A$, the operator $\He{ A } = A + A\tr$ is defined.
+$\texttt{diag}(A_1, A_2)$ is a block-diagonal matrix with block-diagonal matrices $A_1$ and $A_2$.
+For a partitioned matrix, the symbol $\star$ stands for symmetric blocks.
+$\1_n = \bmx{1 & \dots & 1}-\tr \in \NN^{n \times 1}$
 
@@
 $$
-with $\hat{u}$ the lmit of the input signal, $x \in \RR^{n_x}$ the state vector, $\bar{u} \in \RR^{n_u}$ the raw output of the controller, $\Phi \in \RR^{n_q}$ the representation of the non-linearity of the system \textbf{ADD MORE DETAILS ABOUT THE NON-LINEARITY SINCE ITS NOLCOS} and $d \in \RR^{n_d}$ a constant disturbance input.
+with $\hat{u}$ the lmit of the input signal, $x \in \RR^{n_x}$ the state vector, $\bar{u} \in \RR^{n_u}$ the raw output of the controller, $\Phi$ the representation of the non-linearity of the system \textcolor{red}{Please help in adding details about $\Phi$ since it's NOLCOS} and $d \in \RR^{n_d}$ a constant disturbance input.
 
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{img/first_scheme}
+    \includegraphics[width=0.45\textwidth]{Figures/first_scheme}
     \caption{Feedback system}
@@
 with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega = \phi(\nu)$.
-Finally, conditions \ref{eq:nn-equations} can be rewritten as:
+Finally, conditions \eqref{eq:nn-equations} can be rewritten as:
 \begin{equation*}
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{img/second_scheme}
+    \includegraphics[width=0.45\textwidth]{Figures/second_scheme}
     \centering
@@
 \end{equation}
-with $u \in \RR^{n_u}$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \ref{eq:first-N} as:
+with $u \in \RR^{n_u}$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \eqref{eq:first-N} as:
 \begin{equation}\label{eq:last-N}
@@
 \end{equation}
-All the non-linearity of the system and the controller are selected as saturation functions.
-With respect to \cite{css-extended}, Lemma 3, we declare the local sector conditions that hold for this function:
 
-\emph{For $i \in \left\{ 1, \dots, l \right\}$ given a matrix $G^i \in \RR^{n_{\phi_i} \times n_x}$, if $x(k)$ belongs to the set}
-$$
-S^i = \left\{ x(k) \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}_* \leq G^i (x(k) - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
-$$
-
-\emph{Then the following quadratic constraint holds for any diagonal positive definite matrix $T^i \in \RR^{n_{\phi_i} \times n_{\phi_i}}$:}
-\begin{equation}\label{eqn:first_sec_cond}
-  \left[ \nu^i(k) - \omega^i(k) \right]\tr T^i \left[ G^i (x(k) - x_*) - (\omega^i(k) - \omega^i_*) \right] \leq 0
-\end{equation}
-
-The system is compliant with the assumptions of \textbf{Lemma 2} in \cite{css-extended} that allow us to define the following expressions regarding the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller given $x_*$, defining the matrices:
+The system is compliant with the assumptions of \citep[Lemma 2]{css-extended} that allow us to define the following expressions regarding the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller given $x_*$, defining the matrices:
 \begin{equation}
@@
 \end{equation}
+All the non-linearity of the system and the controller are selected as saturation functions.
+With respect to \citep[Lemma 3]{css-extended}, we declare the local sector conditions that hold for this function:
+
+\emph{For $i \in \left\{ 1, \dots, l \right\}$ given a matrix $G^i \in \RR^{n_{\phi_i} \times n_x}$, if $x(k)$ belongs to the set}
+$$
+S^i = \left\{ x(k) \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}_* \leq G^i (x(k) - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
+$$
+
+\emph{Then the following quadratic constraint holds for any diagonal positive definite matrix $T^i \in \RR^{n_{\phi_i} \times n_{\phi_i}}$:}
+\begin{equation}\label{eqn:first_sec_cond}
+  \left[ \nu^i(k) - \omega^i(k) \right]\tr T^i \left[ G^i (x(k) - x_*) - (\omega^i(k) - \omega^i_*) \right] \leq 0
+\end{equation}
+Where $\bar{\nu}^i$ is the saturation level of the $i$-th layer.
+
+\textcolor{red}{Proof needed?}
+
 In order to have a lighter notation the dead-zone vector  will be defined as $\psi = \nu - \omega$.
@@
 $\tilde{x} = x - x_*$.
-Considering \ref{eqn:equilibrium} we have $\psi_* = \nu_* - \omega_* = 0$ hence $\tilde{\psi} = \psi$.
-The sector condition \ref{eqn:first_sec_cond} can be rewritten with no loss of generality as:
+Considering \eqref{eqn:equilibrium} we have $\psi_* = \nu_* - \omega_* = 0$ hence $\tilde{\psi} = \psi$.
+The sector condition \eqref{eqn:first_sec_cond} can be rewritten with no loss of generality as:
 \begin{equation}\label{eqn:sec_cond}
@@
 \section{Main Results}
-Taking inspiration from \textbf{Proposition 1} and \textbf{Lemma 3} in \cite{css-extended} we design the ETM for the output of each layer of the controller and the final output.
-The original idea was to trigger an event whenever the sector conditions were violated.
+
+Taking inspiration from \citep[Proposition 1, Lemma 3]{css-extended} we design the ETM for the output of each layer of the controller and the final output.
 
@@
 \end{equation}
-In the precedent work \cite{css-extended} the function $f$ is defined on the basis of the saturation sector conditions in the following quadratic form with respect to \ref{eqn:sec_cond}:
+In the precedent work \cite{css-extended} the function $f$ is defined on the basis of the saturation sector conditions in the following quadratic form with respect to \eqref{eqn:sec_cond}:
 \begin{equation}\label{eqn:base_omega}
@@
 Inspired by the work of \cite{data-driven} we define a dynamic quantity to be implemented in the ETM triggering condition.
-The idea is to lower the conservatism of the controller by introducing a time-varying threshold that will adapt to the current state of the system reducing the overall number of events and computational burden.
-The new triggering function for the new \emph{dynamic-ETM} is defined as:
+The idea is to lower the conservatism of the controller by introducing a time-varying threshold that will adapt to the current state of the system and controller reducing the overall number of events and computational burden.
+The new triggering function for the new \emph{dynamic-ETM} for the $i$-th layer is defined as:
 \begin{equation}\label{eqn:dynamic_trig}
@@
 
-\textbf{Remark} \emph{Similarly as in Lemma 3 of \cite{data-driven} we will demonstrate how $\eta^i_k$ being the dynamic threshold of layer $i$ at time instant $k$ is non-negative for any $\rho^i \in [0, 1)$, $\eta^i_0 \geq\ \forall k \in \NN$}
+\begin{remark} \emph{Similarly as in \citep[Lemma 3]{data-driven} we will demonstrate how $\eta^i_k$ being the dynamic threshold of layer $i$ at time instant $k$ is non-negative for any $\rho^i \in [0, 1)$, $\eta^i_0 \geq\ \forall k \in \NN$} \end{remark}
 
-\textbf{Proof} We consider the dynamics \ref{eqn:eta_dynamics} and the ETM triggering rule \ref{eqn:dynamic_trig} we prove the remark by recursion:\\
+\begin{proof} We consider the dynamics \eqref{eqn:eta_dynamics} and the ETM triggering rule \eqref{eqn:dynamic_trig} we prove the remark by recursion:\\
 \emph{Initial condition}: Assuming $\eta^i_0 \geq 0$ we may have two possible outcomes:
@@
 If $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i_k$ once again the event is triggered, we update $(\tilde{\psi}, \tilde{\nu}) \to (\bar{\psi}, \bar{\nu})$, the triggering function switches sign $f^i(\tilde{x}, \bar{\psi}, \bar{\nu}) < \rho^i \eta^i_k \to \eta^i_{k+1} \geq 0$.\\
-Hence, $\eta^i_k \geq 0$ implies $\eta^i_{k+1} \geq 0$ concluding the recursion proof.
+Hence, $\eta^i_k \geq 0$ implies $\eta^i_{k+1} \geq 0$ concluding the recursion proof. \end{proof}
+\textcolor{red}{Is the recursion really necessary? Can't I just do the proof for any k?}
 
@@
    \mathbf{\Psi} &= \text{diag} \pmx{ f^1 & \dots & f^l }\\ 
-   \boldsymbol{\eta} &= \text{diag} \pmx{ \eta^1 & \dots & \eta^l }   
+   \boldsymbol{\eta} &= \bmx{ {\eta^1}\tr & \dots & {\eta^l}\tr }\tr   
 \end{aligned}
@@
 \begin{aligned}
-    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \mathbf{\Psi}\\
-    \mathbf{\Psi} &> \bR \boldsymbol{\eta}
+    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \mathbf{\Psi} \1_l\\
+    \mathbf{\Psi} \1_l &> \bR \boldsymbol{\eta}
 \end{aligned}
@@
 After the implementation of the dynamic threshold a further improvement of the triggering condition has been taken in consideration.
-More precisely by denoting $\xi^i = (\tilde{x}, \tilde{\psi}^i, \tilde{\nu}^i)$ we rewrite the even triggering condition $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i$ by taking in consideration \ref{eqn:base_omega} as
+More precisely by denoting $\xi^i = (\tilde{x}, \tilde{\psi}^i, \tilde{\nu}^i)$ we rewrite the even triggering condition $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i$ by taking in consideration \eqref{eqn:base_omega} as
 \begin{equation}
@@
 
-\textbf{Lemma 1}: \emph{Considering the triggering policy \ref{eqn:dynamic_trig}, $R^i$ such that $\bmx{\tilde{x} & \tilde{\psi}^i & \tilde{\nu}^i}\tr = R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr \to {\xi^i}\tr = R^i \xi\tr$ and assuming \ref{eqn:incremental} holds $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists X^i \in \RR^{nx + 2n_{\phi_i} \times nx + 2n_{\phi_i}}, N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ such that:
+\begin{lemma}: \emph{Considering the triggering policy \eqref{eqn:dynamic_trig}, $R^i$ such that $\bmx{\tilde{x} & \tilde{\psi}^i & \tilde{\nu}^i}\tr\!
+= \!R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr\!
+\to {\xi^i}\tr = R^i \xi\tr$ and assuming \eqref{eqn:incremental} holds $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists \!
+X^i\!
+\in\!
+\RR^{nx + 2n_{\phi_i} \times nx + 2n_{\phi_i}}, $ $N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ such that:
 \begin{equation}\label{eqn:finsler}
-     {R^i}\tr He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0
+     {R^i}\tr\! He\left\{X^i - \Omega^i\right\}\! R^i + He\left\{ \smallpmat{N^i_1\\ N^i_2\\ N^i_3}\! \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0
 \end{equation}
 Then we have ${\xi^i}\tr X^i \xi^i \leq {\xi^i}\tr \Omega^i \xi^i$}
+\end{lemma}
 
-\textbf{Proof} By pre and post multiplying for $\xi$ the previous expression we end up with
+\begin{proof} By pre and post multiplying for $\xi$ the previous expression we end up with
 $$
-    {\xi^i}\tr He\left\{X^i - \Omega^i\right\} \xi^i + He\left\{ \xi\tr \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id} \xi \right\} \leq 0
+    {\xi^i}\tr\! He \!\left\{\!X^i \!-\! \Omega^i\right\} \xi^i + He\!\left\{ \xi\tr\! \smallpmat{N^i_1\\ N^i_2\\ N^i_3}\! \smallpmat{R N_{\nu x} & \Id - R & -\Id} \xi\! \right\}\! \leq 0
 $$
-By expanding the term $\pmx{R N_{\nu x} & \Id - R & -\Id} \xi$ we end up with condition \ref{eqn:incremental} meaning that the expression is reduced to
+By expanding the term $\pmx{R N_{\nu x} & \Id - R & -\Id} \xi$ we end up with condition \eqref{eqn:incremental} meaning that the expression is reduced to
 $$
@@
 \end{equation*}
+\end{proof}
 
@@
 \begin{aligned}
-    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \bX \\
-    \bX &> \bR \boldsymbol{\eta}\end{aligned}
+    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \bX \1_l\\
+    \bX \1_l&> \bR \boldsymbol{\eta}\end{aligned}
 \end{equation}
@@
 We will first introduce a more compact notation for the system dynamics as long as some auxiliary matrices:
-\begin{equation}
+\begin{equation}\label{eqn:auxiliary}
 \begin{aligned}
@@
     \bar{B} &= -B N_{u \omega} R\\
-\end{aligned};\ \ 
-R_{\nu} = \bmx{
-  \Id & \0 & \0 & \0\\
-  \0 & \Id & \0 & \0\\
-  R N_{\nu x} & \Id - R & \0 & \0\\
-}
+    \tilde{x}^+ &= \bar{A}\tilde{x} + \bar{B}\tilde{\psi} + C \tilde{\Phi}\\
+R_{\nu} &= \bmx{
+  \Id & 0 & 0 \\
+  0 & \Id & 0 \\
+  R N_{\nu x} & \Id - R & 0\\
+} \in \RR^{n_x + n_{\phi} + n_q \times n_x + 2 n_{\phi}}\\
+R_s &\in \RR^{n_x + n_{\phi} + n_q \times n_x + n_q}
+\end{aligned}
 \end{equation}
 
-\textbf{Theorem 1} \emph{ Consider the control system in \ref{eq:system-dynamics} and the controller $\pi_{\text{ETM}}$ in \ref{eq:nn-equations}.
-Assuming the existence of the matrices $P = P\tr > 0 \in \RR^{n_x \times n_x}$,} $T = \text{diag} \pmx{T^1 & \dots & T^l} \in \RR^{n_{\phi} \times n_{\phi}} > 0$, \emph{$Z = \pmx{{Z^1}\tr & \dots & {Z^l}\tr}\tr \in \RR^{n_{\phi} \times n_x}$, $\bR > 0$, $\bX$, $N^i_1, N^i_2, N^i_3 \ i \in \left\{ 1, \dots, l \right\}$, \textcolor{red}{Aggiungi condizioni sui settori} and a scalar $\alpha > 0$ such that the following matrix inequalities hold:}
-\begin{equation}\label{eqn:lyapunov}
-  \bmx{
-    \bar{A}\tr P \bar{A} - P & \bar{A}\tr P \bar{B} & \bar{A}\tr P C & 0\\
-    \star & \bar{B}\tr P \bar{B} & \bar{B}\tr P C & 0\\
-    \star & \star & C\tr P C & 0\\
-    \star & \star & \star & 2 (\bR - \Id)
+\begin{theorem} \emph{ Consider the control system in \eqref{eq:system-dynamics} and the controller $\pi_{\text{ETM}}$ in \eqref{eq:nn-equations}.
+Assuming the existence of the matrices $P = P\tr > 0 \in \RR^{n_x \times n_x}$,} $T = \text{diag} \pmx{T^1 & \dots & T^l} \in \RR^{n_{\phi} \times n_{\phi}} > 0$, \emph{$Z = \smallpmat{{Z^1}\tr & \dots & {Z^l}\tr}\tr \in \RR^{n_{\phi} \times n_x}$, $\bR > 0$, $\bX$, $N^i_1, N^i_2, N^i_3 \ i \in \left\{ 1, \dots, l \right\}$, $S \in \RR^{n_x \times n_x}, Q \in \RR^{n_x \times n_q}, W \in \RR^{n_q \times n_q}$ and a scalar $\alpha > 0$ such that the matrix inequalities \eqref{eqn:lyapunov}, \eqref{eqn:inclusion}, \eqref{eqn:finsler} hold, where $\hat{\nu}_j^i = \min(|-\bar{\nu}_j^i - \nu^i_{*, j}|,|\bar{\nu}_j^i - \nu^i_{*, j}|)$, $Z^i = T^i G^i,\ i \in \left\{ 1, \dots, l \right\}$\\
+Then the system controlled by $\pi_{\text{ETM}}$ is locally stable around $x_*$ and the set $\cE(P, x_*) = \left\{ x \in \RR^{n_x} : \tilde{x}\tr P \tilde{x} \leq 1\right\}$ is an estimate of the ROA.}
+\end{theorem}
+
+\begin{figure*}
+\begin{align}\label{eqn:lyapunov}
+  &\bmx{
+  \begin{array}{c|c}
+  \bmx{\bar{A}\tr\\
+  \bar{B}\tr\\
+  C\tr
+  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
+  P & 0 & 0\\
+  0 & 0 & 0\\
+  0 & 0 & 0
   } + R_s\tr \bmx{
@@
     \star & W
-  }  R_s - (\1_l \otimes R_{\nu})\tr He(\bX) (\1_l \otimes R_{\nu}) < 0
-\end{equation}
-\begin{equation}\label{eqn:inclusion}
-  \bmx{
+  }  R_s - (\1_l \otimes R_{\nu})\tr He(\bX) (\1_l \otimes R_{\nu}) \hspace{.1em}& 0\\[1.5em]
+\hline \\[-0.9em]
+  0 & \hspace{.1em} 2 (\bR - \Id)
+  \end{array}} < 0\\
+  &\bmx{
     P & {Z_j^i}\tr\\
     \star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
-  } \geq 0 \ \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\}
-\end{equation}
-\begin{equation}\label{eqn:finsler-conditions}
-  {R^i}\tr He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\}
-\end{equation}
-\emph{Where $\hat{\nu}_j^i = \min(|-\bar{\nu}_j^i - \nu^i_{*, j}|,|\bar{\nu}_j^i - \nu^i_{*, j}|)$, $Z^i = T^i G^i,\ i \in \left\{ 1, \dots, l \right\} $}
+  } \geq 0 \ \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\} \label{eqn:inclusion}\\
+  &{R^i}\tr\!
+He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}
+\end{align}
+\vspace{.5em}
+\hrule
+\end{figure*}
 
+\begin{proof} We start by taking into account a candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + He(\1_l\tr\!\boldsymbol{\eta})$.
+We start by commenting $R_{\nu}$ and $R_s$ in \eqref{eqn:auxiliary}, we notice how $\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\nu}\tr}\tr = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}\tr}\tr$.
+The matrix $\bmx{S & Q\\ \star & W}$ represents the quadratic sector conditions specific to $\Psi$ with respect to $\tilde{x}$ and $\tilde{\Psi}$, $R_s$ is a transformation matrix such that: $\bmx{\tilde{x}\tr & \tilde{\Psi}}\tr = R_s \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}\tr}\tr$
+\end{proof}
 \subsection{Optimization procedure}
@@
 
-\printbibliography
+%\bibliographystyle{plain}
+\bibliography{Bibliography/biblio}
 

### 2024-12-12 Paper draft/Paper_ifac.tex
Rename forlders

\documentclass{ifacconf}

\usepackage{natbib}
\usepackage{graphicx}     
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{enumerate}
\usepackage{svg}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[hidelinks]{hyperref}
\usepackage{svg}
\usepackage{subcaption}
\input{math_commands}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{remark}{Remark}
\newtheorem{problem}{Problem}
\newtheorem{example}{Example}
\newtheorem{property}{Property}

\newcommand{\bphi}{\pmb{\phi}}

\begin{document}
	\begin{frontmatter}

		\title{Noninvasive dynamic output-feedback stabilization of multistable systems} 
		% Title, preferably not more than 10 words.

		\thanks[footnoteinfo]{This research is carried out by the infinite utility of $k$-contraction}

		\author[First]{Samuele Zoboli} 		
        \author[Second]{Andreu Cecilia}
	\author[First]{Sophie Tarbouriech}

		\address[Second]{Universitat Polit\`ecnica de Catalunya, Avinguda Diagonal, 647, 08028 Barcelona, Spain, (e-mail: andreu.cecilia@upc.edu)}
  \address[First]{LAAS-CNRS, Universit\'e de Toulouse, CNRS, Toulouse, (e-mail: samuele.zoboli@laas.fr, sophie.tarbouriech@laas.fr)}

		\begin{abstract}                % Abstract of not more than 250 words.
   Artificial intelligence, machine learning, engaging keywords.
		\end{abstract}

		\begin{keyword}
			LMI, Controller design, Matrix inertia, Nonlinear systems, Contraction, $k$-contraction.
		\end{keyword}

	\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
{\color{red} To be written}

\medskip
\emph{Notation:}  $\RR_{\geq0}\defeq[0,\infty)$ and
$\NN \defeq \{0,1,2,\ldots\}$. $|\cdot|$ denotes the standard Euclidean norm.
Given $x\in \RR^n$, $y\in \RR^m$, we set 
$(x,y)\defeq (x^\top , y^\top)^\top$.
The inertia of a matrix $P$ 
is defined by the triplet of integers
$
\cine(P) \defeq (\pi_{-}(P),  \pi_{0}(P), \pi_{+}(P)),
$
where $\pi_{-}({P})$, $\pi_{+}(P)$ and $\pi_{0}(P)$ denote the numbers of eigenvalues of $P$ with negative,
positive and zero real part, respectively, counting multiplicities.

\section{Problem formulation}
We consider nonlinear systems of the form
\begin{equation}\label{eqn:original_system}
\begin{aligned}
\dot x &=   A x  + D\phi(Ex) + Bu,  \\
 y&= Cx+ G\phi(Ex),
\end{aligned}
\end{equation}
where $x\in\RR^{n_x}$ is the state vector,  $u\in\RR^{n_u}$ the controlled input, $y\in\RR^{n_y}$ the measured outputs and $\phi:\RR^{n_v}\to\RR^{n_v}$ is a sufficiently smooth vector field. Additionally, to recover efficiently solvable conditions, we assume the existence of known matrices $R=R^\top, Q=Q^\top$ and $S$ such that 
\begin{equation}\label{eqn:generic_abstraction}
    \pmx{\Id_{n_x}\\\pderiv{\phi}{x}(v)}^\top\pmx{R&S^\top\\S&Q}\pmx{\Id_{n_x}\\\pderiv{\phi}{x}(v)}\preceq 0 
\end{equation}
holds for all $v\in\cV\subseteq \RR^{n_v}$ where $\cV\defeq\{v\in\RR^{n_v}:v=E x, x\in\cX\subseteq \RR^{n_x}\}$. We recall that such a class of nonlinearities includes common and interesting family of functions, such as (shifted) monotonic and differentially sector-bounded ones \citep{zoboli2024quadratic}. 
Finally, we assume that in the unforced case, namely when $u=0$, system \eqref{eqn:original_system} presents a set of fixed points $\cF_x^\star$. More formally, we assume the existence of a nonempty set of equilibria
\begin{equation}\label{eqn:eq_points}
 \cF_x^\star := \{x\in\RR^{n_x} \ | \ Ax+D\phi(E x)= 0\}.
\end{equation}
We highlight that $\cF^\star$ can include multiple elements, that is, system \eqref{eqn:original_system}  may present multiple equilibrium points for the unforced case $u=0$.

We aim to design a feedback control policy that makes (at least) one point $x^\star\in\cF_x^\star$ asymptotically stable. In other words, we want to make the set $\cF_x^\star$ attractive and forward-invariant.

One possibility is to reformulate the problem as a standard stabilization one, see e.g. \citep{isidori2017stabilization}. Differently put, the problem could be solved by selecting a specific point  $x^\star\in\cF_x^\star$ and designing a control law making $x^\star$ globally asymptotically stable. However, we highlight that the considered problem relaxes the more common stabilization problem in three directions. First, the goal is not to stabilize the system at a specific point in $\cF_x^\star$, we only need the system to converge to $\cF_x^\star$. Second, the proposed objective allows for the existence of multiple attractive equilibrium points for the closed-loop system.  Third, the set of equilibrium points $\cF_x^\star$ may be unknown to the designer. This third point is particularly relevant for feedback policies that include an integral action, e.g. \citep{Astolfi2017,GIACCAGLI2023111106}, which often require an unknown non-zero steady-state control signal to be generated by the integral component.
Another possibility is to invoke set-stability approaches, e.g. \citep{SHIRIAEV20001709}. However, we remark that $\cF_x^\star$ may be disconnected. Moreover, its elements are given by the (possibly unknown) equilibria of the unforced system dynamics. These facts impose additional restrictions that can impact the effectiveness and applicability of set-stability approaches.

Given the abovementioned constraints, in what follows, we present a solution that does not require knowledge of the open-loop equilibria and guarantees their stability via $2$-contraction \citep{Michael1995}. Our approach is based on the recent findings in \citep{ cecilia2023generalized, zoboli2024dynamic} and inspired by the techniques of derivative control and washout filters.

\section{Main result}
\subsection{Proposed approach}
We start by highlighting a necessary property controllers must satisfy to guarantee the open-loop equilibria are preserved by the closed loop.
From \eqref{eqn:original_system}, it is clear that for the closed-loop system to stabilize at  $x^\star\in\cF_x^\star$ the control input must vanish at such point.
% dynamic output-feedback law $u=\alpha(y,\xi)$ with state $\xi\in\RR^{n_\xi}$ must satisfy 
% \begin{equation}\label{eqn:noninvasive_cond}
% \alpha(Cx^\star+G\phi(Ex^\star),\xi^\star) = 0, \qquad \forall x^\star \in \cF^\star,
% \end{equation}
% where $\xi^\star$ is the equilibrium state of the controller corresponding to $x^\star$.
We will denote a control policy that satisfies such a property as \textit{noninvasive}, e.g. \citep{ANGELI2024152}. 
\begin{definition}[Noninvasive controller]
Consider the nonlinear system 
\[
\dot x = f(x,u), \qquad y =h(x),
\]
with $x\in\RR^{n_x}, y\in\RR^{n_y}$ and $u\in\RR^{n_u}$. Let 
\[
\cF^\star\defeq\{x\in\RR^{n_x}:f(x,0)=0\}.
\]
The dynamic output-feedback law $u=\alpha(y,\xi)$  with state $\xi\in\RR^{n_\xi}$ is said to be \textit{ noninvasive} if 
\begin{equation}\label{eqn:noninvasive_cond}
\alpha(h(x^\star),\xi^\star) = 0, \qquad \forall x^\star \in \cF^\star,
\end{equation}
where $\xi^\star$ is the controller equilibrium associated to $x^\star$.
\end{definition}

In this paper, we propose ensuring such a requirement via integral action. More specifically, we extend the dynamics of \eqref{eqn:original_system} as follows
\begin{equation}\label{eqn:extended_dyn}
\begin{aligned}
\pmx{\dot x\\ \dot z}&= \pmx{A & 0\\0&0} \pmx{x\\ z} + \pmx{B\\\Id_{n_u}} u + \pmx{D\\0} \phi\left(\pmx{E&0} \pmx{x\\ z} \right),  \\
 \pmx{y\\z}&= \pmx{C & 0\\0 & \Id_{n_u}}\pmx{x\\ z} + \pmx{G\\0} \phi\left(\pmx{E&0}\pmx{x\\ z} \right).
\end{aligned}
\end{equation}
At any equilibrium point $(x^\star,z^\star)$ the $z$-dynamics impose $u=0$, and thus $x^\star \in \cF_x^\star$.

\subsection{Controller design}
Given the above discussion, we aim to design a dynamic-output feedback controller stabilizing the extended dynamics \eqref{eqn:extended_dyn} to a (possibly nonunique) equilibrium. Moreover, we leverage 2-contraction to obtain local exponential stability of equilibria \citep{Muldowney1990}. In particular, we exploit the techniques of \citep{zoboli2024dynamic} to efficiently design a noninvasive linear dynamic output-feedback controller guaranteeing 2-contractivity of the extended closed loop. More formally, we select $n_\xi = n_x+n_u$ and we consider the extended system
\begin{equation}
\begin{array}{rl}
        \label{eqn:nonlin_sys_ctrl}
        \dot x &=  A x +  D\phi(Ex) + Bu, \\
        y&= Cx + G\phi(Ex),\\
        \dot z &= u,\\
        \dot \xi &= M\xi + N y + Fz,\\
        u&=K y + L\xi + Hz,
    \end{array}
\end{equation}
which can be equivalently represented as
\begin{subequations}
\begin{equation}
\begin{array}{rl}
        \label{eqn:nonlin_sys_closed_loop}
        \dot \eta= f(\eta)=\bA \eta + \bD\phi(\bE\eta)
    \end{array}
\end{equation}
with $\eta = \col(x,z,\xi)\in\RR^{n_\eta}$, $n_\eta = n_x+n_u+n_\xi = 2n_\xi$ and
\begin{equation}
\begin{array}{rl}
        \label{eqn:nonlin_sys_closed_loop_matrices}
        % \bA&= \pmx{\smallpmat{A & 0 \\ 0& 0}+\smallpmat{B\\\Id_{n_z}}K\smallpmat{C & F} & \smallpmat{B\\\Id_{n_z}}L\\ N\smallpmat{C & F} & M},  \\
        %  \bD &= \pmx{D_e  + \smallpmat{B\\\Id_{n_z}}KG \\ NG}, \quad
        %  \bE = \pmx{E & 0 & 0}.
        %  \\
          \bA&= \pmx{A+BKC & BH & BL\\
          KC & H & L\\
          NC & F & M},  \\
         \bD &= \pmx{D+BKG\\KG\\NG}, \quad
         \bE = \pmx{E & 0 & 0}.
    \end{array}
\end{equation}
\end{subequations}
We also remark that \eqref{eqn:generic_abstraction} implies the existence of matrices $\bR,\bS,\bQ$ such that
\begin{equation}\label{eqn:generic_abstraction_ctrl}
    \pmx{\Id_{n_\eta}\\\pderiv{\phi}{\eta}(\bE\eta)}^\top\pmx{\bR&\bS^\top\\\bS&\bQ}\pmx{\Id_{n_\eta}\\\pderiv{\phi}{\eta}(\bE\eta)}\preceq 0
\end{equation}
holds for all $\eta\in\cS\subseteq \RR^{n_\eta}$ with $\cS\defeq\{\eta\RR^{n_\eta}: \bE\eta\in\cV\}$, see \citep[Section 2.5]{zoboli2024dynamic}. 
The following result proposes a design for noninvasive linear dynamic output-feedback controllers based on matrix inequalities.
\begin{figure*}
    \begin{equation}\label{eqn:2_contraction_BMI}
    \begin{array}{rl}
      A_e&\defeq \pmx{A&0\\0&0},\qquad  B_e \defeq \pmx{B\\\Id_{n_u}},\qquad C_e \defeq \pmx{C&0\\0&\Id_{n_u}},\qquad 
    D_e \defeq  \pmx{D\\0}, \qquad G_e \defeq  \pmx{G\\0}, \quad 
        \Phi \defeq \pmx{\Pi_2 & \Id_{n_\xi}\\ \Id_{n_\xi} & \Pi_1},
     \\&\\
        \widehat A &\defeq\pmx{A_e\Pi_2+B_e \widehat L && A_e + B_e  \widehat K C_e \\ \widehat M && \Pi_1 A_e+ \widehat N  C_e }, \quad  
        \widehat D \defeq \pmx{D_e +B_e \widehat K G_e\\ \Pi_1 D_e  + \widehat N G_e}, 
        \quad  \widehat R \defeq \pmx{ \widehat R_{11} &  \widehat R_{12}^\top\\
        \star & \widehat R_{22}}, \quad \widehat S^\top \defeq \pmx{\widehat S_1^\top \\ \widehat S_2^\top},
        \\&\\
        &
        \pmx{\widehat R_{11} - \overbar R& \widehat R_{12}^\top - \alpha \Pi_2 R^\top & \widehat S_{1}^\top-\alpha \Pi_2 S^\top\\
         \star & \widehat R_{22}-\alpha R & \widehat S_{2}^\top -\alpha S^\top \\
         \star&\star& \widehat Q-\alpha Q} +
         \He{\pmx{\Upsilon_1\\\Upsilon_2\\\Upsilon_3}\pmx{\pderiv{\phi}{x}(v)& 0 & -\Id_{n_x}}}\preceq 0 , \qquad \forall v \in \cV,
        \\&\\
           & \pmx{
   \He{A_e \Pi_2+B_e \widehat L} -2\beta_1 \Pi_2 - \widehat R_{11} && A_e+B_e \widehat K C_e  + \widehat M^\top-2\beta_1 \Id_{n_\xi} -  \widehat R_{12}^\top && D_e +B_e \widehat K G_e -  \widehat S_1^\top\\
   \star && \He{ \Pi_1 A_e +  \widehat N  C_e } -2\beta_1 \Pi_1   -\widehat R_{22}&& \Pi_1 D_e  + \widehat N G_e -\widehat S_2\\
   \star &&\star&& - \widehat Q
    }\prec 0,\\&\\
         &\pmx{
    -2\beta_0 \Sigma+ \He{\Gamma_1 \widehat A} - \widehat R & & \Sigma- \Gamma_1 \Phi+ \widehat A^\top \Gamma_2^\top  &  &\Gamma_1 \widehat D +  \widehat A^\top \Gamma_3^\top - \widehat S^\top\\
    \star  && -\He{\Gamma_2 \Phi}  & &\Gamma_2 \widehat D -\Phi\Gamma_3^\top \\
    \star && \star& &\He{\Gamma_3 \widehat D} - \widehat Q
    }\prec 0  .\\&\\
    \end{array}    
    \end{equation}
    \hrule
\end{figure*}
\begin{proposition}
    \label{prp:2_contr_BMI}
     Consider  \eqref{eqn:nonlin_sys_closed_loop} with \eqref{eqn:nonlin_sys_closed_loop_matrices} %\eqref{eqn:closed_loop}
and let \eqref{eqn:generic_abstraction_ctrl} hold for all  $x\in\cF$  with given matrices $R, S, Q$. Select $Z\preceq 0$ such that $Z\preceq R$ and assume there exist $ \overbar R_{11}=\overbar R^\top,$ $ \widehat R_{11}=\widehat R_{11}^\top,$ $ \widehat R_{12},$ $ \widehat R_{22}=\widehat R_{22}^\top,$ $ \widehat S_1,$ $ \widehat S_2,$ $ \widehat Q=\widehat Q^\top,$ scalars $\alpha>0,\beta_0\in\RR$, $\beta_1<0$,  symmetric matrices $\Pi_1=\Pi_1^\top, \Pi_2=\Pi_2^\top \in \RR^{n_\xi\times n_\xi}$, matrices $\widehat K \in \RR^{n_u\times n_y}$, $\widehat L\in \RR^{n_u\times n_\xi}$, $\widehat M\in \RR^{n_\xi\times n_\xi}$, $\widehat N \in \RR^{n_\xi\times n_\xi}$,  matrices $\Gamma_1, \Gamma_2, \Gamma_3$  and $\Sigma=\Sigma^\top\succ 0$  such that 
\begin{enumerate}[i)]
        \item The matrix $ R_Z \defeq \smallpmat{\overbar R && \alpha \Pi_2Z\\ \star && \alpha Z}$ is negative semidefinite;
         \item Inequality \eqref{eqn:2_contraction_BMI}\footnotemark[1] holds with $\beta_0+\beta_1<0$;
        \item $\cine(\Gamma_2)=(1, 0, 2n_\xi-1)$.
    \end{enumerate}
    \footnotetext[1]{Given at the top of the next page.}
     Then, 
     there exist square invertible matrices $U,V\in \RR^{n_\xi\times n_\xi}$ such that $UV^\top = \Id_{n_\xi}- \Pi_1\Pi_2$
    and a scalar $\Delta^\star>0$ such that the control matrices 
    \begin{equation}
    \label{eqn:ctrl_matrices}
    \begin{array}{rl}
          \pmx{K & H}&= \widehat K,  \\
         \pmx{N& F}&= U^{-1} (\widehat N - \Pi_1 B \widehat K) ,\\
           L & = (\widehat L - \widehat K C \Pi_2) V^{-\top},\\
         M&=U^{-1}\!\pmx{\Id_{n_\xi}\\-\Pi_1}^{\!\top}\!\!\pmx{\widehat M & \widehat N C\\
         B \widehat L & -A+ B\widehat  KC}\pmx{\Id_{n_\xi}\\-\Pi_2}V^{-\top}\!\!\!\!\!\!
    \end{array}
\end{equation}
    ensure all trajectories of any perturbed closed loop
\begin{equation}
\begin{array}{rl}
        \label{eqn:nonlin_sys_closed_loop_pert}
        \dot \eta= \widetilde f(\eta) =\bA \eta + \bD\phi(\bE\eta)+ \delta(\eta)
    \end{array}
\end{equation} 
evolving in a forward-invariant bounded compact set $\cK\subseteq\cS$ with  $\delta(\eta)=\col(\delta_x(\eta),0,0)$,  $\delta_x(\eta)\in \cC^1$, $|\pderiv{\delta_x}{\eta}(\eta)|<\Delta^\star$  for all $\eta \in \cK$,
converge to an equilibrium $\eta^\star\in \cF^\star_\eta$ with
    \begin{equation}
        \label{eqn:closed_loop_eq_set}
        \cF^\star_\eta \defeq \{\eta\in\cK: Ax+D\phi(E x) + \delta_x(x)= 0\}.
    \end{equation}
\end{proposition}
The proof of Proposition~\ref{prp:2_contr_BMI} is presented in Section~\ref{sec:proof_prp1}.

Controllers issued from Proposition~\ref{prp:2_contr_BMI} ensure convergence of the closed-loop trajectories to open-loop equilibria  even in the presence of (sufficiently slow-varying) state-dependent uncertainties. This robust noninvasive behavior differs from the one in \citep{ANGELI2024152}, where the open-loop dynamics must be perfectly known, or the derivative of the state must be measurable.  Moreover, while the matrices $R,S,Q$ in \eqref{eqn:generic_abstraction} (and therefore $\bR,\bS,\bQ$ in \eqref{eqn:generic_abstraction_ctrl}) are assumed to be known, extensions of Proposition~\ref{prp:2_contr_BMI} to cases of unknown quadratic abstractions can be proposed by following the results in \citep[Section 2]{zoboli2024dynamic}.

Nonetheless, Proposition~\ref{prp:2_contr_BMI} presents two main drawbacks. Firstly, it requires imposing matrix inertia constraints, see item iii). A detailed discussion on methods to impose such constraints in the optimization problem is presented in \citep[Section 3]{zoboli2024dynamic}. Secondly, it requires closed-loop trajectories to evolve in a forward invariant bounded compact set $\cK$. This necessity arises from the fact that the proposed result is grounded on the notion of $2$-contraction, which guarantees convergence to nonunique equilibria if the system evolves in forward invariant bounded compact sets \citep[Theorem 2.5]{Michael1995}, \citep[Lemma 6]{cecilia2023generalized}. Therefore, we now present an LMI-based approach allowing the \textit{a-posteriori} verification of such a requirement.

\begin{figure*}
    \begin{equation}
    \label{eqn:ellipsoid_LMI}
    \pmx{\tau_1 P_{\cE_1}- \tau_2 P_{\cE_2} + \bA_s^\top P_{\cE_1} + P_{\cE_1} \bA_s - \bX && P_{\cE_1}  -\bY^\top && 0\\\star && - \bZ&&0\\
    \star && \star  && -\tau_1 + \tau_2 }\prec 0.
\end{equation}
    \vspace{.5em}
\hrule
\end{figure*}
\begin{proposition}\label{prp:invariant_set}
    Consider \eqref{eqn:nonlin_sys_closed_loop} and let $\bA = \bA_s +\widetilde \bA$ with $\bA_s$ Hurwitz. Suppose there exist matrices $\bX=\bX^\top$, $\bY$, $\bZ=\bZ^\top$, a set $\cS\subseteq\RR^{n_\eta}$ and a vector $\eta^\star\in\cS$ such that the local sector condition
    \begin{equation}
        \label{eqn:local_sect_eta_reshaped}
        \pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}^\top\pmx{\bX &\bY^\top\\\bY&\bZ}\pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}\le 0
    \end{equation}
    holds for all $\eta\in\cS$, with $\psi(v)\defeq \bD \phi(v)+\widetilde \bA\eta$. 
    If there exist positive scalars $\tau_1$, $\tau_2$, two matrices $P_{\cE_1}\succeq P_{\cE_2}\succ 0$ and a matrix $\Omega\in\RR^{n_\eta \times n_u}$ such that \eqref{eqn:ellipsoid_LMI}\footnotemark[2] holds and such that the ellipsoidal set $\cE_2\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_2} (\eta-\eta^\star) \le 1\} \subseteq\cS$, then $\cE_1\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_1} (\eta-\eta^\star) \le 1\}$ is  attractive and forward invariant for \eqref{eqn:nonlin_sys_closed_loop} with region of attraction including the  non-empty set $\widetilde\cE\defeq \cE_2\setminus\cE_1$, for all $\eta\in \cS$.
\end{proposition}
The proof of Proposition~\ref{prp:invariant_set} is presented in Section~\ref{sec:proof_prp2}. 
\footnotetext[2]{Given at the top of the next page.}

\begin{remark}
    Note that the loop transformation $\bA=\bA_s+\widetilde \bA$,  $\psi(v)\defeq \bD \phi(v)+\widetilde \bA\eta$ is a necessary step for the satisfaction of \eqref{eqn:ellipsoid_LMI}. Indeed, after the removal of the quadratic abstraction entry $\bX$, the top-left entry of \eqref{eqn:ellipsoid_LMI} requires
    \[
    \tau_1 P_{\cE_1}- \tau_2 P_{\cE_2} + \bA_s^\top P_{\cE_1} + P_{\cE_1} \bA_s \prec 0.
    \]
    Since $P_{\cE_1}\succeq P_{\cE_2}$ this implies
    \[
    (\tau_1-\tau_2) P_{\cE_1} + \bA_s^\top P_{\cE_1} + P_{\cE_1} \bA_s \prec 0,
    \]
    which requires $\bA_s$ Hurwitz, given that $\tau_1>\tau_2$ from the bottom-right entry of \eqref{eqn:ellipsoid_LMI}. Moreover, since the pair $(\bA, \bD)$ may not be controllable, the more classical loop transformation involving the addition and subtraction of a virtual stabilizing input $u_s = \bK\eta$ and the definition of a reshaped nonlinearity $\psi(v)=\phi(v) - \bK \eta$ may not be applicable. Thus, $\psi(v)$ must include $\bD$ in its definition.
\end{remark}
%
While useful, the results of Proposition~\ref{prp:invariant_set} present two main drawbacks. First, they require the knowledge of a nondifferential quadratic abstraction \eqref{eqn:local_sect_eta_reshaped} for the \emph{reshaped closed loop}. Second, they do not guarantee the outer ellipsoid $\cE_2\subseteq\cS$. We first address the issues emerging from the loop transformation by proposing a result connecting quadratic abstractions on the original system to ones on the reshaped loop.
\begin{lemma}\label{lem:orig2reshaped_abs}
    Consider \eqref{eqn:nonlin_sys_closed_loop} and suppose there exist matrices $X=X$, $Y$, $Z=Z$, a set $\cS\subseteq\RR^{n_\eta}$ and a vector $\eta^\star\in\cS$ such that the local sector condition
    \begin{equation}
        \label{eqn:local_sect_eta}
        \pmx{\eta - \eta^\star\\ \phi(\bE\eta)-\phi(\bE\eta^\star)}^\top\pmx{X &Y^\top\\Y&Z}\pmx{\eta - \eta^\star\\ \phi(\bE\eta)-\phi(\bE\eta^\star)}\le 0
    \end{equation}
    holds for all $\eta\in\cS$. If there exists a matrix $\Omega\in\RR^{n_\eta\times n_v}$ such that
        \begin{equation}
    \label{eqn:D_full}
    \pmx{\bD^\top \Omega + \Omega^\top \bD && \Omega^\top\\
    \star && \Id_{n_u}}\succ 0,
\end{equation}
then $\bD^\top\bD\succ 0$ and \eqref{eqn:local_sect_eta_reshaped} holds with
\begin{equation}
    \label{eqn:original2reshaped_abs_matrices}
    \begin{array}{rl}
        \overbar \bD &= (\bD^\top \bD)^{-1}\bD^\top,\\
         \bX &= X- Y^\top \overbar \bD\widetilde \bA - \widetilde \bA^\top  \overbar \bD^\top Y - \widetilde \bA^\top \overbar \bD^\top Z \overbar \bD\widetilde \bA,\\
         \bY &= \overbar \bD Y + \overbar \bD^\top Z \overbar \bD\widetilde \bA,\\
         \bZ &= \overbar \bD^\top Z \overbar \bD.
    \end{array}
\end{equation}
\end{lemma}
The proof of Lemma~\ref{lem:orig2reshaped_abs} is presented in Section~\ref{sec:proof_lem1}.

Lemma~\ref{lem:orig2reshaped_abs} allows translating quadratic abstractions for the original closed loop to abstractions for the transformed one. We highlight that constraint \eqref{eqn:D_full} can be dropped by assuming $\bD$ has full column rank.

While representing an important step in improving the applicability of
Proposition~\ref{prp:invariant_set}, Lemma~\ref{lem:orig2reshaped_abs} requires knowledge of matrices $X,Y,Z$ in \eqref{eqn:local_sect_eta} and it does not guarantee the outer ellipsoid $\cE_2\subseteq\cS$.
One possible workaround to both issues is to exploit the knowledge of the differential abstraction \eqref{eqn:generic_abstraction} to derive a global incremental one. This is formalized in the following lemma.
\begin{lemma}
    \label{lem:diff2incr_abstraction}
    If there exist matrices $R_v,S_v, Q_v$ with $Q_v\succ 0$ such that $\phi:\RR^{n_v}\to\RR^{n_v}$ satisfies
    \begin{equation}\label{eqn:diff_abstr}
    \pmx{\Id_{n_v}\\\pderiv{\phi}{v}(v)}^\top\pmx{R_v&S_v^\top\\S_v&Q_v}\pmx{\Id_{n_v}\\\pderiv{\phi}{v}(v)}\preceq 0,
    \end{equation}
    for all $v\in \RR^{n_v}$, then 
    \begin{equation}\label{eqn:incr_abstr}
        \pmx{v_1-v_2\\\phi(v_1)-\phi(v_2)}^\top\pmx{R_v&S_v^\top\\S_v&Q_v}\pmx{v_1-v_2\\\phi(v_1)-\phi(v_2)}\preceq 0,
    \end{equation}
    holds for all $v_1,v_2\in\RR^{n_v}$.
\end{lemma}
The proof of Lemma~\ref{lem:diff2incr_abstraction} is presented in Section~\ref{sec:proof_lem2}.

Given the relation $v=Ex=\bE\eta$, Lemma~\ref{lem:diff2incr_abstraction} shows that one can select $X = \gamma\bR = \gamma \bE^\top R_v \bE$, $Y=\gamma \bS = \gamma \bE S_v$ and $Z=\gamma \bQ = \gamma Q_v$ in \eqref{eqn:ellipsoid_LMI} for any $\gamma>0$, see for instance \citep[Section 2.5]{zoboli2024dynamic}. Indeed, since \eqref{eqn:incr_abstr} holds for all $v_1,v_2$, it holds also for $v_2=\bE\eta^\star$ and obtain $\cS=\RR^{n_\eta}$ in Proposition~\ref{prp:invariant_set}, ensuring $\cE_2\subseteq\cS$. Then, Lemma~\ref{lem:orig2reshaped_abs} allows mapping the obtained abstraction to \eqref{eqn:local_sect_eta_reshaped}. Given \eqref{eqn:original2reshaped_abs_matrices}, the requirement of $Q_v\succ0$ does not restrict the set of solutions to the LMI, as the central block of \eqref{eqn:ellipsoid_LMI} imposes the same constraint in the form $\bZ\succ 0$.

Unfortunately, the exploitation of Lemma~\ref{lem:diff2incr_abstraction} is very conservative and still presents drawbacks. Importantly, \eqref{eqn:incr_abstr} holds for any pair $v_1,v_2$. As such, the resulting abstraction represents an incremental sector condition rather than a local one as per \eqref{eqn:local_sect_eta_reshaped}. Therefore, while holding also for the ellipsoid center $\eta^\star$, it is not specific to such a point. This may lead to increased conservatives and numerical problems since the resulting ellipsoids are not anchored to a specific point, i.e., they could be interpreted as incremental ones.

Given the above discussion, we argue that estimating the matrices $\bX,\bY,\bZ$ given a desired open-loop center $x^\star$ is a more reliable solution. This approach only requires minor knowledge of local sector bounds for the nonlinearity $\phi$. Moreover, it allows including constraints guaranteeing $\cE_2$ is included in the known local sector. We formalize the proposed technique in the following theorem.

\begin{figure*}
    \begin{subequations}
    \label{eqn:local_sector_and_inclusion}
    \begin{equation}
    \label{eqn:original2reshaped_abs_matrices}
    \begin{array}{rl}
         \widehat X &= \bE^\top X_v\bE - \bE^\top Y_v^\top \overbar \bD\widetilde \bA - \widetilde \bA^\top  \overbar \bD^\top Y_v \bE - \widetilde \bA^\top \overbar \bD^\top Z_v \overbar \bD\widetilde \bA,\\
         \widehat Y &= \overbar \bD Y + \overbar \bD^\top Z \overbar \bD\widetilde \bA,\\
         \widehat Z &= \overbar \bD^\top Z \overbar \bD.
    \end{array}
\end{equation}
\begin{align}
    \label{eqn:local_sec_open_loop_polytopic}
    &\pmx{\Id_{n_v}\\\Lambda}^\top\pmx{X_v&Y_v^\top\\Y_v&Z_v}\pmx{\Id_{n_v}\\\Lambda} \preceq 0,  \qquad \forall \Lambda: \Theta_1\preceq \Lambda\preceq\Theta_2\\
    \notag\\
\label{eqn:local_sec_closed_loop_polytopic}
    &\pmx{ 0 && 0 && 0\\
     0 && \bX - \widehat X && \bY^\top - \widehat Y^\top \\
     \0 && \bY - \widehat Y && \bZ- \widehat Z }+\He{\pmx{\Upsilon_1 \\ \Upsilon_2\\
    \Upsilon_3 
    }\pmx{\Id_{n_\eta} & -\bA & \Id_{n_v}}}\preceq 0\\\notag\\
    \label{eqn:ellipsoid_inclusion_polytopic}
    &\pmx{P_{\cE_2} &\bE_i^\top\\ \bE_i & (\overbar v -\underline v)_i^2}\succeq 0, \qquad i=1, \dots, n_v.
\end{align}
\hrule
\end{subequations}
\end{figure*}
\begin{theorem}\label{thm:invariant_ellipsoid_local_sect}
Let $\Theta_1,\Theta_2\in\RR^{n_v\times n_v}$ be given diagonal matrices such that $\Theta_1\preceq\Theta_2$. Assume there exist vectors $\underline v, \overbar v, x^\star$ such that $v^\star = Ex^\star \in[\underline v, \overbar v]$ and
\begin{equation}
    \label{eqn:local_sector_v}
   (\phi(v)-\phi(v^\star) -\Theta_1(v-v^\star))_i
   (\phi(v)-\phi(v^\star) -\Theta_2(v-v^\star))_i\le0
\end{equation}
for all $i=1,\dots,n_v$ and for all $v\in[\underline v, \overbar v]$, where $(\cdot)_i$ represents the $i^{\text{th}}$ element of the argument vector or the $i^{\text{th}}$ row of the argument matrix. If there exist positive scalars $\tau_1$, $\tau_2$, a matrix $\Omega\in\RR^{n_\eta\times n_v}$, two matrices $P_{\cE_1}\succeq P_{\cE_2}\succ 0$, matrices $\Upsilon_i$ with $i=,\dots, 8$  and matrices $X_v=X_v^\top$, $Y_v$, $ Z_v= Z_v^\top$, $\bX=\bX^\top$, $\bY$, $ \bZ= \bZ^\top$ such that \eqref{eqn:ellipsoid_LMI} , \eqref{eqn:D_full}
and \eqref{eqn:local_sector_and_inclusion} hold wit $\overbar \bD$ as in \eqref{eqn:original2reshaped_abs_matrices}, then $\cE_1\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_1} (\eta-\eta^\star) \le 1\}$ is  attractive and forward invariant for \eqref{eqn:nonlin_sys_closed_loop} with region of attraction including the  non-empty set $\widetilde\cE\defeq \cE_2\setminus\cE_1$, with 
${\color{red}\eta^\star=\col(x^\star, 0, 0)}$ and
$\cE_2\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_2} (\eta-\eta^\star) \le 1\}$.
\end{theorem}
The proof of Theorem~\ref{thm:invariant_ellipsoid_local_sect} is presented in Section~\ref{sec:proof_thm1}.

Theorem~\ref{thm:invariant_ellipsoid_local_sect} considers the abstraction matrices $\bX,\bY,\bZ$, $X_v, Y_v, Z_v$ as problem variables. To reduce the size of the optimization problem, one could select $\bX=\widehat X$,  $ \bY =  \widehat Y$  and $\bZ= \widehat Z$, thus removing \eqref{eqn:local_sec_closed_loop_polytopic}. However, the inclusion of \eqref{eqn:local_sec_closed_loop_polytopic} leaves more degrees of freedom to the optimization problem, which can exploit entries in the directions of the integrator and controller states. Under the selection mentioned previously, these entries would be set to $0$ by the structure of $\bE$. We also highlight that \eqref{eqn:local_sec_open_loop_polytopic} is typically hard to solve. However, if we constrain the diagonal elements of $Z_v$ to be nonpositive, it is sufficient to check the condition for all
matrices $\Lambda$ corresponding to the finitely many vertices
of the matrix interval $[\Theta_1,\Theta_2]$ \citep{ebihara2021stability}. This can be easily done b, e.g., decomposing $Z_v\defeq Z_d + \widetilde Z^\top + \widetilde Z$ with diagonal $Z_d\preceq 0$ and lower triangular $\widetilde Z$.

\begin{remark}
    Since the differential quadratic abstraction \eqref{eqn:generic_abstraction} holds for all $v\in\cV$ and given the degree of freedom arising from letting matrices $X_v,Y_v,Z_v$ as decision variables, one can select $\underline v, \overbar v$ and compute $\Theta_1,\Theta_2$ for Theorem~\ref{thm:invariant_ellipsoid_local_sect} such that the resulting polytope is included in $\cV$. This choice connects the regions where 2-contractivity holds and the invariant ellipsoid is looked for.
\end{remark}

\section{Comparison with the literature}\label{sec:comparison_lit}
We highlight that alternative solutions to the considered stabilization problem can be found on the literature. Thus, we find it necessary to discuss the main differences between these results and the proposal of this work.

The main property that characterizes a noninvasive control is that the input signal is zero at steady-state. In this paper, we achieved this property through the noninvasive condition in \eqref{eqn:noninvasive_cond}. An alternative way for obtaining zero steady-state input is to design a feedback policy that uses the derivative of the state, that is $u = \alpha(\dot x)$ with $\alpha(0)=0$. This approach have been explored by some authors under the name of state-derivative feedback \citep{ZAHEER2021109495,Arthur2019} and was generalized to its derivative-output version in \citep{ZAHEER2021109495,ARTHUR202040}. We highlight that these results only provide design conditions for linear systems or linearized systems around a unique equilibrium point. Thus, these approaches have very limited applicability to nonlinear systems with multiple uncertain equilibrium points or with initial conditions far from an equilibrium point.

Recently, the authors in \citep{ANGELI2024152} applied $2$-contraction theory for designing state-derivative controllers. $2$-contraction enabled the development of non-local design conditions for multistable systems. A similar framework to the one proposed in \citep{ANGELI2024152} can be recovered from \eqref{eqn:original_system} by selecting $C=A, G=D$ and the static control policy $u=Ky$. Nonetheless, the proposal in \citep{ANGELI2024152} is not free of limitations. First, it requires a direct measurement of the derivative of the state or an exact knowledge of the system vector field. Second, the design conditions are based on mathematical objects known as matrix compounds \citep{bar2023compound}, which usually explode in dimension and may destroy advantageous structural properties of the system.

Our proposal can be understood as an extension of the washout filter discussed in \citep{GARRARD1977497, WANG19951213, Hassouneh2004}. Indeed, we achieve a zero steady-state input by designing the controller as a linear filter with a blocking zero in the origin. Thus, we do not require any measurement of the state or output derivative, and we can operate directly with the system output. 
Nevertheless, we remark that our approach is not free from tradeoffs. Indeed, the addition of input integrators does not allow any of the system's outputs to be regulated. From output regulation theory, the number of regulated outputs must be smaller than or equal to the number of inputs, see e.g. \citep{isidori2003robust}. By adding integrator dynamics on all input signals, such a regulation limit is reached.
We also stress that the addition of input-integrator dynamics may cause the loss of controllability properties. This is shown by the following simple example.
Consider the planar linear time invariant system
\begin{equation}
    \label{eqn:ctrlb_example_sys}
    \begin{array}{rl}
     \dot x_1&= x_1+u  \\
     \dot x_2 &= x_1 
\end{array}
\end{equation}
with $x=\col(x_1,x_2)\in\RR^2$ and $u\in\RR$. It can be easily verified that \eqref{eqn:ctrlb_example_sys} is controllable. However, by adding the input-integrator dynamics, the resulting extended system
\[
    \begin{array}{rl}
     \dot x_1&= x_1+u  \\
     \dot x_2 &= x_1 \\
     \dot z &= u,
\end{array}
\]
has a simply stable uncontrollable mode. However, the $2$-contraction framework proposed in this paper only requires a generalized notion of stabilizability properties \citep{cecilia2023generalized}.
We also highlight that, different from previous results with washout filter, the combination of $2$-contraction \citep{cecilia2023generalized} with the design conditions provided in \citep{zoboli2024dynamic} provides for the first time LMI design conditions for the proposed stabilization problem which are:
\begin{itemize}
    \item non-local (in the sense that the state can be initialized far from any equilibrium point);
    \item can be implemented in nonlinear systems with multiple equilibria;
    \item can be implemented just with the system output.
\end{itemize}

\section{Numerical simulation}
As an illustration of the proposed we consider a second order system of the form \eqref{eqn:original_system} with the following matrices
$$
\begin{aligned}
    A = \begin{bmatrix}
        0 & 1\\ 0 & 0
    \end{bmatrix}, \ B = \begin{bmatrix}
        0 \\ 1
    \end{bmatrix}, \ D = \begin{bmatrix}
        0 \\ 1
    \end{bmatrix}, \ C = E= \begin{bmatrix}
        1 & 0
    \end{bmatrix}, G = 1
\end{aligned}
$$
and the following nonlinearity:
$$
\phi(v)=\cos(v+\theta),
$$
where $\theta\in\RR$ is a constant unknown by the designer. For this particular system, defining $x_1$ and $x_2$ and $x = \begin{bmatrix}
    x_1 & x_2
\end{bmatrix}^\top$, we have the following set of equilibrium points:
$$
\cF^\star = \{x\in\RR^2 \ | \ \cos(x_1+\theta)=0, \ x_2 = 0\}.
$$
In the unforced case $u=0$, the considered system can be understood as a frictionless pendulum with $x_1$ being the pendulum angle shifted by $\theta$ and $x_2$ is the angular velocity. Therefore, the set $\cF^\star$ is not attractive and for any initial condition outside $\cF^\star$ the system converges to a non-trivial limit cycle. With this in mind, the objective is to make the set $\cF^\star$ attractive and invariant.

We highlight that the proposed stabilization problem can be trivially solved using the state-derivative feedback discussed in Section~\ref{sec:comparison_lit}. Indeed, any feedback term of the form $u = -\kappa x_2$, with a sufficiently large $\kappa>0$, would solve the considered stabilization problem, since at any equilibrium we have $x_2 = 0$. Nonetheless, we highlight that we have no direct measure of the states or its derivative, only of the signal $y = x_1 + \cos(x_1+\theta)$. Thus, the state derivative feedback methodology cannot be directly implemented. Naturally, one could try to design an observer for the system and then use the estimation of the second state, denoted $\hat x_2$, as the state derivative feedback, that is  $u = -\kappa \hat x_2$. Nonetheless, the design of such an observer is not trivial due to the unknown constant $\theta$. If one tries to, somehow, design an observer robust to the unknown parameter $\theta$, it will significantly increase the noise sensitivity of the feedback policy. Additionally, the extended system:
$$
\begin{aligned}
    \dot x_1 &= x_2 \\
    \dot x_2 &= \cos(x_2+x_3) + u \\
    \dot x_3 &= 0 \\
    y &= x_1 + \cos(x_1+x_3)
\end{aligned}
$$
is not observable at any equilibrium point with $u=0$, since we have $\dot y = 0$ and $y = x_1 + \cos(x_1+x_3) = x_1 + \cos(x_1+n x_3)$ for any positive and odd integer $n$. Thus, the design of an observer that jointly estimates the states $x$ and the unknown parameter $\theta$ is not a trivial task if we impose a zero steady-state input. Considering these difficulties, the purpose of this section is to show how the problem can be directly solved with the framework proposed in this work.

Controller:
\begin{equation*}
\begin{aligned}
    M &= \smallbmat{
        0.29 &  -2.04  &  1.45\\
   -2.04 &  -3.34 &  -0.97\\
    1.45 &  -0.97  &  1.23},  &N &= \smallbmat{
        -13.91  & -0.17\\
   -3.45  & -1.42 \\
   10.52 &  -4.3}, \\
    K &= \smallbmat{
        -3.77 &  -4.38},  
    &L &= \smallbmat{
        -16.02 &  -0.68 & -13.8},
\end{aligned}
\end{equation*}

\begin{figure}
\includesvg[width=1\linewidth]{Figures/closed-loop}
 \caption{Constant theta case, closed loop}
         \label{fig:constant_theta}
\end{figure}

\begin{figure*}
	\centering
	\begin{subfigure}{0.42\linewidth}
		\includesvg[width=\linewidth]{Figures/open_loop_theta_change}
	\end{subfigure}
	\begin{subfigure}{0.42\linewidth}
		\includesvg[width=\linewidth]{Figures/closed_loop_theta_change}
	\end{subfigure}
 \begin{subfigure}{0.42\linewidth}
		\includesvg[width=\linewidth]{Figures/closed_loop_theta_change_zoom}
	\end{subfigure}
 \caption{Open-loop and closed-loop with no noise and noise (variance $0.05$.)}
         \label{fig:theta_change}
\end{figure*}

\section{Proofs}
\subsection{Proof of Proposition~\ref{prp:2_contr_BMI}}\label{sec:proof_prp1}
Note that the extended matrices $A_e, B_e, C_e, D_e, G_e$ in \eqref{eqn:2_contraction_BMI} describe the $(x,z)$-dynamics in \eqref{eqn:extended_dyn}. Therefore, by \citep[Theorem 2]{zoboli2024dynamic} adapted with the result on known quadratic abstraction of \citep[Corollary 2]{zoboli2024dynamic}, the matrix inequalities in \eqref{eqn:2_contraction_BMI} paired with items i)-iii) of Proposition~\ref{prp:2_contr_BMI} guarantee $\cine(\Sigma_1)=(1,0,2n_\eta-1)$ and
\begin{equation}
\label{eqn:2_contr_cond_closed_loop}
\begin{cases}
     \left( \bA+ \bD\pderiv{\phi}{\eta}(\bE\eta) \right)^\top P_0 +  P_0 \left(\bA+ \bD\pderiv{\phi}{\eta}(\bE\eta) \right) \prec 2\mu_0 P_0,\\
     \left( \bA+ \bD\pderiv{\phi}{\eta}(\bE\eta) \right)^\top P_1 +  P_1 \left(\bA+ \bD\pderiv{\phi}{\eta}(\bE\eta) \right) \prec 2\mu_1 P_1,\\
     \mu_0+\mu_1<0
\end{cases}
\end{equation}
for all $\eta\in\cK$ with $P_0=\Sigma_0, P_1=\Sigma_1, \beta_0=\mu_0,\beta_1=\mu_1$. 
Since $\cK$ is bounded and compact, there exist strictly positive scalars $\varepsilon_0,\varepsilon_1$ such that
\[
\begin{cases}
    \pderiv{ f}{\eta}(\eta)^\top P_0 +  P_0 \pderiv{ f}{\eta}(\eta) \preceq 2\mu_0 P_0 - \varepsilon_0 \Id_{n_\eta},\\
    \pderiv{ f}{\eta}(\eta)^\top P_1 +  P_1 \pderiv{ f}{\eta}(\eta)\preceq 2\mu_1 P_1 - \varepsilon_1 \Id_{n_\eta},\\
     \mu_0+\mu_1<0.
\end{cases}
\]
Note that $|\pderiv{\delta}{\eta}(\eta)|=|\pderiv{\delta_x}{\eta}(\eta)|$ for all $\eta \in \cK$. By \citep[Lemma 8]{zoboli2024dynamic} and the relative proof, there exist strictly positive $\gamma_0,\gamma_1$ small enough and $\Delta^\star_0\in (0,\varepsilon_0\gamma_0),\,  \Delta^\star_1\in(0,\varepsilon_1\gamma_1)$ such that 
\[
\begin{cases}
    \pderiv{\widetilde f}{\eta}(\eta)^\top P_0 +  P_0 \pderiv{\widetilde f}{\eta}(\eta) \prec 2\overbar \mu_0 P_0,\\
     \pderiv{\widetilde f}{\eta}(\eta)^\top P_1 +  P_1 \pderiv{\widetilde f}{\eta}(\eta) \prec 2\overbar \mu_1 P_1,
\end{cases}
\]
with $ \overbar \mu_0 = \mu_0+\frac{\gamma_0}2, \overbar \mu_1 = \mu_1+\frac{\gamma_1}2$. Let such upperbounds on $\gamma_0,\gamma_1$ be denoted by $\overbar \gamma_0, \overbar \gamma_1$. Since $\mu_0+\mu_1<0$, by continuity there exists $\overbar \gamma>0$ such that  $\mu_0+\mu_1+\overbar \gamma<0$. By selecting $ \gamma_0=\gamma_1 = \min(\overbar \gamma_0, \overbar \gamma_1, \overbar \gamma)$ and  $\Delta^\star < \min(\varepsilon_0 \gamma, \varepsilon_1  \gamma)$ we obtain 
\[
\begin{cases}
    \pderiv{\widetilde f}{\eta}(\eta)^\top P_0 +  P_0 \pderiv{\widetilde f}{\eta}(\eta) \prec 2\overbar \mu_0 P_0,\\
     \pderiv{\widetilde f}{\eta}(\eta)^\top P_1 +  P_1 \pderiv{\widetilde f}{\eta}(\eta) \prec 2\overbar \mu_1 P_1,\\
     \overbar \mu_0 + \overbar \mu_1<0.
\end{cases}
\]
Consequently, \citep[Theorem 5]{cecilia2023generalized} proves 2-contractivity of the closed loop \eqref{eqn:nonlin_sys_closed_loop_pert} and \citep[Theorem 2.5]{Michael1995} paired with boundedness and forward- invariance of $\cK$ guarantees the trajectories of \eqref{eqn:nonlin_sys_closed_loop} converge to a (possibly nonunique) equilibrium point. The result is proven by noting that all equilibria of \eqref{eqn:nonlin_sys_closed_loop_pert} satisfy $0=\smallpmat{KC &H& L}\eta + KG \phi(\bE \eta)=u$.

\subsection{Proof of Proposition~\ref{prp:invariant_set}} \label{sec:proof_prp2}
% If \eqref{eqn:nonincremental_abstraction} holds, \eqref{eqn:ellipsoid_LMI} and $\gamma>0$ imply
% \[
%     \pmx{\tau P_\cE+ \bA^\top P_\cE + P_\cE \bA  & P_\cE \bB  & 0\\\bB^\top P_\cE  & 0 &0\\
%     0 & 0 & -\tau \lambda}\preceq 0.
% \]
% Therefore, left and right multiplication by the vector $\nu=\col(\eta,\phi(\bE\eta),1)$ and its transpose paired with the dynamics \eqref{eqn:nonlin_sys_closed_loop}  yields
% \[
% \dot \eta^\top P_\cE \eta + \eta^\top P_\cE \dot \eta +\tau (\eta^\top P_\cE\eta-\lambda)\le 0.
% \]
% On the boundary of $\cE$ we have $\eta^\top P_\cE\eta = \lambda$, thus implying 
% \[
% \dot \eta^\top P_\cE \eta + \eta^\top P_\cE \dot \eta\le 0.
% \]
% Consider the function $V(\eta)=\eta^\top P_\cE \eta$.  Given the above inequality, we have $\dot V(\eta) = 0$  for all $\eta\in \partial\cE$ where $\partial\cE  =\{\eta\in\RR^{n_\eta}:V(\eta)=\eta^\top P_\cE\eta =\lambda\}$. Therefore, all trajectories starting in $\cE$ remain in $\cE$, thus proving forward-invariance.

{
\label{sec:proof_prp2}
Let $\widetilde \eta \defeq \eta-\eta^\star$, $\widetilde \phi \defeq \phi(\bE\eta)-\phi(\bE\eta^\star)$ and $\widetilde \psi \defeq \psi(\bE\eta)-\psi(\bE\eta^\star) = \bD\widetilde \phi + \widetilde \bA\widetilde \eta $.
According to the dynamics \eqref{eqn:nonlin_sys_closed_loop} and the relation $\bA=\bA_s+\widetilde\bA$, we have
\[
\dot{\widetilde\eta} = \bA_s\widetilde \eta +\widetilde \psi.
\]
If \eqref{eqn:local_sect_eta_reshaped} holds for all $\eta\in\cS$, right and left multiplication of \eqref{eqn:ellipsoid_LMI} by $\nu=\col(\widetilde \eta,\widetilde\psi)$ and its transpose implies
\[
    \nu^\top\pmx{\tau_1 P_{\cE_1}-\tau_2 P_{\cE_2}+ \He{P_{\cE_1} \bA}  & P_{\cE_1}   & 0\\P_{\cE_1}  & 0 &0\\
    0 & 0 & -\tau_1+\tau_2 }\nu\prec 0.
\]
By expanding the product, we have
\begin{multline*} 
\He{\widetilde \eta^\top  P_{\cE_1} \left(\bA_s\widetilde \eta +  \widetilde\psi\right)}  \\+\tau_1 (\widetilde\eta^\top P_{\cE_1}\widetilde\eta-1) + \tau_2 (1-\widetilde\eta^\top P_{\cE_2}\widetilde\eta)< 0.
\end{multline*}
Therefore,  by \eqref{eqn:nonlin_sys_closed_loop} we have 
\begin{equation*}
    \He{\dot {\widetilde \eta}^\top P_{\cE_1} \widetilde \eta} +\tau_1 (\eta^\top P_{\cE_1}\eta-1)- \tau_2 (\eta^\top P_{\cE_2}\eta-1) \le 0,
\end{equation*}
for all $\eta\in\cS$.
Since $\tau_1,\tau_2$ are positive, if $\cE_2\subseteq\cS$
the above inequality implies
\begin{equation}\label{eqn:lyap_set_decrease}
    \dot {\widetilde \eta}^\top P_{\cE_1}\widetilde \eta + \widetilde\eta^\top P_{\cE_1} \dot {\widetilde \eta}< 0,
\end{equation}
for all $\eta\in \widetilde\cE$. 
Note that the constraint $ P_{\cE_1}\succeq  P_{\cE_2}$ ensures $\widetilde\cE$ is not empty. Indeed, it implies 
\[\widetilde\eta^\top  P_{\cE_1}\widetilde\eta \ge \widetilde\eta^\top  P_{\cE_2}\widetilde\eta 
\]
and therefore 
$\cE_1\subseteq \cE_2$.
Consequently, $\widetilde\cE$ contains at least the boundary of $\cE_1$ and thus it is not empty.

Consider now the function $V(\eta)=\widetilde\eta^\top P_{\cE_1} \widetilde\eta$.  Given \eqref{eqn:lyap_set_decrease}, for all $\eta \in \widetilde\cE$ we have $\dot V(\eta) < 0$. Therefore, all trajectories starting in $\widetilde\cE$ converge to $\cE_1$, thus proving its attractiveness and forward-invariance.}

\subsection{Proof of Lemma~\ref{lem:orig2reshaped_abs}} \label{sec:proof_lem1}
If \eqref{eqn:D_full} holds, a Schur complement on the bottom-right entry provides
\[
\bD^\top \Omega +\Omega^\top \bD -\Omega^\top\Omega\succ 0.
\]
By adding and subtracting $\bD^\top \bD$, we obtain
\[
\bD^\top \bD\succ (\Omega-\bD)^\top(\Omega-\bD).
\]
Therefore, since $(\Omega-\bD)^\top(\Omega-\bD)\succeq 0$, $\bD^\top \bD\succ 0$. Let $\widetilde \eta \defeq \eta-\eta^\star$, $\widetilde \phi \defeq \phi(\bE\eta)-\phi(\bE\eta^\star)$ and $\widetilde \psi \defeq \psi(\bE\eta)-\psi(\bE\eta^\star) = \bD\widetilde \phi + \widetilde \bA\widetilde \eta $. Since $\bD^\top \bD$ is invertible, we have
\begin{align*}
    \bD^\top \widetilde \psi &= \bD^\top \bD \widetilde \phi + \bD^\top \widetilde \bA \widetilde \eta\\
      (\bD^\top \bD)^{-1}\bD^\top(\widetilde \psi - \widetilde \bA \widetilde \eta) &= \widetilde \phi \\
     \overbar \bD(\widetilde \psi - \widetilde \bA \widetilde \eta) &= \widetilde \phi 
\end{align*}
If \eqref{eqn:local_sect_eta} holds, the result is proven by substituting the above relation in \eqref{eqn:local_sect_eta} and collecting the terms in $\widetilde \eta$ and $\widetilde \psi$.

\subsection{Proof of Lemma~\ref{lem:diff2incr_abstraction}} \label{sec:proof_lem2}
Left and right multiplication of \eqref{eqn:diff_abstr} by $\widetilde v \defeq v_1-v_2$ and expansion of the product yields
\[
\widetilde v^\top R_v\widetilde v + 2 \widetilde v^\top S_v^\top \pderiv{\phi}{v}(v)\widetilde v +\widetilde v^\top \pderiv{\phi}{v}(v)^\top Q_v \pderiv{\phi}{v}(v)\widetilde v \le 0.
\]
Since $Q\succ 0$, by Schur complement we obtain
\begin{equation}\label{eqn:diff2incre_abs}
\pmx{\widetilde v^\top R_v\widetilde v + 2 \widetilde v^\top S_v^\top \pderiv{\phi}{v}(v)\widetilde v && \widetilde v^\top \pderiv{\phi}{v}(v)^\top\\  \pderiv{\phi}{v}(v)\widetilde v && -Q_v^{-1}} \preceq 0,
\end{equation}
for all $v,\widetilde v \in \RR^{n_v}$.
Consider now the function $F:[0,1]\to\RR^{n_v}$ defined as
\[
F(s) \defeq \phi(v_2+s(v_1-v_2)) = \phi(v_2+s\widetilde v).
\]
By the fundamental theorem of calculus, we have
\[
F(1)-F(0)=\phi(v_1)-\phi(v_2) = \int_0^1 \pderiv{\phi}{s}(v_2+s\widetilde v)\, ds \, \widetilde v.
\]
Since \eqref{eqn:diff2incre_abs} holds for all $v\in \RR^{n_v}$, selecting $v = v_2-s\widetilde v$ and integrating over $s$ yields
\[
\pmx{\widetilde v^\top R_v\widetilde v + 2 \widetilde v^\top S_v^\top (\phi(v_1)-\phi(v_2)) &&  (\phi(v_1)-\phi(v_2))^\top\\   (\phi(v_1)-\phi(v_2)) && -Q_v^{-1}} \preceq 0.
\]
The proof is concluded with by Schur complement on the bottom-right entry of the above matrix and by collecting the vectors $\col(\widetilde v, \phi(v_1)-\phi(v_2)) $ and its transpose.

\subsection{Proof of Theorem~\ref{thm:invariant_ellipsoid_local_sect}}\label{sec:proof_thm1}
By Proposition~\ref{prp:invariant_set}, the result is proven if \eqref{eqn:local_sector_v}, \eqref{eqn:D_full} and \eqref{eqn:local_sector_and_inclusion} imply the local abstraction \eqref{eqn:local_sect_eta_reshaped} for $\eta^\star=\col(x^\star, 0, 0)$ and $\cE_2\subseteq \cS$ with $\cS\defeq\{\eta\in\RR^{n_\eta}: \bE\eta\in[\underline v,\overbar v]\}$. Consider \eqref{eqn:local_sector_v} first and define $\widetilde v = v-v^\star = E(x-x^\star)=E\widetilde x$ and $\widetilde \phi = \phi(v)-\phi(v^\star)$. By \citep[Lemma 2]{de2023event}, inequality \eqref{eqn:local_sector_v} implies 
\begin{equation}
    \label{eqn:incremental_abs_v_Pi}
    \pmx{\widetilde v\\\widetilde \phi}^\top\pmx{X_v&Y_v^\top\\Y_v&Z_v}\pmx{\widetilde v\\\widetilde \phi}\le 0
\end{equation}
for all $v\in[\underline v,\overbar v]$ and all $X_v,Y_v,Z_v$ such that \eqref{eqn:local_sec_open_loop_polytopic} holds.
Since $\widetilde v=\bE\widetilde \eta$ and $\widetilde \phi = \phi(\bE\eta)-\phi(\bE\eta^\star)$, \eqref{eqn:incremental_abs_v_Pi} implies
\begin{equation}
    \label{eqn:incremental_abs_eta_Pi}
    \pmx{\widetilde \eta\\\widetilde \phi}^\top\pmx{\bE^\top X_v \bE &\bE^\top Y_v^\top\\Y_v\bE &Z_v}\pmx{\widetilde \eta\\\widetilde \phi}\le 0
\end{equation}
for all $\eta\in\cS$ and all $X_v,Y_v,Z_v$ such that \eqref{eqn:local_sec_open_loop_polytopic} holds. Therefore, by Lemma~\ref{lem:orig2reshaped_abs} with $X=\bE^\top X_v \bE$, $Y=Y_v\bE$ and $Z = Z_v$
we have 
\begin{equation}
    \label{eqn:local_sect_eta_reshaped_hat}
    \pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}^\top\pmx{\widehat X &\widehat Y^\top\\\widehat Y& \widehat Z}\pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}\le 0,
\end{equation}
for all $\eta\in\cS$. 
Consider now \eqref{eqn:local_sec_closed_loop_polytopic}. Right and left multiplication  of  by $\nu = \col(\dot{\widetilde \eta}, \widetilde \eta, \widetilde \psi)$ and its transpose 
yields
\[
\pmx{\widetilde \eta\\\widetilde \psi}^\top\pmx{\bX & \bY^\top\\\bY&\bZ}\pmx{\widetilde \eta\\\widetilde \psi}\le \pmx{\widetilde \eta\\\widetilde \psi}^\top\pmx{\widehat X & \widehat Y^\top\\\widehat Y&\widehat Z}\pmx{\widetilde \eta\\\widetilde \psi}\le 0,
\]
for all $\eta\in\cS$. 
Therefore, \eqref{eqn:incremental_abs_v_Pi} implies the local abstraction \eqref{eqn:local_sect_eta_reshaped} holds for all $\eta\in\cS$.

Consider now \eqref{eqn:ellipsoid_inclusion_polytopic}.
Note that $\eta\in\cS$ if $|\bE\eta-\bE\eta^\star|_i \le |\overline v -\underline v|_i$ for all vector entries $i=1,\dots,n_v$. A Schur complement on the bottom-right block of \eqref{eqn:ellipsoid_inclusion_polytopic} yields
\[
P_{\cE_2}\succeq \frac{1}{(\overbar v -\underline v)_i^\top  (\overbar v -\underline v)_i}\bE_i^\top \bE_i
\]
and in turn
\[
\widetilde \eta^\top P_{\cE_2} \widetilde \eta^\top\ge \frac{1}{(\overbar v -\underline v)^\top  (\overbar v -\underline v)} \widetilde \eta^\top\bE_i^\top \bE_i\widetilde \eta^\top.
\]
 Consequently, if $\eta\in\cE_2$, 
 \[
(\overbar v -\underline v)_i^\top  (\overbar v -\underline v)_i \ge \widetilde \eta^\top\bE_i^\top \bE_i\widetilde \eta^\top
\]
and thus $\eta\in\cS$. Therefore $\cE_2\subseteq \cS$ and this concludes the proof.
\section{Conclusions}
Future work will also explore discrete-time extensions, building on the recent studies \citep{zoboli2024k}.

\bibliography{Bibliography/first}

\end{document}

### 2024-12-12 Paper draft/backup_1.tex
Rename forlders

--- 2024-12-12 Paper draft/Paper_ifac.tex
@@
 
-		\author[First]{Samuele Zoboli} 		
-        \author[Second]{Andreu Cecilia}
-	\author[First]{Sophie Tarbouriech}
+		\author[First]{Andreu Cecilia} 		
+        \author[Second]{Samuele Zoboli}
+	\author[Second]{Sophie Tarbouriech}
 
-		\address[Second]{Universitat Polit\`ecnica de Catalunya, Avinguda Diagonal, 647, 08028 Barcelona, Spain, (e-mail: andreu.cecilia@upc.edu)}
-  \address[First]{LAAS-CNRS, Universit\'e de Toulouse, CNRS, Toulouse, (e-mail: samuele.zoboli@laas.fr, sophie.tarbouriech@laas.fr)}
+		\address[First]{Universitat Polit\`ecnica de Catalunya, Avinguda Diagonal, 647, 08028 Barcelona, Spain, (e-mail: andreu.cecilia@upc.edu)}
+  \address[Second]{LAAS-CNRS, Universit\'e de Toulouse, CNRS, Toulouse, (e-mail: samuele.zoboli@laas.fr, sophie.tarbouriech@laas.fr)}
 
@@
 \end{equation}
-holds for all $v\in\cV\subseteq \RR^{n_v}$ where $\cV\defeq\{v\in\RR^{n_v}:v=E x, x\in\cX\subseteq \RR^{n_x}\}$.
+holds for all $v\in\cV$ where $\cV\defeq\{v\in\RR^{n_v}:v=E x, x\in\cX\subseteq \RR^{n_x}\}$.
 We recall that such a class of nonlinearities includes common and interesting family of functions, such as (shifted) monotonic and differentially sector-bounded ones \citep{zoboli2024quadratic}.
-Finally, we assume that in the unforced case, namely when $u=0$, system \eqref{eqn:original_system} presents a set of fixed points $\cF_x^\star$.
+Finally, we assume that in the unforced case, namely when $u=0$, system \eqref{eqn:original_system} presents a set of fixed points $\Omega$.
 More formally, we assume the existence of a nonempty set of equilibria
 \begin{equation}\label{eqn:eq_points}
- \cF_x^\star := \{x\in\RR^{n_x} \ | \ Ax+D\phi(E x)= 0\}.
+ \Omega := \{x\in\RR^{n_x} \ | \ Ax+D\phi(E x)= 0\}.
 \end{equation}
-We highlight that $\cF^\star$ can include multiple elements, that is, system \eqref{eqn:original_system}  may present multiple equilibrium points for the unforced case $u=0$.
+We highlight that $\Omega$ can include multiple elements, that is, system \eqref{eqn:original_system}  may present multiple equilibrium points for the unforced case $u=0$.
 
-We aim to design a feedback control policy that makes (at least) one point $x^\star\in\cF_x^\star$ asymptotically stable.
-In other words, we want to make the set $\cF_x^\star$ attractive and forward-invariant.
+We aim to design a feedback control policy that makes (at least) one point $x^*\in\Omega$ asymptotically stable. In other words, we want to make the set $\Omega$ attractive and forward-invariant.
 
-One possibility is to reformulate the problem as a standard stabilization one, see e.g.
+One could imagine reframing the problem as a standard stabilization one, see e.g.
 \citep{isidori2017stabilization}.
-Differently put, the problem could be solved by selecting a specific point  $x^\star\in\cF_x^\star$ and designing a control law making $x^\star$ globally asymptotically stable.
+In other words, the problem could be solved by selecting a specific point  $x^*\in\Omega$ and designing a control law making $x^*$ globally asymptotically stable.
 However, we highlight that the considered problem relaxes the more common stabilization problem in three directions.
-First, the goal is not to stabilize the system at a specific point in $\cF_x^\star$, we only need the system to converge to $\cF_x^\star$.
+First, the goal is not to stabilize the system at a specific point in $\Omega$, we only need the system to converge to $\Omega$.
 Second, the proposed objective allows for the existence of multiple attractive equilibrium points for the closed-loop system.
-Third, the set of equilibrium points $\cF_x^\star$ may be unknown to the designer.
+Third, the set of equilibrium points $\Omega$ may be unknown to the designer.
 This third point is particularly relevant for feedback policies that include an integral action, e.g.
 \citep{Astolfi2017,GIACCAGLI2023111106}, which often require an unknown non-zero steady-state control signal to be generated by the integral component.
-Another possibility is to invoke set-stability approaches, e.g.
+Another possibility could be to invoice set-stability approaches, e.g.
 \citep{SHIRIAEV20001709}.
-However, we remark that $\cF_x^\star$ may be disconnected.
+However, we remark that $\Omega$ may be disconnected.
 Moreover, its elements are given by the (possibly unknown) equilibria of the unforced system dynamics.
@@
 
-Given the abovementioned constraints, in what follows, we present a solution that does not require knowledge of the open-loop equilibria and guarantees their stability via $2$-contraction \citep{Michael1995}.
+Given the abovementioned constraints, in what follows, we present a solution that does not require knowledge of the open-loop equilibria and guarantees their stability via the recent concept of $2$-contraction \citep{Michael1995}.
 Our approach is based on the recent findings in \citep{ cecilia2023generalized, zoboli2024dynamic} and inspired by the techniques of derivative control and washout filters.
@@
 We start by highlighting a necessary property controllers must satisfy to guarantee the open-loop equilibria are preserved by the closed loop.
-From \eqref{eqn:original_system}, it is clear that for the closed-loop system to stabilize at  $x^\star\in\cF_x^\star$ the control input must vanish at such point.
+From \eqref{eqn:original_system}, it is clear that for the closed-loop system to stabilize at  $x^\star\in\Omega$ the control input must vanish at such point.
 % dynamic output-feedback law $u=\alpha(y,\xi)$ with state $\xi\in\RR^{n_\xi}$ must satisfy 
 % \begin{equation}\label{eqn:noninvasive_cond}
-% \alpha(Cx^\star+G\phi(Ex^\star),\xi^\star) = 0, \qquad \forall x^\star \in \cF^\star,
+% \alpha(Cx^\star+G\phi(Ex^\star),\xi^\star) = 0, \qquad \forall x^\star \in \Omega,
 % \end{equation}
@@
 \[
-\cF^\star\defeq\{x\in\RR^{n_x}:f(x,0)=0\}.
+\Omega_x\defeq\{x\in\RR^{n_x}:f(x,0)=0\}.
 \]
@@
 \begin{equation}\label{eqn:noninvasive_cond}
-\alpha(h(x^\star),\xi^\star) = 0, \qquad \forall x^\star \in \cF^\star,
+\alpha(h(x^\star),\xi^\star) = 0, \qquad \forall x^\star \in \Omega_x,
 \end{equation}
-where $\xi^\star$ is the controller equilibrium associated to $x^\star$.
+where $\xi^\star$ is the controller equilibrium state associated to $x^\star$.
 \end{definition}
@@
 \end{equation}
-At any equilibrium point $(x^\star,z^\star)$ the $z$-dynamics impose $u=0$, and thus $x^\star \in \cF_x^\star$.
+At any equilibrium point $(x^\star,z^\star)$ the $z$-dynamics impose $u=0$, and thus $x^\star \in \Omega$.
+
+We remark that such an approach does not allow any of the system's outputs to be regulated.
+Indeed, from output regulation theory, the number of regulated outputs must be smaller than or equal to the number of inputs, see e.g.
+\cite{isidori2003robust}.
+By adding integrator dynamics on all input signals, such a regulation limit is reached.
+We also stress that the addition of input-integrator dynamics may cause the loss of controllability properties. This is shown by the following simple example.
+Consider the planar linear time invariant system
+\begin{equation}
+    \label{eqn:ctrlb_example_sys}
+    \begin{array}{rl}
+     \dot x_1&= x_1+u  \\
+     \dot x_2 &= x_1 
+\end{array}
+\end{equation}
+with $x=\col(x_1,x_2)\in\RR^2$ and $u\in\RR$.
+It can be easily verified that \eqref{eqn:ctrlb_example_sys} is controllable.
+However, by adding the input-integrator dynamics, the resulting extended system
+\[
+    \begin{array}{rl}
+     \dot x_1&= x_1+u  \\
+     \dot x_2 &= x_1 \\
+     \dot z &= u,
+\end{array}
+\]
+has a simply stable uncontrollable mode. Luckily, the $2$-contraction framework proposed in this paper only requires a generalized notion of stabilizability properties \citep{cecilia2023generalized}.
 
@@
 \end{equation}
-holds for all $\eta\in\cS\subseteq \RR^{n_\eta}$ with $\cS\defeq\{\eta\RR^{n_\eta}: \bE\eta\in\cV\}$, see \citep[Section 2.5]{zoboli2024dynamic}. 
+holds for all $\eta\in\RR^{n_\eta}$, see \citep[Section 2.5]{zoboli2024dynamic}. 
 The following result proposes a design for noninvasive linear dynamic output-feedback controllers based on matrix inequalities.
@@
       A_e&\defeq \pmx{A&0\\0&0},\qquad  B_e \defeq \pmx{B\\\Id_{n_u}},\qquad C_e \defeq \pmx{C&0\\0&\Id_{n_u}},\qquad 
-    D_e \defeq  \pmx{D\\0}, \qquad G_e \defeq  \pmx{G\\0}, \quad 
-        \Phi \defeq \pmx{\Pi_2 & \Id_{n_\xi}\\ \Id_{n_\xi} & \Pi_1},
+    D_e \defeq  \pmx{D\\0}, \qquad G_e \defeq  \pmx{G\\0},
      \\&\\
-        \widehat A &\defeq\pmx{A_e\Pi_2+B_e \widehat L && A_e + B_e  \widehat K C_e \\ \widehat M && \Pi_1 A_e+ \widehat N  C_e }, \quad  
-        \widehat D \defeq \pmx{D_e +B_e \widehat K G_e\\ \Pi_1 D_e  + \widehat N G_e}, 
-        \quad  \widehat R \defeq \pmx{ \widehat R_{11} &  \widehat R_{12}^\top\\
-        \star & \widehat R_{22}}, \quad \widehat S^\top \defeq \pmx{\widehat S_1^\top \\ \widehat S_2^\top},
-        \\&\\
-        &
-        \pmx{\widehat R_{11} - \overbar R& \widehat R_{12}^\top - \alpha \Pi_2 R^\top & \widehat S_{1}^\top-\alpha \Pi_2 S^\top\\
-         \star & \widehat R_{22}-\alpha R & \widehat S_{2}^\top -\alpha S^\top \\
-         \star&\star& \widehat Q-\alpha Q} +
-         \He{\pmx{\Upsilon_1\\\Upsilon_2\\\Upsilon_3}\pmx{\pderiv{\phi}{x}(v)& 0 & -\Id_{n_x}}}\preceq 0 , \qquad \forall v \in \cV,
+        \widehat A &\defeq\pmx{A_eY+B_e \widehat L && A_e + B_e  \widehat K C_e \\ \widehat M && X A_e+ \widehat N  C_e }, \qquad \qquad  
+        \widehat D \defeq \pmx{D_e +B_e \widehat K G_e\\ X D_e  + \widehat N G_e}, \qquad \qquad
+        \Phi \defeq \pmx{Y & \Id_{n_\xi}\\ \Id_{n_\xi} & X},
         \\&\\
            & \pmx{
-   \He{A_e \Pi_2+B_e \widehat L} -2\beta_1 \Pi_2 - \widehat R_{11} && A_e+B_e \widehat K C_e  + \widehat M^\top-2\beta_1 \Id_{n_\xi} -  \widehat R_{12}^\top && D_e +B_e \widehat K G_e -  \widehat S_1^\top\\
-   \star && \He{ \Pi_1 A_e +  \widehat N  C_e } -2\beta_1 \Pi_1   -\widehat R_{22}&& \Pi_1 D_e  + \widehat N G_e -\widehat S_2\\
-   \star &&\star&& - \widehat Q
+   \He{A_e Y+B_e \widehat L} -2\beta_1 Y - \widehat R_{11} && A_e+B_e \widehat K C_e  + \widehat M^\top-2\beta_1 \Id_{n_\xi} -  \alpha YR^\top && D_e +B_e \widehat K G_e -  \alpha YS^\top\\
+   \star && \He{ X A_e +  \widehat N  C_e } -2\beta_1 X   -\alpha R&& X D_e  + \widehat N G_e -\alpha S\\
+   \star &&\star&& - \alpha Q
     }\prec 0,\\&\\
@@
     \star  && -\He{\Gamma_2 \Phi}  & &\Gamma_2 \widehat D -\Phi\Gamma_3^\top \\
-    \star && \star& &\He{\Gamma_3 \widehat D} - \widehat Q
+    \star && \star& &\He{\Gamma_3 \widehat D} - \alpha  Q
     }\prec 0  .\\&\\
@@
      Consider  \eqref{eqn:nonlin_sys_closed_loop} with \eqref{eqn:nonlin_sys_closed_loop_matrices} %\eqref{eqn:closed_loop}
-and let \eqref{eqn:generic_abstraction_ctrl} hold for all  $x\in\cF$  with given matrices $R, S, Q$.
-Select $Z\preceq 0$ such that $Z\preceq R$ and assume there exist $ \overbar R_{11}=\overbar R^\top,$ $ \widehat R_{11}=\widehat R_{11}^\top,$ $ \widehat R_{12},$ $ \widehat R_{22}=\widehat R_{22}^\top,$ $ \widehat S_1,$ $ \widehat S_2,$ $ \widehat Q=\widehat Q^\top,$ scalars $\alpha>0,\beta_0\in\RR$, $\beta_1<0$,  symmetric matrices $\Pi_1=\Pi_1^\top, \Pi_2=\Pi_2^\top \in \RR^{n_\xi\times n_\xi}$, matrices $\widehat K \in \RR^{n_u\times n_y}$, $\widehat L\in \RR^{n_u\times n_\xi}$, $\widehat M\in \RR^{n_\xi\times n_\xi}$, $\widehat N \in \RR^{n_\xi\times n_\xi}$,  matrices $\Gamma_1, \Gamma_2, \Gamma_3$  and $\Sigma=\Sigma^\top\succ 0$  such that
+and let \eqref{eqn:generic_abstraction_ctrl} hold for all  $x\in\cX$  with given matrices $R, S, Q$.
+Select $Z\preceq 0$ such that $Z\preceq R$ and assume there exist $ \widehat R_{11}=\widehat R_{11}^\top,$   scalars $\alpha>0,\beta_0\in\RR$, $\beta_1<0$,  symmetric matrices $X=X^\top, Y=Y^\top \in \RR^{n_\xi\times n_\xi}$, matrices $\widehat K \in \RR^{n_u\times n_y}$, $\widehat L\in \RR^{n_u\times n_\xi}$, $\widehat M\in \RR^{n_\xi\times n_\xi}$, $\widehat N \in \RR^{n_\xi\times n_\xi}$,  matrices $\Gamma_1, \Gamma_2, \Gamma_3$  and $\Sigma=\Sigma^\top\succ 0$  such that
 \begin{enumerate}[i)]
-        \item The matrix $ R_Z \defeq \smallpmat{\overbar R && \alpha \Pi_2Z\\ \star && \alpha Z}$ is negative semidefinite;
-         \item Inequality \eqref{eqn:2_contraction_BMI}\footnotemark[1] holds with $\beta_0+\beta_1<0$;
+        \item The matrix $\widehat R \defeq \smallpmat{\widehat R_{11} && \alpha YZ\\ \star && \alpha Z}$ is negative semidefinite;
+         \item Inequality \eqref{eqn:2_contraction_BMI}\footnotemark[1] holds with $\widehat S = \smallpmat{  \alpha SY & \alpha S}$ and $\beta_0+\beta_1<0$;
         \item $\cine(\Gamma_2)=(1, 0, 2n_\xi-1)$.
@@
      Then, 
-     there exist square invertible matrices $U,V\in \RR^{n_\xi\times n_\xi}$ such that $UV^\top = \Id_{n_\xi}- \Pi_1\Pi_2$
+     there exist square invertible matrices $U,V\in \RR^{n_\xi\times n_\xi}$ such that $UV^\top = \Id_{n_\xi}- XY$
     and a scalar $\Delta^\star>0$ such that the control matrices 
@@
           \pmx{K & H}&= \widehat K,  \\
-         \pmx{N& F}&= U^{-1} (\widehat N - \Pi_1 B \widehat K) ,\\
-           L & = (\widehat L - \widehat K C \Pi_2) V^{-\top},\\
-         M&=U^{-1}\!\pmx{\Id_{n_\xi}\\-\Pi_1}^{\!\top}\!\!\pmx{\widehat M & \widehat N C\\
-         B \widehat L & -A+ B\widehat  KC}\pmx{\Id_{n_\xi}\\-\Pi_2}V^{-\top}\!\!\!\!\!\!
+         \pmx{N& F}&= U^{-1} (\widehat N - X B \widehat K) ,\\
+           L & = (\widehat L - \widehat K C Y) V^{-\top},\\
+         M&=U^{-1}\!\pmx{\Id_{n_\xi}\\-X}^{\!\top}\!\!\pmx{\widehat M & \widehat N C\\
+         B \widehat L & -A+ B\widehat  KC}\pmx{\Id_{n_\xi}\\-Y}V^{-\top}\!\!\!\!\!\!
     \end{array}
@@
 \end{equation} 
-evolving in a forward-invariant bounded compact set $\cK\subseteq\cS$ with  $\delta(\eta)=\col(\delta_x(\eta),0,0)$,  $\delta_x(\eta)\in \cC^1$, $|\pderiv{\delta_x}{\eta}(\eta)|<\Delta^\star$  for all $\eta \in \cK$,
-converge to an equilibrium $\eta^\star\in \cF^\star_\eta$ with
+evolving in a forward-invariant bounded compact set $\cK\subsetneq \RR^{2 n_\xi}$ with  $\delta(\eta)=\col(\delta_x(\eta),0,0)$,  $\delta_x(\eta)\in \cC^1$, $|\pderiv{\delta_x}{\eta}(\eta)|<\Delta^\star$  for all $\eta \in \cK$,
+converge to an equilibrium $\eta^\star\in \Omega_\eta$ with
     \begin{equation}
         \label{eqn:closed_loop_eq_set}
-        \cF^\star_\eta \defeq \{\eta\in\cK: Ax+D\phi(E x) + \delta_x(x)= 0\}.
+        \Omega_\eta \defeq \{\eta\in\cK: Ax+D\phi(E x) + \delta_x(x)= 0\}.
     \end{equation}
@@
 A detailed discussion on methods to impose such constraints in the optimization problem is presented in \citep[Section 3]{zoboli2024dynamic}.
-Secondly, it requires closed-loop trajectories to evolve in a forward invariant bounded compact set $\cK$.
+Secondly, it requires closed-loop trajectories to evolve in forward invariant bounded compact sets.
 This necessity arises from the fact that the proposed result is grounded on the notion of $2$-contraction, which guarantees convergence to nonunique equilibria if the system evolves in forward invariant bounded compact sets \citep[Theorem 2.5]{Michael1995}, \citep[Lemma 6]{cecilia2023generalized}.
 Therefore, we now present an LMI-based approach allowing the \textit{a-posteriori} verification of such a requirement.
+% \begin{proposition}\label{prp:invariant_set}
+%     Consider \eqref{eqn:nonlin_sys_closed_loop} and assume there exist matrices $\bR=\bR^\top$, $\bS$ and $\bQ$ such that 
+%     \begin{equation}\label{eqn:nonincremental_abstraction}
+%     \pmx{\eta\\\phi(\bE\eta)}^\top\pmx{\bR&\bS^\top\\\bS&\bQ}\pmx{\eta\\\phi(\bE\eta)}\preceq 0
+% \end{equation}
+% for all $\eta\in\RR^{n_\eta}$. If there exist scalars $\tau\in\RR$, $\lambda>0$, $\gamma>0$ and a matrix $P_\cE\succ 0$ such that 
+% \begin{equation}
+%     \label{eqn:ellypsoid_LMI}
+%     \pmx{\tau P_\cE+ \bA^\top P_\cE + P_\cE \bA - \gamma \bR & P_\cE \bD -\gamma \bS^\top & 0\\\bD^\top P_\cE -\gamma \bS & -\gamma \bQ&0\\
+%     0 & 0 & -\tau \lambda}\preceq 0,
+% \end{equation}
+% then the ellipsoidal set $\cE=\{\eta\in\RR^{n_\eta}:\eta^\top P_\cE \eta \le \lambda\}$ is forward invariant for \eqref{eqn:nonlin_sys_closed_loop}.
+% \end{proposition}
 
+{
 \begin{figure*}
+    \begin{subequations}\label{eqn:invariance_LMIs}
         \begin{equation}
-    \label{eqn:ellipsoid_LMI}
-    \pmx{\tau_1 P_{\cE_1}- \tau_2 P_{\cE_2} + \bA_s^\top P_{\cE_1} + P_{\cE_1} \bA_s - \bX && P_{\cE_1}  -\bY^\top && 0\\\star && - \bZ&&0\\
-    \star && \star  && -\tau_1 + \tau_2 }\prec 0.
+        \label{eqn:ellypsoid_inclusion}
+\lambda_3 \ge \lambda_1, \qquad  P_{\cE_1}- \frac{\lambda_3}{\lambda_2} P_{\cE_2}\succeq 0,
     \end{equation}
+\begin{equation}
+    \label{eqn:ellypsoid_LMI}
+    \pmx{\tau_1 P_{\cE_1}- \tau_2 P_{\cE_2}+ \bA^\top P_{\cE_1} + P_{\cE_1} \bA - \gamma \bR && P_{\cE_1} \bD -\gamma \bS^\top && 0\\\star && -\gamma \bQ&&0\\
+    \star && \star  && -\tau_1 \lambda_1+ \tau_2 \lambda_2}\prec 0.
+\end{equation}
+    \end{subequations}
     \vspace{.5em}
@@
 \begin{proposition}\label{prp:invariant_set}
-    Consider \eqref{eqn:nonlin_sys_closed_loop} and let $\bA = \bA_s +\widetilde \bA$ with $\bA_s$ Hurwitz.
-Suppose there exist matrices $\bX=\bX^\top$, $\bY$, $\bZ=\bZ^\top$, a set $\cS\subseteq\RR^{n_\eta}$ and a vector $\eta^\star\in\cS$ such that the local sector condition
-    \begin{equation}
-        \label{eqn:local_sect_eta_reshaped}
-        \pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}^\top\pmx{\bX &\bY^\top\\\bY&\bZ}\pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}\le 0
+    Consider \eqref{eqn:nonlin_sys_closed_loop} and assume there exist matrices $\bR=\bR^\top$, $\bS$ and $\bQ$ such that 
+    \begin{equation}\label{eqn:nonincremental_abstraction}
+    \pmx{\eta\\\phi(\bE\eta)}^\top\pmx{\bR&\bS^\top\\\bS&\bQ}\pmx{\eta\\\phi(\bE\eta)}\preceq 0
 \end{equation}
-    holds for all $\eta\in\cS$, with $\psi(v)\defeq \bD \phi(v)+\widetilde \bA\eta$. 
-    If there exist positive scalars $\tau_1$, $\tau_2$, two matrices $P_{\cE_1}\succeq P_{\cE_2}\succ 0$ and a matrix $\Omega\in\RR^{n_\eta \times n_u}$ such that \eqref{eqn:ellipsoid_LMI}\footnotemark[2] holds and such that the ellipsoidal set $\cE_2\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_2} (\eta-\eta^\star) \le 1\} \subseteq\cS$, then $\cE_1\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_1} (\eta-\eta^\star) \le 1\}$ is  attractive and forward invariant for \eqref{eqn:nonlin_sys_closed_loop} with region of attraction including the  non-empty set $\widetilde\cE\defeq \cE_2\setminus\cE_1$, for all $\eta\in \cS$.
+for all $\eta\in\RR^{n_\eta}$.
+If there exist positive scalars $\tau_1$, $\tau_2$, $\lambda_1$, $\lambda_2$, $\gamma>0$ and two matrices $P_{\cE_1}\succ 0, P_{\cE_2}\succ 0$ such that \eqref{eqn:invariance_LMIs}\footnotemark[2] holds, the ellipsoidal set $\cE_1\defeq\{\eta\in\RR^{n_\eta}:\eta^\top P_{\cE_1} \eta \le \lambda_1\}$ is attractive and forward invariant for \eqref{eqn:nonlin_sys_closed_loop}, with region of attraction including the set $\cS\defeq\{\eta\in\RR^{n_\eta}: \eta^\top P_{\cE_1}\eta\ge\lambda_1 \land \eta^\top P_{\cE_2}\eta\le\lambda_2\}$.
 \end{proposition}
+}
 The proof of Proposition~\ref{prp:invariant_set} is presented in Section~\ref{sec:proof_prp2}.
@@
 
-\begin{remark}
-    Note that the loop transformation $\bA=\bA_s+\widetilde \bA$,  $\psi(v)\defeq \bD \phi(v)+\widetilde \bA\eta$ is a necessary step for the satisfaction of \eqref{eqn:ellipsoid_LMI}.
-Indeed, after the removal of the quadratic abstraction entry $\bX$, the top-left entry of \eqref{eqn:ellipsoid_LMI} requires
-    \[
-    \tau_1 P_{\cE_1}- \tau_2 P_{\cE_2} + \bA_s^\top P_{\cE_1} + P_{\cE_1} \bA_s \prec 0.
-    \]
-    Since $P_{\cE_1}\succeq P_{\cE_2}$ this implies
-    \[
-    (\tau_1-\tau_2) P_{\cE_1} + \bA_s^\top P_{\cE_1} + P_{\cE_1} \bA_s \prec 0,
-    \]
-    which requires $\bA_s$ Hurwitz, given that $\tau_1>\tau_2$ from the bottom-right entry of \eqref{eqn:ellipsoid_LMI}.
-Moreover, since the pair $(\bA, \bD)$ may not be controllable, the more classical loop transformation involving the addition and subtraction of a virtual stabilizing input $u_s = \bK\eta$ and the definition of a reshaped nonlinearity $\psi(v)=\phi(v) - \bK \eta$ may not be applicable.
-Thus, $\psi(v)$ must include $\bD$ in its definition.
-\end{remark}
-%
-While useful, the results of Proposition~\ref{prp:invariant_set} present two main drawbacks.
-First, they require the knowledge of a nondifferential quadratic abstraction \eqref{eqn:local_sect_eta_reshaped} for the \emph{reshaped closed loop}.
-Second, they do not guarantee the outer ellipsoid $\cE_2\subseteq\cS$.
-We first address the issues emerging from the loop transformation by proposing a result connecting quadratic abstractions on the original system to ones on the reshaped loop.
-\begin{lemma}\label{lem:orig2reshaped_abs}
-    Consider \eqref{eqn:nonlin_sys_closed_loop} and suppose there exist matrices $X=X$, $Y$, $Z=Z$, a set $\cS\subseteq\RR^{n_\eta}$ and a vector $\eta^\star\in\cS$ such that the local sector condition
-    \begin{equation}
-        \label{eqn:local_sect_eta}
-        \pmx{\eta - \eta^\star\\ \phi(\bE\eta)-\phi(\bE\eta^\star)}^\top\pmx{X &Y^\top\\Y&Z}\pmx{\eta - \eta^\star\\ \phi(\bE\eta)-\phi(\bE\eta^\star)}\le 0
-    \end{equation}
-    holds for all $\eta\in\cS$. If there exists a matrix $\Omega\in\RR^{n_\eta\times n_v}$ such that
-        \begin{equation}
-    \label{eqn:D_full}
-    \pmx{\bD^\top \Omega + \Omega^\top \bD && \Omega^\top\\
-    \star && \Id_{n_u}}\succ 0,
-\end{equation}
-then $\bD^\top\bD\succ 0$ and \eqref{eqn:local_sect_eta_reshaped} holds with
-\begin{equation}
-    \label{eqn:original2reshaped_abs_matrices}
-    \begin{array}{rl}
-        \overbar \bD &= (\bD^\top \bD)^{-1}\bD^\top,\\
-         \bX &= X- Y^\top \overbar \bD\widetilde \bA - \widetilde \bA^\top  \overbar \bD^\top Y - \widetilde \bA^\top \overbar \bD^\top Z \overbar \bD\widetilde \bA,\\
-         \bY &= \overbar \bD Y + \overbar \bD^\top Z \overbar \bD\widetilde \bA,\\
-         \bZ &= \overbar \bD^\top Z \overbar \bD.
-    \end{array}
-\end{equation}
-\end{lemma}
-The proof of Lemma~\ref{lem:orig2reshaped_abs} is presented in Section~\ref{sec:proof_lem1}.
+Proposition~\ref{prp:invariant_set} relies on a classical quadratic abstraction \eqref{eqn:nonincremental_abstraction}, while Proposition~\ref{prp:2_contr_BMI} exploits a differential one, i.e.
+\eqref{eqn:generic_abstraction_ctrl}.
+In the following technical lemma, we show how classical quadratic abstractions can be recovered from differential ones.
 
-Lemma~\ref{lem:orig2reshaped_abs} allows translating quadratic abstractions for the original closed loop to abstractions for the transformed one.
-We highlight that constraint \eqref{eqn:D_full} can be dropped by assuming $\bD$ has full column rank.
-
-While representing an important step in improving the applicability of
-Proposition~\ref{prp:invariant_set}, Lemma~\ref{lem:orig2reshaped_abs} requires knowledge of matrices $X,Y,Z$ in \eqref{eqn:local_sect_eta} and it does not guarantee the outer ellipsoid $\cE_2\subseteq\cS$.
-One possible workaround to both issues is to exploit the knowledge of the differential abstraction \eqref{eqn:generic_abstraction} to derive a global incremental one.
-This is formalized in the following lemma.
 \begin{lemma}
@@
     \end{equation}
-    for all $v\in \RR^{n_v}$, then 
+    for all $v\in \RR^{n_v}$. Then, 
     \begin{equation}\label{eqn:incr_abstr}
@@
 \end{lemma}
-The proof of Lemma~\ref{lem:diff2incr_abstraction} is presented in Section~\ref{sec:proof_lem2}.
+The proof of Lemma~\ref{lem:diff2incr_abstraction} is presented in Section~\ref{sec:proof_lem1}.
 
-Given the relation $v=Ex=\bE\eta$, Lemma~\ref{lem:diff2incr_abstraction} shows that one can select $X = \gamma\bR = \gamma \bE^\top R_v \bE$, $Y=\gamma \bS = \gamma \bE S_v$ and $Z=\gamma \bQ = \gamma Q_v$ in \eqref{eqn:ellipsoid_LMI} for any $\gamma>0$, see for instance \citep[Section 2.5]{zoboli2024dynamic}.
-Indeed, since \eqref{eqn:incr_abstr} holds for all $v_1,v_2$, it holds also for $v_2=\bE\eta^\star$ and obtain $\cS=\RR^{n_\eta}$ in Proposition~\ref{prp:invariant_set}, ensuring $\cE_2\subseteq\cS$.
-Then, Lemma~\ref{lem:orig2reshaped_abs} allows mapping the obtained abstraction to \eqref{eqn:local_sect_eta_reshaped}.
-Given \eqref{eqn:original2reshaped_abs_matrices}, the requirement of $Q_v\succ0$ does not restrict the set of solutions to the LMI, as the central block of \eqref{eqn:ellipsoid_LMI} imposes the same constraint in the form $\bZ\succ 0$.
-
-Unfortunately, the exploitation of Lemma~\ref{lem:diff2incr_abstraction} is very conservative and still presents drawbacks.
-Importantly, \eqref{eqn:incr_abstr} holds for any pair $v_1,v_2$.
-As such, the resulting abstraction represents an incremental sector condition rather than a local one as per \eqref{eqn:local_sect_eta_reshaped}.
-Therefore, while holding also for the ellipsoid center $\eta^\star$, it is not specific to such a point.
-This may lead to increased conservatives and numerical problems since the resulting ellipsoids are not anchored to a specific point, i.e., they could be interpreted as incremental ones.
-
-Given the above discussion, we argue that estimating the matrices $\bX,\bY,\bZ$ given a desired open-loop center $x^\star$ is a more reliable solution.
-This approach only requires minor knowledge of local sector bounds for the nonlinearity $\phi$.
-Moreover, it allows including constraints guaranteeing $\cE_2$ is included in the known local sector.
-We formalize the proposed technique in the following theorem.
-
-\begin{figure*}
-    \begin{subequations}
-    \label{eqn:local_sector_and_inclusion}
-    \begin{equation}
-    \label{eqn:original2reshaped_abs_matrices}
-    \begin{array}{rl}
-         \widehat X &= \bE^\top X_v\bE - \bE^\top Y_v^\top \overbar \bD\widetilde \bA - \widetilde \bA^\top  \overbar \bD^\top Y_v \bE - \widetilde \bA^\top \overbar \bD^\top Z_v \overbar \bD\widetilde \bA,\\
-         \widehat Y &= \overbar \bD Y + \overbar \bD^\top Z \overbar \bD\widetilde \bA,\\
-         \widehat Z &= \overbar \bD^\top Z \overbar \bD.
-    \end{array}
-\end{equation}
-\begin{align}
-    \label{eqn:local_sec_open_loop_polytopic}
-    &\pmx{\Id_{n_v}\\\Lambda}^\top\pmx{X_v&Y_v^\top\\Y_v&Z_v}\pmx{\Id_{n_v}\\\Lambda} \preceq 0,  \qquad \forall \Lambda: \Theta_1\preceq \Lambda\preceq\Theta_2\\
-    \notag\\
-\label{eqn:local_sec_closed_loop_polytopic}
-    &\pmx{ 0 && 0 && 0\\
-     0 && \bX - \widehat X && \bY^\top - \widehat Y^\top \\
-     \0 && \bY - \widehat Y && \bZ- \widehat Z }+\He{\pmx{\Upsilon_1 \\ \Upsilon_2\\
-    \Upsilon_3 
-    }\pmx{\Id_{n_\eta} & -\bA & \Id_{n_v}}}\preceq 0\\\notag\\
-    \label{eqn:ellipsoid_inclusion_polytopic}
-    &\pmx{P_{\cE_2} &\bE_i^\top\\ \bE_i & (\overbar v -\underline v)_i^2}\succeq 0, \qquad i=1, \dots, n_v.
-\end{align}
-\hrule
-\end{subequations}
-\end{figure*}
-\begin{theorem}\label{thm:invariant_ellipsoid_local_sect}
-Let $\Theta_1,\Theta_2\in\RR^{n_v\times n_v}$ be given diagonal matrices such that $\Theta_1\preceq\Theta_2$.
-Assume there exist vectors $\underline v, \overbar v, x^\star$ such that $v^\star = Ex^\star \in[\underline v, \overbar v]$ and
-\begin{equation}
-    \label{eqn:local_sector_v}
-   (\phi(v)-\phi(v^\star) -\Theta_1(v-v^\star))_i
-   (\phi(v)-\phi(v^\star) -\Theta_2(v-v^\star))_i\le0
-\end{equation}
-for all $i=1,\dots,n_v$ and for all $v\in[\underline v, \overbar v]$, where $(\cdot)_i$ represents the $i^{\text{th}}$ element of the argument vector or the $i^{\text{th}}$ row of the argument matrix.
-If there exist positive scalars $\tau_1$, $\tau_2$, a matrix $\Omega\in\RR^{n_\eta\times n_v}$, two matrices $P_{\cE_1}\succeq P_{\cE_2}\succ 0$, matrices $\Upsilon_i$ with $i=,\dots, 8$  and matrices $X_v=X_v^\top$, $Y_v$, $ Z_v= Z_v^\top$, $\bX=\bX^\top$, $\bY$, $ \bZ= \bZ^\top$ such that \eqref{eqn:ellipsoid_LMI} , \eqref{eqn:D_full}
-and \eqref{eqn:local_sector_and_inclusion} hold wit $\overbar \bD$ as in \eqref{eqn:original2reshaped_abs_matrices}, then $\cE_1\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_1} (\eta-\eta^\star) \le 1\}$ is  attractive and forward invariant for \eqref{eqn:nonlin_sys_closed_loop} with region of attraction including the  non-empty set $\widetilde\cE\defeq \cE_2\setminus\cE_1$, with
-${\color{red}\eta^\star=\col(x^\star, 0, 0)}$ and
-$\cE_2\defeq\{\eta\in\RR^{n_\eta}:(\eta-\eta^\star)^\top P_{\cE_2} (\eta-\eta^\star) \le 1\}$.
-\end{theorem}
-The proof of Theorem~\ref{thm:invariant_ellipsoid_local_sect} is presented in Section~\ref{sec:proof_thm1}.
-
-Theorem~\ref{thm:invariant_ellipsoid_local_sect} considers the abstraction matrices $\bX,\bY,\bZ$, $X_v, Y_v, Z_v$ as problem variables.
-To reduce the size of the optimization problem, one could select $\bX=\widehat X$,  $ \bY =  \widehat Y$  and $\bZ= \widehat Z$, thus removing \eqref{eqn:local_sec_closed_loop_polytopic}.
-However, the inclusion of \eqref{eqn:local_sec_closed_loop_polytopic} leaves more degrees of freedom to the optimization problem, which can exploit entries in the directions of the integrator and controller states.
-Under the selection mentioned previously, these entries would be set to $0$ by the structure of $\bE$.
-We also highlight that \eqref{eqn:local_sec_open_loop_polytopic} is typically hard to solve.
-However, if we constrain the diagonal elements of $Z_v$ to be nonpositive, it is sufficient to check the condition for all
-matrices $\Lambda$ corresponding to the finitely many vertices
-of the matrix interval $[\Theta_1,\Theta_2]$ \citep{ebihara2021stability}.
-This can be easily done b, e.g., decomposing $Z_v\defeq Z_d + \widetilde Z^\top + \widetilde Z$ with diagonal $Z_d\preceq 0$ and lower triangular $\widetilde Z$.
+Lemma~\ref{lem:diff2incr_abstraction} shows that, if $\phi(0)=0$, then the selection $v_2 = 0$  and positive definiteness of $\bQ$ allow recovering \eqref{eqn:nonincremental_abstraction} from \eqref{eqn:generic_abstraction_ctrl} by selecting $v_1 = \bE\eta$ and recalling that \eqref{eqn:generic_abstraction_ctrl} can be obtained from inequalities of the form \eqref{eqn:diff_abstr}, see \citep[Section 2.5]{zoboli2024dynamic}.
 
 \begin{remark}
-    Since the differential quadratic abstraction \eqref{eqn:generic_abstraction} holds for all $v\in\cV$ and given the degree of freedom arising from letting matrices $X_v,Y_v,Z_v$ as decision variables, one can select $\underline v, \overbar v$ and compute $\Theta_1,\Theta_2$ for Theorem~\ref{thm:invariant_ellipsoid_local_sect} such that the resulting polytope is included in $\cV$.
-This choice connects the regions where 2-contractivity holds and the invariant ellipsoid is looked for.
+    The results of Proposition~\ref{prp:invariant_set} guarantee the existence of an invariant ellipsoid centered in the origin.
+However, if the open-loop equilibria are known, such an ellipsoid can be shifted by requiring \eqref{eqn:nonincremental_abstraction} to hold for the error variables $\eta-\eta^\star$ and $\phi(\bE\eta)-\phi(\bE\eta^\star)$.
+Thanks to the incremental nature of \eqref{eqn:incr_abstr}, Lemma~\ref{lem:diff2incr_abstraction} can still be used to derive the necessary quadratic abstraction from the differential formulation \eqref{eqn:generic_abstraction_ctrl}.
 \end{remark}
@@
 Recently, the authors in \citep{ANGELI2024152} applied $2$-contraction theory for designing state-derivative controllers.
-$2$-contraction enabled the development of non-local design conditions for multistable systems.
+$2$-contraction enabled the development of semi-global design conditions for multistable systems.
 A similar framework to the one proposed in \citep{ANGELI2024152} can be recovered from \eqref{eqn:original_system} by selecting $C=A, G=D$ and the static control policy $u=Ky$.
@@
 Thus, we do not require any measurement of the state or output derivative, and we can operate directly with the system output.
-Nevertheless, we remark that our approach is not free from tradeoffs.
-Indeed, the addition of input integrators does not allow any of the system's outputs to be regulated.
-From output regulation theory, the number of regulated outputs must be smaller than or equal to the number of inputs, see e.g.
-\citep{isidori2003robust}.
-By adding integrator dynamics on all input signals, such a regulation limit is reached.
-We also stress that the addition of input-integrator dynamics may cause the loss of controllability properties. This is shown by the following simple example.
-Consider the planar linear time invariant system
-\begin{equation}
-    \label{eqn:ctrlb_example_sys}
-    \begin{array}{rl}
-     \dot x_1&= x_1+u  \\
-     \dot x_2 &= x_1 
-\end{array}
-\end{equation}
-with $x=\col(x_1,x_2)\in\RR^2$ and $u\in\RR$.
-It can be easily verified that \eqref{eqn:ctrlb_example_sys} is controllable.
-However, by adding the input-integrator dynamics, the resulting extended system
-\[
-    \begin{array}{rl}
-     \dot x_1&= x_1+u  \\
-     \dot x_2 &= x_1 \\
-     \dot z &= u,
-\end{array}
-\]
-has a simply stable uncontrollable mode. However, the $2$-contraction framework proposed in this paper only requires a generalized notion of stabilizability properties \citep{cecilia2023generalized}.
-We also highlight that, different from previous results with washout filter, the combination of $2$-contraction \citep{cecilia2023generalized} with the design conditions provided in \citep{zoboli2024dynamic} provides for the first time LMI design conditions for the proposed stabilization problem which are:
+However, we highlight that, different from previous results with washout filter, the combination of $2$-contraction \citep{cecilia2023generalized} with the design conditions provided in \citep{zoboli2024dynamic} provides, for the first time, LMI design conditions for the proposed stabilization problem which are:
 \begin{itemize}
@@
 $$
-\cF^\star = \{x\in\RR^2 \ | \ \cos(x_1+\theta)=0, \ x_2 = 0\}.
+\Omega = \{x\in\RR^2 \ | \ \cos(x_1+\theta)=0, \ x_2 = 0\}.
 $$
 In the unforced case $u=0$, the considered system can be understood as a frictionless pendulum with $x_1$ being the pendulum angle shifted by $\theta$ and $x_2$ is the angular velocity.
-Therefore, the set $\cF^\star$ is not attractive and for any initial condition outside $\cF^\star$ the system converges to a non-trivial limit cycle.
-With this in mind, the objective is to make the set $\cF^\star$ attractive and invariant.
+Therefore, the set $\Omega$ is not attractive and for any initial condition outside $\Omega$ the system converges to a non-trivial limit cycle.
+With this in mind, the objective is to make the set $\Omega$ attractive and invariant.
 
@@
 Note that the extended matrices $A_e, B_e, C_e, D_e, G_e$ in \eqref{eqn:2_contraction_BMI} describe the $(x,z)$-dynamics in \eqref{eqn:extended_dyn}.
-Therefore, by \citep[Theorem 2]{zoboli2024dynamic} adapted with the result on known quadratic abstraction of \citep[Corollary 2]{zoboli2024dynamic}, the matrix inequalities in \eqref{eqn:2_contraction_BMI} paired with items i)-iii) of Proposition~\ref{prp:2_contr_BMI} guarantee $\cine(\Sigma_1)=(1,0,2n_\eta-1)$ and
+Therefore, by \citep[Theorem 2]{zoboli2024dynamic} adapted with the result on known quadratic abstraction of \citep[Proposition 3]{zoboli2024dynamic}, the matrix inequalities in \eqref{eqn:2_contraction_BMI} paired with items i)-iii) of Proposition~\ref{prp:2_contr_BMI} guarantee $\cine(\Sigma_1)=(1,0,2n_\eta-1)$ and
 \begin{equation}
@@
 \subsection{Proof of Proposition~\ref{prp:invariant_set}} \label{sec:proof_prp2}
-% If \eqref{eqn:nonincremental_abstraction} holds, \eqref{eqn:ellipsoid_LMI} and $\gamma>0$ imply
+% If \eqref{eqn:nonincremental_abstraction} holds, \eqref{eqn:ellypsoid_LMI} and $\gamma>0$ imply
 % \[
@@
 \label{sec:proof_prp2}
-Let $\widetilde \eta \defeq \eta-\eta^\star$, $\widetilde \phi \defeq \phi(\bE\eta)-\phi(\bE\eta^\star)$ and $\widetilde \psi \defeq \psi(\bE\eta)-\psi(\bE\eta^\star) = \bD\widetilde \phi + \widetilde \bA\widetilde \eta $.
-According to the dynamics \eqref{eqn:nonlin_sys_closed_loop} and the relation $\bA=\bA_s+\widetilde\bA$, we have
+If \eqref{eqn:nonincremental_abstraction} holds, \eqref{eqn:ellypsoid_LMI} and $\gamma>0$ imply
 \[
-\dot{\widetilde\eta} = \bA_s\widetilde \eta +\widetilde \psi.
+    \pmx{\tau_1 P_{\cE_1}-\tau_2 P_{\cE_2}+ \He{P_{\cE_1} \bA}  & P_{\cE_1} \bB  & 0\\\bB^\top P_{\cE_1}  & 0 &0\\
+    0 & 0 & -\tau_1 \lambda_1+\tau_2 \lambda_2}\prec 0.
 \]
-If \eqref{eqn:local_sect_eta_reshaped} holds for all $\eta\in\cS$, right and left multiplication of \eqref{eqn:ellipsoid_LMI} by $\nu=\col(\widetilde \eta,\widetilde\psi)$ and its transpose implies
+Therefore, left and right multiplication by the vector $\nu=\col(\eta,\phi(\bE\eta),1)$ and its transpose paired with the dynamics \eqref{eqn:nonlin_sys_closed_loop}  yields
 \[
-    \nu^\top\pmx{\tau_1 P_{\cE_1}-\tau_2 P_{\cE_2}+ \He{P_{\cE_1} \bA}  & P_{\cE_1}   & 0\\P_{\cE_1}  & 0 &0\\
-    0 & 0 & -\tau_1+\tau_2 }\nu\prec 0.
+\dot \eta^\top P_{\cE_1} \eta + \eta^\top P_{\cE_1} \dot \eta +\tau_1 (\eta^\top P_{\cE_1}\eta-\lambda_1)- \tau_2 (\eta^\top P_{\cE_2}\eta-\lambda_2)< 0.
 \]
-By expanding the product, we have
-\begin{multline*} 
-\He{\widetilde \eta^\top  P_{\cE_1} \left(\bA_s\widetilde \eta +  \widetilde\psi\right)}  \\+\tau_1 (\widetilde\eta^\top P_{\cE_1}\widetilde\eta-1) + \tau_2 (1-\widetilde\eta^\top P_{\cE_2}\widetilde\eta)< 0.
-\end{multline*}
-Therefore,  by \eqref{eqn:nonlin_sys_closed_loop} we have 
-\begin{equation*}
-    \He{\dot {\widetilde \eta}^\top P_{\cE_1} \widetilde \eta} +\tau_1 (\eta^\top P_{\cE_1}\eta-1)- \tau_2 (\eta^\top P_{\cE_2}\eta-1) \le 0,
-\end{equation*}
-for all $\eta\in\cS$.
-Since $\tau_1,\tau_2$ are positive, if $\cE_2\subseteq\cS$
+Since $\tau_1,\tau_2$ are positive, if $\eta\in \cS$ 
 the above inequality implies
 \begin{equation}\label{eqn:lyap_set_decrease}
-    \dot {\widetilde \eta}^\top P_{\cE_1}\widetilde \eta + \widetilde\eta^\top P_{\cE_1} \dot {\widetilde \eta}< 0,
+    \dot \eta^\top P_{\cE_1} \eta + \eta^\top P_{\cE_1} \dot \eta< 0.
 \end{equation}
-for all $\eta\in \widetilde\cE$. 
-Note that the constraint $ P_{\cE_1}\succeq  P_{\cE_2}$ ensures $\widetilde\cE$ is not empty. Indeed, it implies 
-\[\widetilde\eta^\top  P_{\cE_1}\widetilde\eta \ge \widetilde\eta^\top  P_{\cE_2}\widetilde\eta 
+Note that \eqref{eqn:ellypsoid_inclusion} ensures $\cS$ is not empty. Indeed, \eqref{eqn:ellypsoid_inclusion}  implies 
+\[\frac{1}{\lambda_3}\eta^\top  P_{\cE_1}\eta \ge \frac{1}{\lambda_2}\eta^\top  P_{\cE_2}\eta 
 \]
 and therefore 
-$\cE_1\subseteq \cE_2$.
-Consequently, $\widetilde\cE$ contains at least the boundary of $\cE_1$ and thus it is not empty.
+\[
+  \eta^\top P_{\cE_1}\eta\le\lambda_3\implies \eta^\top P_{\cE_2}\eta\le\lambda_2.
+\]
+Since $\lambda_3\ge\lambda_1$, the set
+\[
+\overbar \cS\defeq\{\eta\in\RR^{n_\eta}: \lambda_1\le \eta^\top P_{\cE_1}\eta\le\lambda_3\},
+\]
+is not empty, as it contains at least the boundary of the ellipsoidal set $\cE_1$, namely
+\[
+\partial \cE_1\defeq\{\eta\in\RR^{n_\eta}: \eta^\top P_{\cE_1}\eta=\lambda_1 \}.
+\]
+Consequently, since $\overbar \cS\subseteq\cS$, $\cS$ is not empty.
 
-Consider now the function $V(\eta)=\widetilde\eta^\top P_{\cE_1} \widetilde\eta$.
-Given \eqref{eqn:lyap_set_decrease}, for all $\eta \in \widetilde\cE$ we have $\dot V(\eta) < 0$.
-Therefore, all trajectories starting in $\widetilde\cE$ converge to $\cE_1$, thus proving its attractiveness and forward-invariance.}
+Consider now the function $V(\eta)=\eta^\top P_{\cE_1} \eta$.
+Given \eqref{eqn:lyap_set_decrease}, for all $\eta \in \cS$ we have $\dot V(\eta) < 0$.
+Therefore, all trajectories starting in $\cS$ converge to $\cE_1$, thus proving its attractiveness and forward-invariance.}
 
-\subsection{Proof of Lemma~\ref{lem:orig2reshaped_abs}} \label{sec:proof_lem1}
-If \eqref{eqn:D_full} holds, a Schur complement on the bottom-right entry provides
-\[
-\bD^\top \Omega +\Omega^\top \bD -\Omega^\top\Omega\succ 0.
-\]
-By adding and subtracting $\bD^\top \bD$, we obtain
-\[
-\bD^\top \bD\succ (\Omega-\bD)^\top(\Omega-\bD).
-\]
-Therefore, since $(\Omega-\bD)^\top(\Omega-\bD)\succeq 0$, $\bD^\top \bD\succ 0$.
-Let $\widetilde \eta \defeq \eta-\eta^\star$, $\widetilde \phi \defeq \phi(\bE\eta)-\phi(\bE\eta^\star)$ and $\widetilde \psi \defeq \psi(\bE\eta)-\psi(\bE\eta^\star) = \bD\widetilde \phi + \widetilde \bA\widetilde \eta $.
-Since $\bD^\top \bD$ is invertible, we have
-\begin{align*}
-    \bD^\top \widetilde \psi &= \bD^\top \bD \widetilde \phi + \bD^\top \widetilde \bA \widetilde \eta\\
-      (\bD^\top \bD)^{-1}\bD^\top(\widetilde \psi - \widetilde \bA \widetilde \eta) &= \widetilde \phi \\
-     \overbar \bD(\widetilde \psi - \widetilde \bA \widetilde \eta) &= \widetilde \phi 
-\end{align*}
-If \eqref{eqn:local_sect_eta} holds, the result is proven by substituting the above relation in \eqref{eqn:local_sect_eta} and collecting the terms in $\widetilde \eta$ and $\widetilde \psi$.
-
-\subsection{Proof of Lemma~\ref{lem:diff2incr_abstraction}} \label{sec:proof_lem2}
+\subsection{Proof of Lemma~\ref{lem:diff2incr_abstraction}} \label{sec:proof_lem1}
 Left and right multiplication of \eqref{eqn:diff_abstr} by $\widetilde v \defeq v_1-v_2$ and expansion of the product yields
@@
 
-\subsection{Proof of Theorem~\ref{thm:invariant_ellipsoid_local_sect}}\label{sec:proof_thm1}
-By Proposition~\ref{prp:invariant_set}, the result is proven if \eqref{eqn:local_sector_v}, \eqref{eqn:D_full} and \eqref{eqn:local_sector_and_inclusion} imply the local abstraction \eqref{eqn:local_sect_eta_reshaped} for $\eta^\star=\col(x^\star, 0, 0)$ and $\cE_2\subseteq \cS$ with $\cS\defeq\{\eta\in\RR^{n_\eta}: \bE\eta\in[\underline v,\overbar v]\}$.
-Consider \eqref{eqn:local_sector_v} first and define $\widetilde v = v-v^\star = E(x-x^\star)=E\widetilde x$ and $\widetilde \phi = \phi(v)-\phi(v^\star)$.
-By \citep[Lemma 2]{de2023event}, inequality \eqref{eqn:local_sector_v} implies
-\begin{equation}
-    \label{eqn:incremental_abs_v_Pi}
-    \pmx{\widetilde v\\\widetilde \phi}^\top\pmx{X_v&Y_v^\top\\Y_v&Z_v}\pmx{\widetilde v\\\widetilde \phi}\le 0
-\end{equation}
-for all $v\in[\underline v,\overbar v]$ and all $X_v,Y_v,Z_v$ such that \eqref{eqn:local_sec_open_loop_polytopic} holds.
-Since $\widetilde v=\bE\widetilde \eta$ and $\widetilde \phi = \phi(\bE\eta)-\phi(\bE\eta^\star)$, \eqref{eqn:incremental_abs_v_Pi} implies
-\begin{equation}
-    \label{eqn:incremental_abs_eta_Pi}
-    \pmx{\widetilde \eta\\\widetilde \phi}^\top\pmx{\bE^\top X_v \bE &\bE^\top Y_v^\top\\Y_v\bE &Z_v}\pmx{\widetilde \eta\\\widetilde \phi}\le 0
-\end{equation}
-for all $\eta\in\cS$ and all $X_v,Y_v,Z_v$ such that \eqref{eqn:local_sec_open_loop_polytopic} holds.
-Therefore, by Lemma~\ref{lem:orig2reshaped_abs} with $X=\bE^\top X_v \bE$, $Y=Y_v\bE$ and $Z = Z_v$
-we have 
-\begin{equation}
-    \label{eqn:local_sect_eta_reshaped_hat}
-    \pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}^\top\pmx{\widehat X &\widehat Y^\top\\\widehat Y& \widehat Z}\pmx{\eta - \eta^\star\\ \psi(\bE\eta)-\psi(\bE\eta^\star)}\le 0,
-\end{equation}
-for all $\eta\in\cS$. 
-Consider now \eqref{eqn:local_sec_closed_loop_polytopic}. Right and left multiplication  of  by $\nu = \col(\dot{\widetilde \eta}, \widetilde \eta, \widetilde \psi)$ and its transpose 
-yields
-\[
-\pmx{\widetilde \eta\\\widetilde \psi}^\top\pmx{\bX & \bY^\top\\\bY&\bZ}\pmx{\widetilde \eta\\\widetilde \psi}\le \pmx{\widetilde \eta\\\widetilde \psi}^\top\pmx{\widehat X & \widehat Y^\top\\\widehat Y&\widehat Z}\pmx{\widetilde \eta\\\widetilde \psi}\le 0,
-\]
-for all $\eta\in\cS$. 
-Therefore, \eqref{eqn:incremental_abs_v_Pi} implies the local abstraction \eqref{eqn:local_sect_eta_reshaped} holds for all $\eta\in\cS$.
-
-Consider now \eqref{eqn:ellipsoid_inclusion_polytopic}.
-Note that $\eta\in\cS$ if $|\bE\eta-\bE\eta^\star|_i \le |\overline v -\underline v|_i$ for all vector entries $i=1,\dots,n_v$.
-A Schur complement on the bottom-right block of \eqref{eqn:ellipsoid_inclusion_polytopic} yields
-\[
-P_{\cE_2}\succeq \frac{1}{(\overbar v -\underline v)_i^\top  (\overbar v -\underline v)_i}\bE_i^\top \bE_i
-\]
-and in turn
-\[
-\widetilde \eta^\top P_{\cE_2} \widetilde \eta^\top\ge \frac{1}{(\overbar v -\underline v)^\top  (\overbar v -\underline v)} \widetilde \eta^\top\bE_i^\top \bE_i\widetilde \eta^\top.
-\]
- Consequently, if $\eta\in\cE_2$, 
- \[
-(\overbar v -\underline v)_i^\top  (\overbar v -\underline v)_i \ge \widetilde \eta^\top\bE_i^\top \bE_i\widetilde \eta^\top
-\]
-and thus $\eta\in\cS$. Therefore $\cE_2\subseteq \cS$ and this concludes the proof.
 \section{Conclusions}

### 2024-12-12 Paper draft/draft.tex
Rename forlders

@@
 For a partitioned matrix, the symbol $\star$ stands for symmetric blocks.
-$\1_n = \bmx{1 & \dots & 1}-\tr \in \NN^{n \times 1}$
+$\1_n = \bmx{1 & \dots & 1}\tr \in \NN^{n \times 1}$.
+We identify with subscript $i$ the $i$-th element of a vector or the $i$-th row of a matrix.
 
@@
 $$
-with $\hat{u}$ the lmit of the input signal, $x \in \RR^{n_x}$ the state vector, $\bar{u} \in \RR^{n_u}$ the raw output of the controller, $\Phi$ the representation of the non-linearity of the system \textcolor{red}{Please help in adding details about $\Phi$ since it's NOLCOS} and $d \in \RR^{n_d}$ a constant disturbance input.
+with $\hat{u}$ the limit of the input signal, $x \in \RR^{n_x}$ the state vector, $\bar{u} \in \RR^{n_u}$ the raw output of the controller, $d \in \RR^{n_d}$ a constant disturbance input  and $\Phi \in \RR^{n_q}$ the representation of the non-linearity of the system.
+\begin{assumption}\label{ass:nonlin}
+We assume that $\Phi$ satisfies some local quadratic abstraction for some $\underline{\mu}, \bar{\mu} > 0 \in \RR^{n_q}, \bar{S}, \bar{Q}, \bar{W} \in \RR^{n_q \times n_q}$
+\begin{equation}\label{eqn:nonlin-set}
+\forall v \in \bar{S} = \left\{ v \in  \RR^{n_q} : - \underline{\mu}_i \leq v_i \leq \bar{\mu}_i, i \in \pmx{1 & \dots & n_q} \right\} 
+\end{equation}
+\begin{equation}\label{eqn:nonlin-inclusion}
+\bmx{x \\ \Phi(x)}\tr \bmx{\bar{S} & \bar{Q} \\ \star & \bar{W}} \bmx{x \\ \Phi(x)} \geq 0
+\end{equation}
+\end{assumption}\textcolor{red}{Please help in adding details about $\Phi$ since it's NOLCOS}.
 
@@
 The controller $\pi$ is defined by:
-\begin{equation}\label{eq:nn-equations}
+\begin{equation}\label{eqn:nn-equations}
   \begin{aligned}
-  \omega^{0}(k) &= x(k) \\
-  \nu^{i}(k) &= W^{i} \omega^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
+  \hat{\omega}^{0}(k) &= x(k) \\
+  \nu^{i}(k) &= W^{i} \hat{\omega}^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
   \omega^{i}(k) &= \phi_i(\nu^i(k))\\
@@
 \end{equation}
-with $\nu^i(k) \in \RR^{n_{\phi_i}}$ the input to the $i$-th activation function $\phi_i: \RR^{n_{\phi_i}} \to \RR^{n_{\phi_i}}$, $\omega^i(k) \in \RR^{n_{\phi_i}}$ the output.
+with $\nu^i(k) \in \RR^{n_{\phi_i}}$ the input to the $i$-th activation function $\phi^i: \RR^{n_{\phi_i}} \to \RR^{n_{\phi_i}}$, $\omega^i(k), \hat{\omega}^i(k) \in \RR^{n_{\phi_i}}$ the current and last forwarded output respectively.
 The weights $W^i \in \RR^{n_{\phi_i} \times n_{\phi_{i-1}}}$ and biases $b^i \in \RR^{n_{\phi_i}}$ define the affine operation of each layer.
-The saturation function application is to be intended element-wise and will be denoted as $\phi_i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$ and $\varphi: \RR \to \RR$ is the scalar activation function assumed to be symmetric and identical for every neuron.
+The saturation function application is to be intended element-wise and will be denoted as $\phi^i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$ and $\varphi: \RR \to \RR$ is the scalar activation function assumed to be symmetric and identical for every neuron.
 Adopting the same notation as in \cite{css-paper} we define the controller policy in a condensed form.
@@
 \begin{equation*}
-  \nu = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi = \bmx{\phi_1(\nu^1) \\ \vdots \\ \phi_l(\nu^l)} 
+  \nu = \bmx{\nu^1 \\ \vdots \\ \nu^l}, \omega = \bmx{\omega^1 \\ \vdots \\ \omega^l}, \phi = \bmx{\phi^1(\nu^1) \\ \vdots \\ \phi^l(\nu^l)} 
 \end{equation*}
 with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega = \phi(\nu)$.
-Finally, conditions \eqref{eq:nn-equations} can be rewritten as:
+Finally, conditions \eqref{eqn:nn-equations} can be rewritten as:
 \begin{equation*}
-  \bmx{u(k)\\ \nu(k)} = N \bmx{x(k) \\ \omega(k) \\ 1} 
+  \bmx{u(k)\\ \nu(k)} = \overbar N \bmx{x(k) \\ \omega(k) \\ 1} 
 \end{equation*}
 where
-\begin{equation}\label{eq:first-N}
-  \bmx{
+\begin{equation}\label{eqn:first-N}
+  \overbar N = \bmx{
     \begin{array}{c | c c c c | c}
@@
 The dynamics will be rewritten as:
-\begin{equation}\label{eq:system-dynamics}
+\begin{equation}\label{eqn:system-dynamics}
   x^{+} = A x + B u + C \Phi(E x) + D d
 \end{equation}
-with $u \in \RR^{n_u}$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \eqref{eq:first-N} as:
-\begin{equation}\label{eq:last-N}
+with $u \in \RR^{n_u}$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \eqref{eqn:first-N} as:
+\begin{equation}\label{eqn:last-N}
   N = \bmx{
@@
     \begin{aligned}
-         R &= (\Id - N_{\nu \omega})^{-1}\\
-         R_{\omega} &= N_{u x} + N_{u \omega} R N_{\nu x}\\
-         R_b &= N_{u \omega} R N_{\nu b} + N_{u b}
+         &R = (\Id - N_{\nu \omega})^{-1}\\
+         &R_{\omega} = N_{u x} + N_{u \omega} R N_{\nu x}\\
+         &R_b = N_{u \omega} R N_{\nu b} + N_{u b}
     \end{aligned}
@@
 
-With $N_{\nu \omega}$ always invertible due to its lower triangular structure.
+With $R$ always invertible due to $N_{\nu \omega}$ always invertible due to its lower triangular structure.
 We find the equilibrium of the state by the use of a numerical solver that minimizes the following expression since it is no longer explicit due to the non-linearity $\Phi$.
@@
 All the non-linearity of the system and the controller are selected as saturation functions.
-With respect to \citep[Lemma 3]{css-extended}, we declare the local sector conditions that hold for this function:
+With respect to \citep[Lemma 3]{css-extended}, we declare the local sector conditions that hold for every state included in the following set:
 
-\emph{For $i \in \left\{ 1, \dots, l \right\}$ given a matrix $G^i \in \RR^{n_{\phi_i} \times n_x}$, if $x(k)$ belongs to the set}
-$$
-S^i = \left\{ x(k) \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}_* \leq G^i (x(k) - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
-$$
+\begin{remark}\label{rem:sec-set}
+\emph{For $i \in \left\{ 1, \dots, l \right\}$ given a matrix $G^i \in \RR^{n_{\phi_i} \times n_x}$, if $x$ belongs to the set}
+\begin{equation}\label{eqn:inclusion-set}
+S = \left\{ x \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}^i_* \leq G^i (x - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
+\end{equation}
 
@@
 \begin{equation}\label{eqn:first_sec_cond}
-  \left[ \nu^i(k) - \omega^i(k) \right]\tr T^i \left[ G^i (x(k) - x_*) - (\omega^i(k) - \omega^i_*) \right] \leq 0
-\end{equation}
+  \left[ \nu^i - \omega^i \right]\tr T^i \left[ G^i (x - x_*) - (\omega^i - \omega^i_*) \right] \leq 0
+\end{equation}\end{remark}
 Where $\bar{\nu}^i$ is the saturation level of the $i$-th layer.
 
-\textcolor{red}{Proof needed?}
-
-In order to have a lighter notation the dead-zone vector  will be defined as $\psi = \nu - \omega$.
+In order to have a lighter notation the dead-zone vector  will be defined as $\psi^i = \nu^i - \omega^i$ and $\hat{\psi}^i = \nu^i - \hat{\omega}^i$.
 Furthermore, all following discussions will take into account incremental variables with respect to their equilibrium, denoted with a tilde, i.e.
@@
 Considering \eqref{eqn:equilibrium} we have $\psi_* = \nu_* - \omega_* = 0$ hence $\tilde{\psi} = \psi$.
-The sector condition \eqref{eqn:first_sec_cond} can be rewritten with no loss of generality as:
+We define $\xi^i = \bmx{\tilde{x}\tr & \tilde{\psi}_i\tr & \tilde{\nu}_i\tr}\tr$ and $\hat{\xi}^i = \bmx{\tilde{x}\tr & \hat{\tilde{\psi}}_i\tr & \tilde{\nu}_i\tr}\tr$, then the overall sector conditions \eqref{eqn:first_sec_cond} can be rewritten with no loss of generality $\forall i \in \pmx{1 & \dots & l}$ as:
 \begin{equation}\label{eqn:sec_cond}
-  \tilde{\psi}\tr T^i \left[ G^i \tilde{x} + \tilde{\psi} - \tilde{\nu} \right] \leq 0
+\begin{aligned}
+    &({\tilde{\psi}^i})\tr T^i \bmx{G^i \tilde{x} + \tilde{\psi}^i - \tilde{\nu}^i} = {\xi^i}\tr \bmx{0 & 0 & 0\\
+    T^i G^i & T^1 & -T^1\\
+    0 & 0 & 0} \xi^i = \\
+    &= {\xi^i}\tr \Omega^i \xi^i \leq 0
+\end{aligned}
 \end{equation}
+
+\begin{remark}\label{rem:sec-cond}
+Note how ensuring ${\xi^i}\tr \Omega^i \xi^i < 0$ guarantees that the sector condition \eqref{eqn:sec_cond} holds.
+\end{remark}
+
 We can express now the new inter-dependency between the variables of interest:
@@
 An ETM works by triggering an event that will update the current layer and propagate its value across the network. 
-The triggering function will take as inputs the incremental variables of the state of the system
-$\tilde{x}$, the current state of the layer $\tilde{\nu}$ and the dead-zone vector $\tilde{\psi}$. The ETM was defined as:
+In the precedent work \cite{css-extended} the triggering function is defined on the basis of the saturation sector conditions in the form \eqref{eqn:sec_cond}. The ETM was defined as:
 \begin{equation}
-  \omega^{i}(k) = \begin{cases}
-    \phi_i(\nu^i(k)) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > 0\\
-    \omega^{i}(k-1) & \text{otherwise}
+  \hat{\omega}^{i}(k) = \begin{cases}
+    \phi^i(\nu^i(k)) & \text{if } {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i} > 0\\
+    \hat{\omega}^{i}(k-1) & \text{otherwise}
   \end{cases}
 \end{equation}
-In the precedent work \cite{css-extended} the function $f$ is defined on the basis of the saturation sector conditions in the following quadratic form with respect to \eqref{eqn:sec_cond}:
-\begin{equation}\label{eqn:base_omega}
-  f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) =
-  \bmx{\tilde{x}\\ \tilde{\psi} \\ \tilde{\nu}}\tr
-  \Omega^i
-  \bmx{\tilde{x}\\ \tilde{\psi} \\ \tilde{\nu}} > 0
-\end{equation}
-With $$
-\Omega^i =  \bmx{
-    0 & 0 & 0\\
-    T^i G^i & T^i & -T^i\\
-    0 & 0 & 0
-  } 
-$$
 
@@
 \begin{equation}\label{eqn:dynamic_trig}
-  \omega^{i}(k) = \begin{cases}
-    \phi_i(\nu^i(k)) & \text{if } f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta_i\\
-    \omega^{i}(k-1) & \text{otherwise}
+  \hat{\omega}^{i}(k) = \begin{cases}
+    \phi^i(\nu^i(k)) & \text{if } {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i} > \rho^i \eta_i\\
+    \hat{\omega}^{i}(k-1) & \text{otherwise}
   \end{cases};
@@
 \begin{equation}\label{eqn:eta_dynamics}
-  \eta^+_i = \rho^i \eta_i - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu})\  \forall i \in \left( 1, \dots, l \right) 
+  {\eta^i}^+ = \rho^i \eta^i - {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i}\quad  \forall i \in \left( 1, \dots, l \right) 
 \end{equation}
-
-\begin{remark} \emph{Similarly as in \citep[Lemma 3]{data-driven} we will demonstrate how $\eta^i_k$ being the dynamic threshold of layer $i$ at time instant $k$ is non-negative for any $\rho^i \in [0, 1)$, $\eta^i_0 \geq\ \forall k \in \NN$} \end{remark}
-
-\begin{proof} We consider the dynamics \eqref{eqn:eta_dynamics} and the ETM triggering rule \eqref{eqn:dynamic_trig} we prove the remark by recursion:\\
-\emph{Initial condition}: Assuming $\eta^i_0 \geq 0$ we may have two possible outcomes:
-if $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) < \rho^i \eta^i_0$ we have that 
-\begin{equation}
-  f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) < \rho^i \eta^i_0 \to \rho^i \eta^i_0 - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) = \eta^i_1 \geq 0
-\end{equation}
-
-If instead $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i_0$ then the event is triggered and the values of the current layer state $(\tilde{\psi}, \tilde{\nu})$ are updated to $(\bar{\psi}, \bar{\nu})$, leading again to the first case scenario where $f^i(\tilde{x}, \bar{\psi}, \bar{\nu}) < \rho^i \eta^i_0 \to \eta^i_1 \geq 0$.\\
-\emph{Recursion step}: Assuming $\eta^i_k \geq 0$ similarly as the initial condition we may have:
-$$
-f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) < \rho^i \eta^i_k \to \rho^i \eta^i_k - f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) = \eta^i_{k+1} \geq 0
-$$
-If $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i_k$ once again the event is triggered, we update $(\tilde{\psi}, \tilde{\nu}) \to (\bar{\psi}, \bar{\nu})$, the triggering function switches sign $f^i(\tilde{x}, \bar{\psi}, \bar{\nu}) < \rho^i \eta^i_k \to \eta^i_{k+1} \geq 0$.\\
-Hence, $\eta^i_k \geq 0$ implies $\eta^i_{k+1} \geq 0$ concluding the recursion proof. \end{proof}
-\textcolor{red}{Is the recursion really necessary? Can't I just do the proof for any k?}
 
@@
    \bR &= \text{diag} \pmx{\rho^1 & \dots & \rho^l }\\
-   \mathbf{\Psi} &= \text{diag} \pmx{ f^1 & \dots & f^l }\\ 
-   \boldsymbol{\eta} &= \bmx{ {\eta^1}\tr & \dots & {\eta^l}\tr }\tr   
+   \mathbf{\Psi} &= \bmx{ {\hat{\xi^1}}\tr \Omega^1 \hat{\xi^1} & \dots & {\hat{\xi^l}}\tr \Omega^l \hat{\xi^l}}\tr\\
+   \boldsymbol{\eta} &= \bmx{ \eta^1 & \dots & \eta^l}\tr   
 \end{aligned}
@@
 \begin{aligned}
-    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \mathbf{\Psi} \1_l\\
-    \mathbf{\Psi} \1_l &> \bR \boldsymbol{\eta}
+    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \mathbf{\Psi}\\
+    \mathbf{\Psi} &> \bR \boldsymbol{\eta}
 \end{aligned}
@@
 \subsection{Finsler Lemma application}
-After the implementation of the dynamic threshold a further improvement of the triggering condition has been taken in consideration.
-More precisely by denoting $\xi^i = (\tilde{x}, \tilde{\psi}^i, \tilde{\nu}^i)$ we rewrite the even triggering condition $f^i(\tilde{x}, \tilde{\psi}, \tilde{\nu}) > \rho^i \eta^i$ by taking in consideration \eqref{eqn:base_omega} as
-\begin{equation}
-    {\xi^i}\tr \Omega^i \xi^i > \rho^i \eta^i
-\end{equation}
+After the implementation of the dynamic threshold a further improvement of the triggering condition has been taken in consideration. By recalling the triggering condition:
+$$
+\hat{{\xi^i}}\tr \Omega^i \hat{\xi^i} > \rho^i \eta^i
+$$
 Intuitively the lower the left-hand term of the inequality the lower will be the number of triggered events.
@@
 
-\begin{lemma}: \emph{Considering the triggering policy \eqref{eqn:dynamic_trig}, $R^i$ such that $\bmx{\tilde{x} & \tilde{\psi}^i & \tilde{\nu}^i}\tr\!
-= \!R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr\!
-\to {\xi^i}\tr = R^i \xi\tr$ and assuming \eqref{eqn:incremental} holds $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists \!
-X^i\!
-\in\!
-\RR^{nx + 2n_{\phi_i} \times nx + 2n_{\phi_i}}, $ $N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ such that:
+\begin{lemma}\label{lem:finsler}: \emph{Considering the triggering policy \eqref{eqn:dynamic_trig}, $R^i$ such that ${\xi^i}\tr = \!R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr\!
+= R^i \xi\tr$, with reference to the structure in \eqref{eqn:sec_cond} and assuming \eqref{eqn:incremental} holds for some $\underline{\xi}, \bar{\xi} \in \RR^{n_x + 2n_{\phi}}$ such that $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists X^i \in\!
+\RR^{(nx + 2n_{\phi_i}) \times (nx + 2n_{\phi_i})},$ $N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ a diagonal matrix $T^i > 0 \in \RR^{n_{\phi_i} \times n_{\phi_i}}, G^i \in \RR^{n_{\phi_i} \times n_x}$ such that:
 \begin{equation}\label{eqn:finsler}
@@
 $$
-    {\xi^i}\tr He(X^i - \Omega^i) \xi^i \leq 0
+    {\xi^i}\tr \He{X^i - \Omega^i} \xi^i \leq 0
 $$
@@
 Assuming we have the $X^i$ matrices $i \in \left\{1, \dots, l \right\}$ we rewrite the triggering conditions for each layer:
-\begin{equation}
-  \omega^{i}(k) = \begin{cases}
-    \phi_i(\nu^i(k)) & \text{if } {\xi^i}\tr X^i \xi > \rho^i \eta^i(k)\\
-    \omega^{i}(k-1) & \text{otherwise}
+\begin{equation}\label{eqn:etm-trigger}
+  \hat{\omega}^{i}(k) = \begin{cases}
+    \phi^i(\nu^i(k)) & \text{if } {\xi^i}\tr X^i \xi > \rho^i \eta^i(k)\\
+    \hat{\omega}^{i}(k-1) & \text{otherwise}
   \end{cases}
 \end{equation}
-By denoting $\bX = \text{diag} \pmx{{R^1}\tr X^1 R^1 & \dots & {R^l}\tr X^l R^l}$, the ETM triggering condition and dynamic thresholds' dynamics can be rewritten in an element-wise fashion as:
+\begin{remark}\label{rem:eta-positive} \emph{Similarly as in \citep[Lemma 3]{data-driven} it is possible to demonstrate how $\eta^i$ being the dynamic threshold of layer $i$ is always non-negative for any $\rho^i \in [0, 1)$, $\eta^i_0 \geq\ 0$} \end{remark}
+With respect to \eqref{eqn:etm-trigger} and \eqref{eqn:eta_dynamics} we have two possible occurrences:
+\begin{itemize}
+    \item $\hat{{\xi^i}}\tr X^i \hat{\xi^i} \leq \rho^i \eta^i$: it's trivial to verify that ${\eta^i}^+ = \rho^i \eta^i - \hat{{\xi^i}}\tr X^i \hat{\xi^i} \geq 0 $
+    \item $\hat{\xi^i}\tr X^i \hat{\xi^i} > \rho^i \eta^i$: in this case an event is triggered and the vector $\hat{\xi^i}$ gets updated to $\xi^i$ after the activation function application, in this situation \eqref{eqn:sec_cond} is met and with respect to remark \ref{rem:sec-cond} and Lemma \ref{lem:finsler} we have ${\xi^i}\tr X^i \xi^i < {\xi^i}\tr \Omega^i \xi^i < 0$.
+Once again we have ${\eta^i}^+ = \rho^i \eta^i - {\xi^i}\tr X^i \xi^i > 0 $
+\end{itemize}
+By denoting $\bX = \text{diag} \pmx{R_1\tr X^1 R_1 & \dots & R_l\tr X^l & R_l} $ and $\xi = \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr$
+$$
+\mathbf{\Psi_x} = \pmx{\Id_l \otimes \xi}\tr\! \bX \pmx{\1_l \otimes \xi}
+$$
+the ETM triggering condition and dynamic thresholds' dynamics can be rewritten in an element-wise fashion as:
 \begin{equation}
 \begin{aligned}
-    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \bX \1_l\\
-    \bX \1_l&> \bR \boldsymbol{\eta}\end{aligned}
+    \boldsymbol{\eta}^+ &= \bR \boldsymbol{\eta} - \mathbf{\Psi_x}\\
+    \mathbf{\Psi_x} &> \bR \boldsymbol{\eta}\end{aligned}
 \end{equation}
@@
   R N_{\nu x} & \Id - R & 0\\
-} \in \RR^{n_x + n_{\phi} + n_q \times n_x + 2 n_{\phi}}\\
-R_s &\in \RR^{n_x + n_{\phi} + n_q \times n_x + n_q}
+} \in \RR^{(n_x + n_{\phi} + n_q) \times (n_x + 2 n_{\phi})}\\
+R_s &\in \RR^{(n_x + n_{\phi} + n_q) \times (n_x + n_q)}
 \end{aligned}
@@
 
-\begin{theorem} \emph{ Consider the control system in \eqref{eq:system-dynamics} and the controller $\pi_{\text{ETM}}$ in \eqref{eq:nn-equations}.
-Assuming the existence of the matrices $P = P\tr > 0 \in \RR^{n_x \times n_x}$,} $T = \text{diag} \pmx{T^1 & \dots & T^l} \in \RR^{n_{\phi} \times n_{\phi}} > 0$, \emph{$Z = \smallpmat{{Z^1}\tr & \dots & {Z^l}\tr}\tr \in \RR^{n_{\phi} \times n_x}$, $\bR > 0$, $\bX$, $N^i_1, N^i_2, N^i_3 \ i \in \left\{ 1, \dots, l \right\}$, $S \in \RR^{n_x \times n_x}, Q \in \RR^{n_x \times n_q}, W \in \RR^{n_q \times n_q}$ and a scalar $\alpha > 0$ such that the matrix inequalities \eqref{eqn:lyapunov}, \eqref{eqn:inclusion}, \eqref{eqn:finsler} hold, where $\hat{\nu}_j^i = \min(|-\bar{\nu}_j^i - \nu^i_{*, j}|,|\bar{\nu}_j^i - \nu^i_{*, j}|)$, $Z^i = T^i G^i,\ i \in \left\{ 1, \dots, l \right\}$\\
-Then the system controlled by $\pi_{\text{ETM}}$ is locally stable around $x_*$ and the set $\cE(P, x_*) = \left\{ x \in \RR^{n_x} : \tilde{x}\tr P \tilde{x} \leq 1\right\}$ is an estimate of the ROA.}
-\end{theorem}
-
 \begin{figure*}
+\begin{subequations}\label{eqn:LMI}
 \begin{align}\label{eqn:lyapunov}
-  &\bmx{
+  &\Xi =\bmx{
   \begin{array}{c|c}
@@
   0 & 0 & 0
-  } + R_s\tr \bmx{
+  } - (\1_l \otimes R_{\nu})\tr \He{\bX} (\1_l \otimes R_{\nu})
+   + R_s\tr \bmx{
     S & Q\\
     \star & W
-  }  R_s - (\1_l \otimes R_{\nu})\tr He(\bX) (\1_l \otimes R_{\nu}) \hspace{.1em}& 0\\[1.5em]
+  }  R_s\hspace{.1em}& 0\\[1.5em]
 \hline \\[-0.9em]
@@
   &{R^i}\tr\!
-He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}
+He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
+  &\bmx{P & {E_i}\tr \\ \star & \hat{\mu_i}^2} \geq 0 \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
 \end{align}
+\end{subequations}
+\vspace{.5em}
+\hrule
+\end{figure*}
+\begin{figure*}
+\begin{equation}\label{eqn:first-step}
+ \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
+  \bmx{\bar{A}\tr\\
+  \bar{B}\tr\\
+  C\tr
+  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
+  P & 0 & 0\\
+  0 & 0 & 0\\
+  0 & 0 & 0
+  } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}
+ - \pmx{\1_l \otimes \xi}\tr \bx \pmx{\1_l \otimes \xi}
+ + \bmx{\tilde{x}\\ \tilde{\Phi}}\tr\! \bmx{S & Q\\ \star & W} \bmx{\tilde{x} \\ \tilde{\Phi}}
+ + \sqrt{\boldsymbol{\eta}}\tr \He{\bR - \Id} \sqrt{\boldsymbol{\eta}}
+\end{equation}
 \vspace{.5em}
@@
 
-\begin{proof} We start by taking into account a candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + He(\1_l\tr\!\boldsymbol{\eta})$.
-We start by commenting $R_{\nu}$ and $R_s$ in \eqref{eqn:auxiliary}, we notice how $\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\nu}\tr}\tr = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}\tr}\tr$.
-The matrix $\bmx{S & Q\\ \star & W}$ represents the quadratic sector conditions specific to $\Psi$ with respect to $\tilde{x}$ and $\tilde{\Psi}$, $R_s$ is a transformation matrix such that: $\bmx{\tilde{x}\tr & \tilde{\Psi}}\tr = R_s \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}\tr}\tr$
+\begin{theorem}\label{thm:stability} \emph{ Consider the control system in \eqref{eqn:system-dynamics} and the controller $\pi_{\text{ETM}}$ in \eqref{eqn:nn-equations}.
+Assuming the existence of the matrices $P = P\tr > 0 \in \RR^{n_x \times n_x}$,} $T = \text{diag} \pmx{T^1 & \dots & T^l} \in \RR^{n_{\phi} \times n_{\phi}} > 0$, \emph{$Z = \smallpmat{{Z^1}\tr & \dots & {Z^l}\tr}\tr \in \RR^{n_{\phi} \times n_x}$, $\bR > 0$, $\bX$, $N^i_1, N^i_2, N^i_3 \ \forall i \in \left\{ 1, \dots, l \right\}$, $S \in \RR^{n_x \times n_x}, Q \in \RR^{n_x \times n_q}, W \in \RR^{n_q \times n_q}$ and a scalar $\alpha > 0$ such that the matrix inequalities \eqref{eqn:lyapunov}, \eqref{eqn:inclusion}, \eqref{eqn:finsler} hold, where $\hat{\nu}_j^i = \min(|-\bar{\nu}_j^i - \nu^i_{*, j}|,|\bar{\nu}_j^i - \nu^i_{*, j}|)$, $
+\hat{\mu}_i = \min(|-\underline{\mu}_i - v_i^*|,|\bar{\mu}_i - v_i^*|)$ and $G^i = (T^i)^{-1} Z^i\ \forall i \in \left\{1, \dots, l \right\}$
+Then the point $(x, \boldsymbol{\eta}) = (x_*, \0)$ is Locally Exponentially Stable (LES) with a ROA that includes the ellipsoid $\cE(P, x_*) = \left\{ x \in \RR^{n_x}, \boldsymbol{\eta} \in \RR^{l} : \tilde{x}\tr P \tilde{x} + \He{\1_l\tr \boldsymbol{\eta}} \leq 1\right\}$}
+\end{theorem}
+
+\begin{proof}
+By exploiting the inequality $$\bmx{\alpha (\hat{\nu}^i_j)^{-2} - T^i_{j, j}}(\hat{\nu}^i_j)^2\bmx{\alpha (\hat{\nu}^i_j)^{-2} - T^i_{j, j}} \geq 0$$ having $(T^i_{j,j})^2(\hat{\nu}^i_j)^2 \geq 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}$ and assuming \eqref{eqn:inclusion} we state:
+$$
+\bmx{
+P & (T^i_{j,j} G^i_j)\tr\\
+\star & (T^i_{j,j})^2(\hat{\nu}^i_j)^2
+} \geq 
+\bmx{
+P & {Z_j^i}\tr\\
+\star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
+} \geq 0
+$$
+That implies
+$$
+\bmx{
+P & 0 & (T^i_{j,j} G^i_j)\tr\\
+0 & 2 \Id & 0\\
+\star & 0 &(T^i_{j,j})^2(\hat{\nu}^i_j)^2
+} \geq 0
+$$
+We initially assume the positivity of $\boldsymbol{\eta}$, by pre and post multiplying for $\bmx{\tilde{x}\tr, \sqrt{\boldsymbol{\eta}}, 1}\tr$ and its transpose and applying Schur's complement we obtain
+$$
+\bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}}\tr \bmx{P & 0\\ 0 & 2 \Id} \bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}} \geq \tilde{x}\tr \frac{(G^i_j)\tr (G^i_j)}{\hat{\nu^i_j}^2} \tilde{x}
+$$
+This ensures $\cE(P, x_*) \subseteq S$ defined in Remark \ref{rem:sec-set}.
+Now assuming \eqref{eqn:finsler_constraint} hold $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$, by the application of Lemma \ref{lem:finsler} we ensure $\boldsymbol{\eta} > 0$ through Remark \ref{rem:eta-positive}.
+Considering \eqref{eqn:nonlin-inclusion-constraint} we have
+$$
+\bmx{P & 0 & E_i\tr \\
+0 & 2 \Id & 0 \\
+E_i & 0 & \hat{\mu_i}^2} \geq 0
+$$
+By expressing $v_i = E_i x$ and pre and post multiplying by $\bmx{\tilde{x}\tr & \tilde{\Phi}(Ex)\tr}\tr$ we have:
+$$
+\bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}}\tr \bmx{P & 0\\ 0 & 2 \Id} \bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}} \geq \tilde{x}\tr \frac{E_i\tr E_i}{\hat{\mu_i}^2} \tilde{x} = \frac{v_i\tr v_i}{\hat{\mu_i}^2} \geq 0
+$$
+Once again this ensures $\cE(P, x_*) \subseteq \bar{S}$ defined in Assumption \ref{ass:nonlin} hence $\cE(P, x_*) \subseteq \left\{ S \cap \bar{S} \right\}$.
+
+We can define the following definite-positive candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + \He{\1_l\tr\!\boldsymbol{\eta}}$. 
+By denoting $R_s$ as a proper transformation matrix such that $\bmx{\tilde{v}\tr & \tilde{\Phi}(Ex)}\tr = R_s \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)}\tr$ then
+\begin{equation}\label{eqn:sector-sign}
+\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)}\tr R_s\tr \bmx{S & Q \\ \star & W} R_s \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)} \geq 0\ \forall x \in \cE(P, x_*)
+\end{equation}
+With respect to \eqref{eqn:auxiliary} and \eqref{eqn:incremental} by denoting $\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\nu}\tr}\tr = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr$, if we pre and post multiply \eqref{eqn:lyapunov} by $\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}\tr & \sqrt{\boldsymbol{\eta}}\tr}\tr$ and its transpose\\
+
+we obtain \eqref{eqn:first-step}. Considering \eqref{eqn:sector-sign} we verify Assumption \ref{ass:nonlin} then by collecting the proper terms and exploiting Kronecker product's properties we have
+\begin{equation}
+\begin{aligned}
+&\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
+  \bmx{\bar{A}\tr\\
+  \bar{B}\tr\\
+  C\tr
+  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
+  P & 0 & 0\\
+  0 & 0 & 0\\
+  0 & 0 & 0
+  } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}} + \\
+  &+ \He{\1_l\tr (\bR - \Id) \boldsymbol{\eta}} - \mathbf{\Psi_x} < 0\\
+\end{aligned}
+\end{equation}
+If all conditions in \eqref{eqn:LMI} are met then \eqref{eqn:lyapunov} implies the following:
+\begin{equation}
+    (\tilde{x}^+)^{\top} P \tilde{x}^+ - \tilde{x}\tr P \tilde{x} + \He{\1_l\tr (\boldsymbol{\eta}^+ - \boldsymbol{\eta})} = \Delta V < 0, 
+\end{equation}
+{\color{blue} for all $(x, \beta)\in \cE(P)$. Therefore, $\cE(P)$ is forward invariant and all trajectories starting inside such a set converge to $(x,0)$, thus concluding the proof.}
+We conclude the proof by stating that $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$ the point $(x_*, \0)$ is LES, and $\Delta V(x, \boldsymbol{\eta}) < 0$, considering that the system is in Discrete Time (DT) then $\cE(P, x_*)$ is an invariant set and it is an inner-approximation of the ROA
 \end{proof}

### 2024-12-12 Paper draft/draft.tex
Theorem proof ended

@@
 
-\address[First]{Università di Trento, Trento, Italy (e-mail: marco.sterlini@studenti.unitn.it) }
+\address[First]{University di Trento, Trento, Italy (e-mail: marco.sterlini@studenti.unitn.it) }
 \address[Second]{LAAS-CNRS, Université de Toulouse, CNRS, Toulouse, France (e-mail:bu) }
@@
 
-We can express now the new inter-dependency between the variables of interest:
+We can express now the new interdependency between the variables of interest:
 \begin{equation}\label{eqn:incremental}
@@
 
-\begin{lemma}\label{lem:finsler}: \emph{Considering the triggering policy \eqref{eqn:dynamic_trig}, $R^i$ such that ${\xi^i}\tr = \!R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr\!
+\begin{lemma}\label{lem:finsler} \emph{Considering the triggering policy \eqref{eqn:dynamic_trig}, $R^i$ such that ${\xi^i}\tr = \!R^i \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr\!
 = R^i \xi\tr$, with reference to the structure in \eqref{eqn:sec_cond} and assuming \eqref{eqn:incremental} holds for some $\underline{\xi}, \bar{\xi} \in \RR^{n_x + 2n_{\phi}}$ such that $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists X^i \in\!
@@
   &{R^i}\tr\!
-He\left\{X^i - \Omega^i\right\} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
+\He{X^i - \Omega^i} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
   &\bmx{P & {E_i}\tr \\ \star & \hat{\mu_i}^2} \geq 0 \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
@@
 \end{figure*}
-\begin{figure*}
-\begin{equation}\label{eqn:first-step}
- \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
-  \bmx{\bar{A}\tr\\
-  \bar{B}\tr\\
-  C\tr
-  } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
-  P & 0 & 0\\
-  0 & 0 & 0\\
-  0 & 0 & 0
-  } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}
- - \pmx{\1_l \otimes \xi}\tr \bx \pmx{\1_l \otimes \xi}
- + \bmx{\tilde{x}\\ \tilde{\Phi}}\tr\! \bmx{S & Q\\ \star & W} \bmx{\tilde{x} \\ \tilde{\Phi}}
- + \sqrt{\boldsymbol{\eta}}\tr \He{\bR - \Id} \sqrt{\boldsymbol{\eta}}
-\end{equation}
-\vspace{.5em}
-\hrule
-\end{figure*}
+% \begin{figure*}
+% \begin{equation}\label{eqn:first-step}
+%  \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}\tr \left( \!
+%   \bmx{\bar{A}\tr\\
+%   \bar{B}\tr\\
+%   C\tr
+%   } P \bmx{\bar{A} & \bar{B} & C} - \bmx{
+%   P & 0 & 0\\
+%   0 & 0 & 0\\
+%   0 & 0 & 0
+%   } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}}
+%  - \pmx{\1_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi}
+%  + \bmx{\tilde{x}\\ \tilde{\Phi}}\tr\! \bmx{S & Q\\ \star & W} \bmx{\tilde{x} \\ \tilde{\Phi}}
+%  + \sqrt{\boldsymbol{\eta}}\tr \He{\bR - \Id} \sqrt{\boldsymbol{\eta}}
+% \end{equation}
+% \vspace{.5em}
+% \hrule
+% \end{figure*}
 
@@
 We can define the following definite-positive candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + \He{\1_l\tr\!\boldsymbol{\eta}}$. 
-By denoting $R_s$ as a proper transformation matrix such that $\bmx{\tilde{v}\tr & \tilde{\Phi}(Ex)}\tr = R_s \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)}\tr$ then
+By denoting $R_s$ as a proper transformation matrix such that $\bmx{\tilde{v}\tr & \tilde{\Phi}(Ex)}\tr = R_s \!
+\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)}\tr$ then the following holds $\forall x \in \cE(P, x_*)$:
 \begin{equation}\label{eqn:sector-sign}
-\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)}\tr R_s\tr \bmx{S & Q \\ \star & W} R_s \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)} \geq 0\ \forall x \in \cE(P, x_*)
+\bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)}\tr \! R_s\tr \! \bmx{S & Q \\ \star & W} R_s \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}(Ex)} \geq 0
 \end{equation}
-With respect to \eqref{eqn:auxiliary} and \eqref{eqn:incremental} by denoting $\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\nu}\tr}\tr = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr$, if we pre and post multiply \eqref{eqn:lyapunov} by $\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}\tr & \sqrt{\boldsymbol{\eta}}\tr}\tr$ and its transpose\\
-
-we obtain \eqref{eqn:first-step}. Considering \eqref{eqn:sector-sign} we verify Assumption \ref{ass:nonlin} then by collecting the proper terms and exploiting Kronecker product's properties we have
+With respect to \eqref{eqn:auxiliary} and \eqref{eqn:incremental} by denoting \\$\xi = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr = R_{\nu} \zeta$, if we pre and post multiply \eqref{eqn:lyapunov} by $\bmx{\zeta\tr & \sqrt{\boldsymbol{\eta}}\tr}\tr$ and its transpose, by expressing the midterm as follows:
+\begin{equation}
+\begin{aligned}
+  &\zeta\tr \pmx{\1_l \otimes R_{\nu}}\tr \He{\bX} \pmx{\1_l \otimes R_{\nu}} \zeta = \\
+  =& \pmx{\1_l 1 \otimes R_{\nu} \zeta}\tr \He{\bX} \pmx{\1_l 1 \otimes R_{\nu} \zeta} = \\
+  =& \pmx{\1_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi} = \\
+  =& \1_l\tr \pmx{\Id_l \otimes \xi}\tr \He{\bX} \pmx{\1_l \otimes \xi} = \1_l\tr \mathbf{\Psi_x}
+\end{aligned}
+\end{equation}
+and considering \eqref{eqn:sector-sign} we verify Assumption \ref{ass:nonlin} then by collecting the proper terms $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$ we have
 \begin{equation}
@@
   } \right) \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\Phi}} + \\
-  &+ \He{\1_l\tr (\bR - \Id) \boldsymbol{\eta}} - \mathbf{\Psi_x} < 0\\
+  &+ \He{\1_l\tr \bmx{(\bR - \Id) \boldsymbol{\eta} - \mathbf{\Psi_x}}} = \\
+  &= (\tilde{x}^+)^{\top} P \tilde{x}^+ - \tilde{x}\tr P \tilde{x} + \He{\1_l\tr (\boldsymbol{\eta}^+ - \boldsymbol{\eta})} = \Delta V < 0, 
 \end{aligned}
 \end{equation}
-If all conditions in \eqref{eqn:LMI} are met then \eqref{eqn:lyapunov} implies the following:
-\begin{equation}
-    (\tilde{x}^+)^{\top} P \tilde{x}^+ - \tilde{x}\tr P \tilde{x} + \He{\1_l\tr (\boldsymbol{\eta}^+ - \boldsymbol{\eta})} = \Delta V < 0, 
-\end{equation}
-{\color{blue} for all $(x, \beta)\in \cE(P)$. Therefore, $\cE(P)$ is forward invariant and all trajectories starting inside such a set converge to $(x,0)$, thus concluding the proof.}
-We conclude the proof by stating that $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$ the point $(x_*, \0)$ is LES, and $\Delta V(x, \boldsymbol{\eta}) < 0$, considering that the system is in Discrete Time (DT) then $\cE(P, x_*)$ is an invariant set and it is an inner-approximation of the ROA
+{\color{blue} Therefore, $\cE(P, x_*)$ is forward invariant and all trajectories starting inside such a set converge to $(x, \0)$, thus concluding the proof.}
 \end{proof}

### 2024-12-12 Paper draft/draft.tex
Optimization sector finished

@@
     \star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
-  } \geq 0 \ \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\} \label{eqn:inclusion}\\
+  } \geq 0 \qquad \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\} \label{eqn:inclusion}\\
   &{R^i}\tr\!
-\He{X^i - \Omega^i} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
-  &\bmx{P & {E_i}\tr \\ \star & \hat{\mu_i}^2} \geq 0 \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
+\He{X^i - \Omega^i} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \qquad \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
+  &\bmx{P & {E_i}\tr \\ \star & \hat{\mu_i}^2} \geq 0 \qquad \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
 \end{align}
@@
 \end{proof}
+
 \subsection{Optimization procedure}
+We state now a corollary whose results will be useful to derive some performances upgrade that will be later discussed.
+
+\begin{corollary}\label{cor:optimization}
+  \emph{Given Theorem \ref{thm:stability}, considering the decision variables $\Sigma, P, T, Z, \bX, \bR, S, Q, W, N^1_i, N^2_i, N^3_i, \beta_i \quad \forall i \in \left\{ 1, \dots, l \right\}$ with $\Sigma$ being a diagonal matrix $\Sigma > 0 \in \RR^{(n_x + 2 n_{\phi} + n_q) \times (n_x + 2 n_{\phi} + n_q)}$, and $\beta_i$ being scalars with the following optimization scheme:}
+  \begin{equation}\label{eqn:optimization}
+  \begin{aligned}
+    &\min \trace{P} + \trace{\Sigma} + \sum_{i=1}^{l} \beta_i\\
+    &\text{subject to } \eqref{eqn:LMI},\\ 
+    &\qquad \qquad \Xi + \Sigma \geq 0,\\
+    &\qquad \qquad \bmx{-\beta_i \Id & X^i\\
+    {X^i}\tr & -\Id} \leq 0 \qquad i \in \left\{ 1, \dots, l \right\}
+  \end{aligned}
+  \end{equation}
+  \emph{Then $(x_*, \0)$ is LES and $\cE(P, x_*)$ is an inner approximation of the ROA.}
+\end{corollary}
+
+By adding these constraints we obtain some desirable performances upgrade:
+\begin{itemize}
+  \item By minimizing the trace of $P$ we are maximizing the volume of the ellipsoid $\cE(P, x_*)$ hence the ROA.
+  \item By minimizing the trace of $\Sigma$ we are minimizing the conservatism of the controller by obtaining a constraint matrix $\Xi$ directly related to the decay rate of the Lyapunov function as close to zero as possible but never greater due to the second condition in \eqref{eqn:optimization}.
+  \item By taking \eqref{eqn:optimization} and applying the Schur complement to the last condition we obtain
+  $$
+  X^i {X^i}\tr \leq \beta_i \Id
+  $$
+  meaning that the smaller the value of $\beta_i$ the smaller will be the value that appears directly in the ETM triggering condition \eqref{eqn:etm-trigger} hence the lower the number of events.
+\end{itemize}
+
+Let us comment the quantity $\alpha$ that appears in \eqref{eqn:inclusion}, if this was considered as a variable then the statements of Theorem \ref{thm:stability} would not hold since we would encounter some bilinear terms.
+By fixing $\alpha$ to a constant value before the LMI evaluation we address the problem.
+The choice of $\alpha$ value is crucial since it will directly affect the conservatism of the solution that the problem will provide.
+The followed approach sees the deployment of a \emph{golden ratio search} to extract a good value.
+The algorithm is based on the bisection method with some basic improvements:
+\begin{itemize}
+  \item Using the golden ratio to divide the interval reduces the number of LMI solutions compared to the bisection method. Converging faster since it narrows the search interval more efficiently.
+  \item It maintains a constant reduction factor for the interval size, leading to more consistent progress towards the optimal value.
+  \item It is particularly useful for unimodal functions like the one we are dealing with after we set the volume of the ROA as the parameter to optimize.
+\end{itemize}
 

### 2024-12-12 Paper draft/draft.tex
Testo simulazioni fatto

@@
 
+{\color{red}{Mi fa cagare l'espressione della minimizzazione così}}
+
 By adding these constraints we obtain some desirable performances upgrade:
@@
 \end{itemize}
+{\color{red}{Vorrei migliorare questa parte}}
 
@@
 \section{Simulations}
+We considered an inverted pendulum system with mass $m = 0.15 kg$, length $l = 0.5 m$, damping coefficient $\mu = 0.05 N ms/rad$, gravitational acceleration $g = 9.81 m/s^2$ and the addition of an integrator in its dynamics.
+The DT model of the system is given with respect to \eqref{eqn:system-dynamics} by:
+\begin{equation}\label{eqn:pendulum}
+\begin{aligned}
+  &x = \bmx{\theta \\ \dot{\theta} \\ z} \quad 
+  A = \bmx{1 & dt & 0\\
+  \frac{g}{l} dt & 1 - \frac{\mu}{m l^2} dt & 0\\
+  1 & 0 & 1}\quad
+  B = \bmx{0\\ \frac{\bar{\tau}}{m l^2} dt\\ 0} \\
+  &C = \bmx{0 \\ \frac{g}{l} dt \\ 0} \quad
+  D = \bmx{0 \\ 0 \\ -1} \quad
+  E = \bmx{1 & 0 & 0}
+\end{aligned}
+\end{equation}
+With the maximum torque available $\bar{\tau} = 5 N m$, the sampling time $dt = 0.02 s$ and the non-linearity $\Phi(Ex) = \text{sin}(Ex) - Ex$.
+We have trained a stabilizing NN controller with Reinforcement Learning (RL) structured as an MLP of $3$ layers with $n_{\phi_1} = n_{\phi_2} = n_{\phi_3} = 32$.
+The addition of the integrator in the system dynamics made the training less trivial and allowed to compensate the intrinsic equilibrium points given by the controller that are generally different from the origin.
+The output of the controller $\bar{u}$ is a scalar, the saturation level $\hat{u}$ of the last ETM (see fig.
+\ref{fig:second_scheme}) is set to $1$ given the nature of RL training that always normalizes the final output of the network.
+We solved many instances of the LMI problem by first applying the ETM with null threshold, dynamic threshold, new ETM triggering matrices found with Lemma \ref{lem:finsler} and finally the matrices that we obtained by the application of Corollary \ref{cor:optimization} observing a considerable upgrade in performances in terms of computational saving.
+The results are shown in the following table:
+
+{\color{red}{Troppo discorsivo?}}
 

### 2024-12-15 Paper draft/draft.tex
Updated schemes

--- 2024-12-12 Paper draft/draft.tex
@@
 
-\title{Draft of paper}
+\title{{\color{red}\textbf{TODO}}}
 
@@
 
-\address[First]{University di Trento, Trento, Italy (e-mail: marco.sterlini@studenti.unitn.it) }
-\address[Second]{LAAS-CNRS, Université de Toulouse, CNRS, Toulouse, France (e-mail:bu) }
+\address[First]{\color{red}\textbf{TODO}}
+\address[Second]{\color{red}\textbf{TODO}}
 
 \begin{abstract}
-\textbf{INSERT ABSTRACT}
+{\color{red}\textbf{TODO}}
 \end{abstract}
@@
 \begin{keyword}
-\textbf{INSERT KEYWORDS}
+{\color{red}\textbf{TODO}}
 \end{keyword}
@@
 \section{Introduction}
-\textbf{INSERT INTRODUCTION}\\
+{\color{red}\textbf{TODO}}\\
 \emph{Notation:} $\NN, \RR^{n}, \RR^{n \times m}$ denote the set of natural non-negative integers, the set of real vectors of dimension $n$ and the set of real matrices of dimension $n \times m$, respectively.
@@
 \section{Problem Formulation}
-We consider the non-linear discrete-time system stabilized by a Neural Network (NN) controller represented in figure \ref{fig:first_scheme} described by:
+We consider the nonlinear discrete-time system stabilized by a Neural Network (NN) controller, as represented in Figure \ref{fig:first_scheme} and described by:
 \begin{equation}
@@
 $$
-    \texttt{sat}(\bar{u}) = \texttt{sign}(\bar{u})\text{min}(|\bar{u}|, \hat{u})
+    \texttt{sat}(\bar{u}) = \texttt{sign}(\bar{u})\text{min}(|\bar{u}|, \widehat{u})
 $$
-with $\hat{u}$ the limit of the input signal, $x \in \RR^{n_x}$ the state vector, $\bar{u} \in \RR^{n_u}$ the raw output of the controller, $d \in \RR^{n_d}$ a constant disturbance input  and $\Phi \in \RR^{n_q}$ the representation of the non-linearity of the system.
+here $\widehat{u}$ represents the limit of the input signal, $x \in \mathbb{R}^{n_x}$ is the state vector, $\bar{u} \in \mathbb{R}^{n_u}$ denotes the raw output of the controller, $d \in \mathbb{R}^{n_d}$ is a constant disturbance input, and $\Phi \in \mathbb{R}^{n_q}$ characterizes the nonlinearities of the system.
 \begin{assumption}\label{ass:nonlin}
@@
 \begin{equation}\label{eqn:nonlin-set}
-\forall v \in \bar{S} = \left\{ v \in  \RR^{n_q} : - \underline{\mu}_i \leq v_i \leq \bar{\mu}_i, i \in \pmx{1 & \dots & n_q} \right\} 
+\forall v \in \bar{S} := \left\{ v \in  \RR^{n_q} : - \underline{\mu}_i \leq v_i \leq \bar{\mu}_i, i \in \pmx{1 & \dots & n_q} \right\} 
 \end{equation}
@@
 \end{figure}
-
-The controller is implemented as a Multi Layer Perceptron (MLP) with $l$ layers, each with $n_{\phi_i}, i \in \left\{ 1, \dots, l\right\}$ neurons, all activation functions are saturation functions, represented by $\phi_i$.
-The objective is to design an Event-Triggering Mechanism (ETM) that allows to ease the computational burden associated with the computation of many non-linear activation functions providing at the same time sufficient sector conditions to handle the non-linearity of the system and the controller.
-The controller $\pi$ is defined by:
+The controller is implemented as a Multi-Layer Perceptron (MLP) with $l$ layers, each containing $n_{\phi_i}$ neurons for $i \in {1, \dots, l}$.
+All activation functions are saturation functions, represented by $\phi_i$.
+The goal is to design an Event-Triggering Mechanism (ETM) that alleviates the computational burden associated with evaluating numerous nonlinear activation functions while simultaneously providing sufficient sector conditions to manage the nonlinearities of both the system and the controller.
+The controller $\pi$ is defined as:
 \begin{equation}\label{eqn:nn-equations}
   \begin{aligned}
-  \hat{\omega}^{0}(k) &= x(k) \\
-  \nu^{i}(k) &= W^{i} \hat{\omega}^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
+  \widehat{\omega}^{0}(k) &= x(k) \\
+  \nu^{i}(k) &= W^{i} \widehat{\omega}^{i - 1}(k) + b^{i}, i \in \left\{ 1, \dots, l \right\}\\
   \omega^{i}(k) &= \phi_i(\nu^i(k))\\
@@
 \end{equation}
-with $\nu^i(k) \in \RR^{n_{\phi_i}}$ the input to the $i$-th activation function $\phi^i: \RR^{n_{\phi_i}} \to \RR^{n_{\phi_i}}$, $\omega^i(k), \hat{\omega}^i(k) \in \RR^{n_{\phi_i}}$ the current and last forwarded output respectively.
-The weights $W^i \in \RR^{n_{\phi_i} \times n_{\phi_{i-1}}}$ and biases $b^i \in \RR^{n_{\phi_i}}$ define the affine operation of each layer.
-The saturation function application is to be intended element-wise and will be denoted as $\phi^i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$ and $\varphi: \RR \to \RR$ is the scalar activation function assumed to be symmetric and identical for every neuron.
-Adopting the same notation as in \cite{css-paper} we define the controller policy in a condensed form.
-Denoting the augmented vectors:
+$\nu^i(k) \in \mathbb{R}^{n_{\phi_i}}$ represents the input to the $i$-th activation function $\phi^i: \mathbb{R}^{n_{\phi_i}} \to \mathbb{R}^{n_{\phi_i}}$, and $\omega^i(k), \widehat{\omega}^i(k) \in \mathbb{R}^{n_{\phi_i}}$ denote the current and last forwarded outputs, respectively.
+The weights $W^i \in \mathbb{R}^{n_{\phi_i} \times n_{\phi_{i-1}}}$ and biases $b^i \in \mathbb{R}^{n_{\phi_i}}$ define the affine transformation of each layer.
+The application of the saturation function is element-wise and is denoted as $\phi^i(\nu^i(k)) = \left[ \varphi(\nu^i_1(k)), \dots, \varphi(\nu^i_{n_{\phi_i}}(k)) \right]$, where $\varphi: \mathbb{R} \to \mathbb{R}$ is a scalar activation function that is assumed to be symmetric and identical for every neuron.
+
+Following the notation in \cite{css-paper}, the controller policy is expressed in a condensed form. Introducing the augmented vectors:
 \begin{equation*}
@@
 \end{equation*}
-with $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \RR^{n_{\phi}} \to \RR^{n_{\phi}}$ the combined non-linearity we have $\omega = \phi(\nu)$.
-Finally, conditions \eqref{eqn:nn-equations} can be rewritten as:
+With $n_{\phi} = \sum_{i=1}^{l} n_{\phi_i}$ and $\phi: \mathbb{R}^{n_{\phi}} \to \mathbb{R}^{n_{\phi}}$ representing the combined nonlinearity, we have $\omega = \phi(\nu)$.
+Finally, the conditions in \eqref{eqn:nn-equations} can be reformulated as:
 \begin{equation*}
@@
 \begin{equation}\label{eqn:first-N}
-  \overbar N = \bmx{
+  \overbar N := \bmx{
     \begin{array}{c | c c c c | c}
@@
 \end{equation}
-Similarly as in \cite{css-extended} we implement an ETM in every layer of the controller, we add also a final one to the output that will be used to further reduce the total number of events and calls to the non-linear functions.
-The system and the controller are now described by figure \ref{fig:second_scheme}.
+Similarly to \cite{css-extended}, we implement an ETM in every layer of the controller, with an additional one at the output.
+This final ETM is designed to further reduce the total number of events and calls to the nonlinear functions.
+The system and controller are now represented in Figure \ref{fig:second_scheme}.
 \begin{figure}[H]
@@
 \end{figure}
-The dynamics will be rewritten as:
+
+The system dynamics are rewritten as:
 \begin{equation}\label{eqn:system-dynamics}
@@
 \end{equation}
-with $u \in \RR^{n_u}$ the output of the controller subject to the ETM. This modification required to redefine the matrix $N$ defined in the controller policy \eqref{eqn:first-N} as:
+where $u \in \mathbb{R}^{n_u}$ is the output of the controller subject to the ETM.
+This modification requires redefining the matrix $\overbar N$, originally defined in the controller policy \eqref{eqn:first-N}, as:
 \begin{equation}\label{eqn:last-N}
-  N = \bmx{
+  N := \bmx{
     \begin{array}{c | c c c c | c}
@@
 \end{equation}
-
-The system is compliant with the assumptions of \citep[Lemma 2]{css-extended} that allow us to define the following expressions regarding the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller given $x_*$, defining the matrices:
+The system adheres to the assumptions outlined in \citep[Lemma 2]{css-extended}, which allow us to derive the following expressions for the equilibrium points $\left( x_*, u_*, \nu_*, \omega_* \right)$ of the system and controller, given $x_*$.
+The matrices are defined as:
 \begin{equation}
@@
 \end{equation}
-
-With $R$ always invertible due to $N_{\nu \omega}$ always invertible due to its lower triangular structure.
-We find the equilibrium of the state by the use of a numerical solver that minimizes the following expression since it is no longer explicit due to the non-linearity $\Phi$.
+Given the lower triangular structure of $N_{\nu \omega}$, the expression for $R$ is always invertible and exists.
+The equilibrium of the state is found using a numerical solver that minimizes the following expression, as it is no longer explicit due to the nonlinearity $\Phi$.
 Given a constant disturbance $\bar{d}$:
@@
 \end{equation}
-All the non-linearity of the system and the controller are selected as saturation functions.
-With respect to \citep[Lemma 3]{css-extended}, we declare the local sector conditions that hold for every state included in the following set:
-
+All the nonlinearities of the system and the controller are chosen to be saturation functions.
+In accordance with \citep[Lemma 3]{css-extended}, we state the local sector conditions that hold for every state within the following set:
 \begin{remark}\label{rem:sec-set}
@@
 \begin{equation}\label{eqn:inclusion-set}
-S = \left\{ x \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}^i_* \leq G^i (x - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
+S := \left\{ x \in \RR^{n_x} : -\bar{\nu}^i - \bar{\nu}^i_* \leq G^i (x - x_*) \leq \bar{\nu}^i - \nu_*^i \right\} 
 \end{equation}
@@
 \end{equation}\end{remark}
-Where $\bar{\nu}^i$ is the saturation level of the $i$-th layer.
+Where $\bar{\nu}^i$ represents the saturation level of the $i$-th layer.
 
-In order to have a lighter notation the dead-zone vector  will be defined as $\psi^i = \nu^i - \omega^i$ and $\hat{\psi}^i = \nu^i - \hat{\omega}^i$.
-Furthermore, all following discussions will take into account incremental variables with respect to their equilibrium, denoted with a tilde, i.e.
-$\tilde{x} = x - x_*$.
-Considering \eqref{eqn:equilibrium} we have $\psi_* = \nu_* - \omega_* = 0$ hence $\tilde{\psi} = \psi$.
-We define $\xi^i = \bmx{\tilde{x}\tr & \tilde{\psi}_i\tr & \tilde{\nu}_i\tr}\tr$ and $\hat{\xi}^i = \bmx{\tilde{x}\tr & \hat{\tilde{\psi}}_i\tr & \tilde{\nu}_i\tr}\tr$, then the overall sector conditions \eqref{eqn:first_sec_cond} can be rewritten with no loss of generality $\forall i \in \pmx{1 & \dots & l}$ as:
+For ease of reference, the dead-zone vector is defined as $\psi^i = \nu^i - \omega^i$ and $\widehat{\psi}^i = \nu^i - \widehat{\omega}^i$.
+Additionally, all subsequent discussions will use incremental variables relative to their equilibrium, denoted with a tilde, i.e., $\tilde{x} = x - x_*$.
+Considering \eqref{eqn:equilibrium}, we have $\psi_* = \nu_* - \omega_* = 0$, hence $\tilde{\psi} = \psi$.
+We define $\xi^i = \begin{bmatrix} \tilde{x}^\top & \tilde{\psi}_i^\top & \tilde{\nu}_i^\top \end{bmatrix}^\top$ and $\widehat{\xi}^i = \begin{bmatrix} \tilde{x}^\top & \widehat{\tilde{\psi}}_i^\top & \tilde{\nu}_i^\top \end{bmatrix}^\top$.
+Then, the overall sector conditions \eqref{eqn:first_sec_cond} can be rewritten, with no loss of generality, for all $i \in {1, \dots, l}$ as:
 \begin{equation}\label{eqn:sec_cond}
@@
 \begin{remark}\label{rem:sec-cond}
-Note how ensuring ${\xi^i}\tr \Omega^i \xi^i < 0$ guarantees that the sector condition \eqref{eqn:sec_cond} holds.
+Note that ensuring ${\xi^i}^\top \Omega^i \xi^i < 0$ guarantees that the sector condition \eqref{eqn:sec_cond} holds.
 \end{remark}
 
-We can express now the new interdependency between the variables of interest:
+We can now express the new interdependency between the variables of interest as follows:
 \begin{equation}\label{eqn:incremental}
@@
 
-Taking inspiration from \citep[Proposition 1, Lemma 3]{css-extended} we design the ETM for the output of each layer of the controller and the final output.
+Inspired by \citep[Proposition 1, Lemma 3]{css-extended}, we design the ETM for the output of each layer of the controller and the final output.
+This is achieved by first incorporating a dynamic threshold and then refining the triggering function using Finsler's lemma.
 
 \subsection{Event-triggering mechanism}
-An ETM works by triggering an event that will update the current layer and propagate its value across the network. 
-In the precedent work \cite{css-extended} the triggering function is defined on the basis of the saturation sector conditions in the form \eqref{eqn:sec_cond}. The ETM was defined as:
+An ETM operates by triggering an event to update the current layer and propagate its value throughout the network.
+In the previous work \cite{css-extended}, the triggering function was designed based on the saturation sector conditions in the form of \eqref{eqn:sec_cond}.
+The ETM was defined as:
 \begin{equation}
-  \hat{\omega}^{i}(k) = \begin{cases}
-    \phi^i(\nu^i(k)) & \text{if } {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i} > 0\\
-    \hat{\omega}^{i}(k-1) & \text{otherwise}
+  \widehat{\omega}^{i}(k) = \begin{cases}
+    \phi^i(\nu^i(k)) & \text{if } {\widehat{\xi^i}}\tr \Omega^i \widehat{\xi^i} > 0\\
+    \widehat{\omega}^{i}(k-1) & \text{otherwise}
   \end{cases}
 \end{equation}
-
-This approach acts as a static ETM triggering an event immediately after the sector conditions are violated.
-Inspired by the work of \cite{data-driven} we define a dynamic quantity to be implemented in the ETM triggering condition.
-The idea is to lower the conservatism of the controller by introducing a time-varying threshold that will adapt to the current state of the system and controller reducing the overall number of events and computational burden.
-The new triggering function for the new \emph{dynamic-ETM} for the $i$-th layer is defined as:
+This approach functions as a static ETM, triggering an event immediately when the sector conditions are violated.
+Building upon the concepts from \cite{data-driven}, we propose a dynamic quantity to be incorporated into the ETM triggering condition.
+The goal is to reduce the conservatism of the controller by introducing a time-varying threshold that adapts to the current state of the system and controller, thereby decreasing the overall number of events and computational burden.
+The new triggering function for the proposed \emph{dynamic-ETM} for the $i$-th layer is defined as:
 \begin{equation}\label{eqn:dynamic_trig}
-  \hat{\omega}^{i}(k) = \begin{cases}
-    \phi^i(\nu^i(k)) & \text{if } {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i} > \rho^i \eta_i\\
-    \hat{\omega}^{i}(k-1) & \text{otherwise}
+  \widehat{\omega}^{i}(k) = \begin{cases}
+    \phi^i(\nu^i(k)) & \text{if } {\widehat{\xi^i}}\tr \Omega^i \widehat{\xi^i} > \rho^i \eta_i\\
+    \widehat{\omega}^{i}(k-1) & \text{otherwise}
   \end{cases};
@@
 
-Where $\eta^i$ characterized by the following dynamics:
+Where $\eta^i$ is characterized by the following dynamics:
 \begin{equation}\label{eqn:eta_dynamics}
-  {\eta^i}^+ = \rho^i \eta^i - {\hat{\xi^i}}\tr \Omega^i \hat{\xi^i}\quad  \forall i \in \left( 1, \dots, l \right) 
+  {\eta^i}^+ = \rho^i \eta^i - {\widehat{\xi^i}}\tr \Omega^i \widehat{\xi^i}\quad  \forall i \in \left( 1, \dots, l \right) 
 \end{equation}
 
-We denote a condensed notation by taking into account all conditions for each layer
+We introduce a condensed notation that incorporates all conditions across each layer.
 \begin{equation}
@@
    \bR &= \text{diag} \pmx{\rho^1 & \dots & \rho^l }\\
-   \mathbf{\Psi} &= \bmx{ {\hat{\xi^1}}\tr \Omega^1 \hat{\xi^1} & \dots & {\hat{\xi^l}}\tr \Omega^l \hat{\xi^l}}\tr\\
+   \mathbf{\Psi} &= \bmx{ {\widehat{\xi^1}}\tr \Omega^1 \widehat{\xi^1} & \dots & {\widehat{\xi^l}}\tr \Omega^l \widehat{\xi^l}}\tr\\
    \boldsymbol{\eta} &= \bmx{ \eta^1 & \dots & \eta^l}\tr   
@@
 \subsection{Finsler Lemma application}
-After the implementation of the dynamic threshold a further improvement of the triggering condition has been taken in consideration. By recalling the triggering condition:
+Following the implementation of the dynamic threshold, a further refinement of the triggering condition has been considered. Recalling the triggering condition:
 $$
-\hat{{\xi^i}}\tr \Omega^i \hat{\xi^i} > \rho^i \eta^i
+\widehat{{\xi^i}}\tr \Omega^i \widehat{\xi^i} > \rho^i \eta^i
 $$
-Intuitively the lower the left-hand term of the inequality the lower will be the number of triggered events.
-Unfortunately $\Omega^i$ structure is non-flexible due to his strict relation with the sector condition expression.
-The following lemma addresses the issue by designing some triggering matrices that guarantee a lower update rate for each layer by the use of Finsler's lemma:
+Intuitively, the smaller the left-hand side of the inequality, the fewer events will be triggered.
+However, the structure of $\Omega^i$ is inflexible due to its strict relationship with the sector condition expression.
+The following lemma addresses this limitation by introducing triggering matrices designed to ensure a lower update rate for each layer, leveraging Finsler's lemma:
 
@@
 = R^i \xi\tr$, with reference to the structure in \eqref{eqn:sec_cond} and assuming \eqref{eqn:incremental} holds for some $\underline{\xi}, \bar{\xi} \in \RR^{n_x + 2n_{\phi}}$ such that $\forall \xi \in \left[ \underline{\xi}, \bar{\xi}\right]$, if $\exists X^i \in\!
-\RR^{(nx + 2n_{\phi_i}) \times (nx + 2n_{\phi_i})},$ $N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ a diagonal matrix $T^i > 0 \in \RR^{n_{\phi_i} \times n_{\phi_i}}, G^i \in \RR^{n_{\phi_i} \times n_x}$ such that:
+\RR^{(nx + 2n_{\phi_i}) \times (nx + 2n_{\phi_i})},$ $N^i_1 \in \RR^{n_x \times n_{\phi_i}}, N^i_2, N^i_3 \in \RR^{n_{\phi_i} \times n_{\phi_i}}$ a diagonal matrix $T^i \succ 0 \in \RR^{n_{\phi_i} \times n_{\phi_i}}, G^i \in \RR^{n_{\phi_i} \times n_x}$ such that:
 \begin{equation}\label{eqn:finsler}
-     {R^i}\tr\! He\left\{X^i - \Omega^i\right\}\! R^i + He\left\{ \smallpmat{N^i_1\\ N^i_2\\ N^i_3}\! \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0
+     {R^i}\tr\! He\left\{X^i - \Omega^i\right\}\! R^i + He\left\{ \smallpmat{N^i_1\\ N^i_2\\ N^i_3}\! \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \preceq 0
 \end{equation}
@@
 
-\begin{proof} By pre and post multiplying for $\xi$ the previous expression we end up with
+\begin{proof} By pre- and post-multiplying for $\xi$ the previous expression we end up with
 $$
@@
 $$
-By expanding the term $\pmx{R N_{\nu x} & \Id - R & -\Id} \xi$ we end up with condition \eqref{eqn:incremental} meaning that the expression is reduced to
+By expanding the term $\pmx{R N_{\nu x} & \Id - R & -\Id} \xi$ we end up with condition \eqref{eqn:incremental}, meaning that the expression is reduced to
 $$
@@
 $$
-By further expanding the expression and considering that it is eventually reduced to a scalar for which it holds that the transpose is equivalent to itself we obtain:  
+By further expanding the expression and noting that it eventually reduces to a scalar, for which the transpose is equivalent to itself, we obtain:
 \begin{equation*}
@@
     &{\xi^i}\tr (X^i + {X^i}\tr) \xi^i \leq {\xi^i}\tr (\Omega^i + {\Omega^i}\tr) \xi^i\\
-    &2 {\xi^i}\tr X^i\xi^ \leq 2 {\xi^i}\tr \Omega^i \xi^i\\
+    &2 {\xi^i}\tr X^i\xi^i \leq 2 {\xi^i}\tr \Omega^i \xi^i\\
     & {\xi^i}\tr X^i \xi^i \leq {\xi^i}\tr \Omega^i \xi^i
@@
 
-Assuming we have the $X^i$ matrices $i \in \left\{1, \dots, l \right\}$ we rewrite the triggering conditions for each layer:
+Given $X^i$ matrices $i \in \left\{1, \dots, l \right\}$, we can rewrite the triggering conditions for each layer as:
 \begin{equation}\label{eqn:etm-trigger}
-  \hat{\omega}^{i}(k) = \begin{cases}
+  \widehat{\omega}^{i}(k) = \begin{cases}
     \phi^i(\nu^i(k)) & \text{if } {\xi^i}\tr X^i \xi > \rho^i \eta^i(k)\\
-    \hat{\omega}^{i}(k-1) & \text{otherwise}
+    \widehat{\omega}^{i}(k-1) & \text{otherwise}
   \end{cases}
 \end{equation}
-\begin{remark}\label{rem:eta-positive} \emph{Similarly as in \citep[Lemma 3]{data-driven} it is possible to demonstrate how $\eta^i$ being the dynamic threshold of layer $i$ is always non-negative for any $\rho^i \in [0, 1)$, $\eta^i_0 \geq\ 0$} \end{remark}
+\begin{remark}\label{rem:eta-positive} \emph{Similarly to \citep[Lemma 3]{data-driven}, it can be demonstrated that $\eta^i$, the dynamic threshold of layer $i$, is always non-negative for any $\rho^i \in [0, 1)$, with the initial condition $\eta^i_0 \geq 0$.} \end{remark}
 With respect to \eqref{eqn:etm-trigger} and \eqref{eqn:eta_dynamics} we have two possible occurrences:
 \begin{itemize}
-    \item $\hat{{\xi^i}}\tr X^i \hat{\xi^i} \leq \rho^i \eta^i$: it's trivial to verify that ${\eta^i}^+ = \rho^i \eta^i - \hat{{\xi^i}}\tr X^i \hat{\xi^i} \geq 0 $
-    \item $\hat{\xi^i}\tr X^i \hat{\xi^i} > \rho^i \eta^i$: in this case an event is triggered and the vector $\hat{\xi^i}$ gets updated to $\xi^i$ after the activation function application, in this situation \eqref{eqn:sec_cond} is met and with respect to remark \ref{rem:sec-cond} and Lemma \ref{lem:finsler} we have ${\xi^i}\tr X^i \xi^i < {\xi^i}\tr \Omega^i \xi^i < 0$.
-Once again we have ${\eta^i}^+ = \rho^i \eta^i - {\xi^i}\tr X^i \xi^i > 0 $
+    \item $\widehat{{\xi^i}}\tr X^i \widehat{\xi^i} \leq \rho^i \eta^i$: It is trivial to verify that ${\eta^i}^+ = \rho^i \eta^i - \widehat{{\xi^i}}\tr X^i \widehat{\xi^i} \geq 0 $
+    \item $\widehat{\xi^i}\tr X^i \widehat{\xi^i} > \rho^i \eta^i$: In this case, an event is triggered, and the vector $\widehat{\xi^i}$ gets updated to $\xi^i$ after the activation function application.
+In this situation, \eqref{eqn:sec_cond} is satisfied, and, with respect to Remark \ref{rem:sec-cond} and Lemma \ref{lem:finsler}, we have ${\xi^i}^\top X^i \xi^i < {\xi^i}^\top \Omega^i \xi^i < 0$.
+Once again, we have ${\eta^i}^+ = \rho^i \eta^i - {\xi^i}^\top X^i \xi^i > 0$.
 \end{itemize}
-By denoting $\bX = \text{diag} \pmx{R_1\tr X^1 R_1 & \dots & R_l\tr X^l & R_l} $ and $\xi = \bmx{\tilde{x} & \tilde{\psi} & \tilde{\nu}}\tr$
-$$
+By denoting 
+\begin{equation}
+    \bX = \text{diag} \pmx{R_1\tr X^1 R_1 & \dots & R_l\tr X^l & R_l}, \qquad \xi = \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\nu}}
+\end{equation}
+\begin{equation}\label{eqn:final-psi}
     \mathbf{\Psi_x} = \pmx{\Id_l \otimes \xi}\tr\! \bX \pmx{\1_l \otimes \xi}
-$$
-the ETM triggering condition and dynamic thresholds' dynamics can be rewritten in an element-wise fashion as:
+\end{equation}
+The ETM triggering condition and the dynamics of the dynamic thresholds can be rewritten in an element-wise fashion as:
 \begin{equation}
@@
 \subsection{Lyapunov conditions}
-In this section we will provide sufficient conditions for the stability of the system and the controller under the form of Linear Matrix Inequalities (LMIs) with the guarantee of local asymptotic stability.
-The solution of the problem will also provide the matrices $X^i$ and the parameters $\rho^i$ that will be used in the ETM triggering conditions along with an estimate of the Region Of Attraction (ROA) of the equilibrium point.
+In this section, we will provide sufficient conditions for the stability of the system and controller in the form of Linear Matrix Inequalities (LMIs), ensuring local asymptotic stability.
+The solution to the problem will also yield the matrices $X^i$ and the parameters $\rho^i$ to be used in the ETM triggering conditions, along with an estimate of the Region of Attraction (ROA) for the equilibrium point.
 
-We will first introduce a more compact notation for the system dynamics as long as some auxiliary matrices:
+We will first introduce a more compact notation for the system dynamics, along with some auxiliary matrices:
 \begin{equation}\label{eqn:auxiliary}
@@
   0 & \hspace{.1em} 2 (\bR - \Id)
-  \end{array}} < 0\\
+  \end{array}} \prec 0\\
   &\bmx{
     P & {Z_j^i}\tr\\
-    \star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
-  } \geq 0 \ \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\} \label{eqn:inclusion}\\
+    \star & 2 \alpha T^i_{j,j} - \alpha^2 (\widehat{\nu}_j^i)^{-2}
+  } \succeq 0 \qquad \forall i \in \left\{ 1, \dots, l \right\}, j \in \left\{ 1, \dots, n_{\phi_i} \right\} \label{eqn:inclusion}\\
   &{R^i}\tr\!
-\He{X^i - \Omega^i} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \leq 0 \ \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
-  &\bmx{P & {E_i}\tr \\ \star & \hat{\mu_i}^2} \geq 0 \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
+\He{X^i - \Omega^i} R^i + He\left\{ \pmx{N^i_1\\ N^i_2\\ N^i_3} \pmx{R N_{\nu x} & \Id - R & -\Id}\right\} \preceq 0 \qquad \forall i \in \left\{ 1, \dots, l \right\} \label{eqn:finsler_constraint}\\
+  &\bmx{P & {E_i}\tr \\ \star & \widehat{\mu_i}^2} \succeq 0 \qquad \forall i \in \left\{1, \dots, n_q \right\} \label{eqn:nonlin-inclusion-constraint}
 \end{align}
@@
 \begin{theorem}\label{thm:stability} \emph{ Consider the control system in \eqref{eqn:system-dynamics} and the controller $\pi_{\text{ETM}}$ in \eqref{eqn:nn-equations}.
-Assuming the existence of the matrices $P = P\tr > 0 \in \RR^{n_x \times n_x}$,} $T = \text{diag} \pmx{T^1 & \dots & T^l} \in \RR^{n_{\phi} \times n_{\phi}} > 0$, \emph{$Z = \smallpmat{{Z^1}\tr & \dots & {Z^l}\tr}\tr \in \RR^{n_{\phi} \times n_x}$, $\bR > 0$, $\bX$, $N^i_1, N^i_2, N^i_3 \ \forall i \in \left\{ 1, \dots, l \right\}$, $S \in \RR^{n_x \times n_x}, Q \in \RR^{n_x \times n_q}, W \in \RR^{n_q \times n_q}$ and a scalar $\alpha > 0$ such that the matrix inequalities \eqref{eqn:lyapunov}, \eqref{eqn:inclusion}, \eqref{eqn:finsler} hold, where $\hat{\nu}_j^i = \min(|-\bar{\nu}_j^i - \nu^i_{*, j}|,|\bar{\nu}_j^i - \nu^i_{*, j}|)$, $
-\hat{\mu}_i = \min(|-\underline{\mu}_i - v_i^*|,|\bar{\mu}_i - v_i^*|)$ and $G^i = (T^i)^{-1} Z^i\ \forall i \in \left\{1, \dots, l \right\}$
+Assuming the existence of the matrices $P = P\tr \succ 0 \in \RR^{n_x \times n_x}$,} $T = \text{diag} \pmx{T^1 & \dots & T^l} \in \RR^{n_{\phi} \times n_{\phi}} \succ 0$, \emph{$Z = \smallpmat{{Z^1}\tr & \dots & {Z^l}\tr}\tr \in \RR^{n_{\phi} \times n_x}$, $\bR > 0$, $\bX$, $N^i_1, N^i_2, N^i_3 \ \forall i \in \left\{ 1, \dots, l \right\}$, $S \in \RR^{n_x \times n_x}, Q \in \RR^{n_x \times n_q}, W \in \RR^{n_q \times n_q}$ and a scalar $\alpha > 0$ such that the matrix inequalities \eqref{eqn:lyapunov}, \eqref{eqn:inclusion}, \eqref{eqn:finsler} hold, where $\widehat{\nu}_j^i = \min(|-\bar{\nu}_j^i - \nu^i_{*, j}|,|\bar{\nu}_j^i - \nu^i_{*, j}|)$, $
+\widehat{\mu}_i = \min(|-\underline{\mu}_i - v_i^*|,|\bar{\mu}_i - v_i^*|)$ and $G^i = (T^i)^{-1} Z^i\ \forall i \in \left\{1, \dots, l \right\}$
 Then the point $(x, \boldsymbol{\eta}) = (x_*, \0)$ is Locally Exponentially Stable (LES) with a ROA that includes the ellipsoid $\cE(P, x_*) = \left\{ x \in \RR^{n_x}, \boldsymbol{\eta} \in \RR^{l} : \tilde{x}\tr P \tilde{x} + \He{\1_l\tr \boldsymbol{\eta}} \leq 1\right\}$}
@@
 \begin{proof}
-By exploiting the inequality $$\bmx{\alpha (\hat{\nu}^i_j)^{-2} - T^i_{j, j}}(\hat{\nu}^i_j)^2\bmx{\alpha (\hat{\nu}^i_j)^{-2} - T^i_{j, j}} \geq 0$$ having $(T^i_{j,j})^2(\hat{\nu}^i_j)^2 \geq 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}$ and assuming \eqref{eqn:inclusion} we state:
+By exploiting the inequality $$\bmx{\alpha (\widehat{\nu}^i_j)^{-2} - T^i_{j, j}}(\widehat{\nu}^i_j)^2\bmx{\alpha (\widehat{\nu}^i_j)^{-2} - T^i_{j, j}} \geq 0$$ having $(T^i_{j,j})^2(\widehat{\nu}^i_j)^2 \geq 2 \alpha T^i_{j,j} - \alpha^2 (\widehat{\nu}_j^i)^{-2}$, and assuming \eqref{eqn:inclusion}, we state:
 $$
@@
 P & (T^i_{j,j} G^i_j)\tr\\
-\star & (T^i_{j,j})^2(\hat{\nu}^i_j)^2
+\star & (T^i_{j,j})^2(\widehat{\nu}^i_j)^2
 } \geq 
@@
 P & {Z_j^i}\tr\\
-\star & 2 \alpha T^i_{j,j} - \alpha^2 (\hat{\nu}_j^i)^{-2}
+\star & 2 \alpha T^i_{j,j} - \alpha^2 (\widehat{\nu}_j^i)^{-2}
 } \geq 0
@@
 0 & 2 \Id & 0\\
-\star & 0 &(T^i_{j,j})^2(\hat{\nu}^i_j)^2
-} \geq 0
+\star & 0 &(T^i_{j,j})^2(\widehat{\nu}^i_j)^2
+} \succeq 0
 $$
-We initially assume the positivity of $\boldsymbol{\eta}$, by pre and post multiplying for $\bmx{\tilde{x}\tr, \sqrt{\boldsymbol{\eta}}, 1}\tr$ and its transpose and applying Schur's complement we obtain
+We initially assume the positivity of $\boldsymbol{\eta}$, by pre- and post-multiplying for $\bmx{\tilde{x}\tr, \sqrt{\boldsymbol{\eta}}, 1}\tr$ and its transpose, and applying Schur's complement, we obtain
 $$
-\bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}}\tr \bmx{P & 0\\ 0 & 2 \Id} \bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}} \geq \tilde{x}\tr \frac{(G^i_j)\tr (G^i_j)}{\hat{\nu^i_j}^2} \tilde{x}
+\bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}}\tr \bmx{P & 0\\ 0 & 2 \Id} \bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}} \succeq \tilde{x}\tr \frac{(G^i_j)\tr (G^i_j)}{\widehat{\nu^i_j}^2} \tilde{x}
 $$
-This ensures $\cE(P, x_*) \subseteq S$ defined in Remark \ref{rem:sec-set}.
-Now assuming \eqref{eqn:finsler_constraint} hold $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$, by the application of Lemma \ref{lem:finsler} we ensure $\boldsymbol{\eta} > 0$ through Remark \ref{rem:eta-positive}.
-Considering \eqref{eqn:nonlin-inclusion-constraint} we have
+This ensures $\cE(P, x_*) \subseteq S$, as defined in Remark \ref{rem:sec-set}.
+Now, assuming \eqref{eqn:finsler_constraint} holds for all $(x, \boldsymbol{\eta}) \in \cE(P, x_*)$, the application of Lemma \ref{lem:finsler} guarantees that $\boldsymbol{\eta} > 0$ as discussed in Remark \ref{rem:eta-positive}.
+Considering \eqref{eqn:nonlin-inclusion-constraint}, we have:
 $$
@@
 0 & 2 \Id & 0 \\
-E_i & 0 & \hat{\mu_i}^2} \geq 0
+E_i & 0 & \widehat{\mu_i}^2} \succeq 0
 $$
-By expressing $v_i = E_i x$ and pre and post multiplying by $\bmx{\tilde{x}\tr & \tilde{\Phi}(Ex)\tr}\tr$ we have:
+By expressing $v_i = E_i x$ and pre- and post-multiplying by $\bmx{\tilde{x}\tr & \tilde{\Phi}(Ex)\tr}\tr$ we have:
 $$
-\bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}}\tr \bmx{P & 0\\ 0 & 2 \Id} \bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}} \geq \tilde{x}\tr \frac{E_i\tr E_i}{\hat{\mu_i}^2} \tilde{x} = \frac{v_i\tr v_i}{\hat{\mu_i}^2} \geq 0
+\bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}}\tr \bmx{P & 0\\ 0 & 2 \Id} \bmx{\tilde{x}\\ \sqrt{\boldsymbol{\eta}}} \succeq \tilde{x}\tr \frac{E_i\tr E_i}{\widehat{\mu_i}^2} \tilde{x} = \frac{v_i\tr v_i}{\widehat{\mu_i}^2} \succeq 0
 $$
-Once again this ensures $\cE(P, x_*) \subseteq \bar{S}$ defined in Assumption \ref{ass:nonlin} hence $\cE(P, x_*) \subseteq \left\{ S \cap \bar{S} \right\}$.
+Once again, this ensures $\cE(P, x_*) \subseteq \bar{S}$ as defined in Assumption \ref{ass:nonlin}, and thus $\cE(P, x_*) \subseteq \left\{ S \cap \bar{S} \right\}$.
 
-We can define the following definite-positive candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + \He{\1_l\tr\!\boldsymbol{\eta}}$. 
-By denoting $R_s$ as a proper transformation matrix such that $\bmx{\tilde{v}\tr & \tilde{\Phi}(Ex)}\tr = R_s \!
-\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)}\tr$ then the following holds $\forall x \in \cE(P, x_*)$:
+We can define the following positive-definite candidate Lyapunov function structured as: $V = \tilde{x}\tr P \tilde{x} + \He{\1_l\tr\!\boldsymbol{\eta}}$. 
+By denoting $R_s$ as a proper transformation matrix such that $\bmx{\tilde{v}\tr & \tilde{\Phi}(Ex)\tr}\tr = R_s \!
+\bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr$ the following holds for all $x \in \cE(P, x_*)$:
 \begin{equation}\label{eqn:sector-sign}
@@
 \end{equation}
-With respect to \eqref{eqn:auxiliary} and \eqref{eqn:incremental} by denoting \\$\xi = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr = R_{\nu} \zeta$, if we pre and post multiply \eqref{eqn:lyapunov} by $\bmx{\zeta\tr & \sqrt{\boldsymbol{\eta}}\tr}\tr$ and its transpose, by expressing the midterm as follows:
+With respect to \eqref{eqn:auxiliary} and \eqref{eqn:incremental} by denoting \\$\xi = R_{\nu} \bmx{\tilde{x}\tr & \tilde{\psi}\tr & \tilde{\Phi}(Ex)\tr}\tr = R_{\nu} \zeta$, if we pre- and post-multiply \eqref{eqn:lyapunov} by $\bmx{\zeta\tr & \sqrt{\boldsymbol{\eta}}\tr}\tr$ and its transpose,we can express the intermediate term as follows:
 \begin{equation}
@@
 \end{equation}
-and considering \eqref{eqn:sector-sign} we verify Assumption \ref{ass:nonlin} then by collecting the proper terms $\forall (x, \boldsymbol{\eta}) \in \cE(P, x_*)$ we have
+Considering \eqref{eqn:sector-sign}, we verify Assumption \ref{ass:nonlin}. Then, by collecting the appropriate terms for all $(x, \boldsymbol{\eta}) \in \cE(P, x_*)$ we have:
 \begin{equation}
@@
 \end{equation}
-{\color{blue} Therefore, $\cE(P, x_*)$ is forward invariant and all trajectories starting inside such a set converge to $(x, \0)$, thus concluding the proof.}
+Therefore, $\cE(P, x_*)$ is forward invariant, and all trajectories starting inside such a set converge to $(x, \0)$, thus concluding the proof.
 \end{proof}
+\begin{table*}[t]
+    \centering
+    \begin{tabular}{c|cccc|c|c|c}
+    Setup & $\lambda_{l1}$  & $\lambda_{l2}$ & $\lambda_{l3}$ & $\lambda_u$ & $\lambda_{\text{tot }}$& $\VV (\cE(P, x_*))$ & $n_{\text{steps}}$\\
+    \hline
+    $\cS 1$ & $96.28$ & $54.80$ & $44.58$ & $28.17$ & $64.84$ & $80.28$ & $323$\\
+    $\cS 2$ & $74.19$ & $40.65$ & $35.48$ & $30.32$ & $49.90$ & $65.78$ & $310$\\
+    $\cS 3$ & $55.53$ & $36.70$ & $32.88$ & $21.29$ & $41.49$ & $81.34$ & $2278$\\
+    $\cS 4$ & $44.90$ & $38.13$ & $35.61$ & $24.64$ & $39.39$ & $80.89$ & $2098$\\
+    \end{tabular}
+    \caption{Average results for $20$ simulations with different setups}
+    \label{tab:results}
+\vspace{.5em}
+\hrule
+\end{table*}
 \subsection{Optimization procedure}
+We now state a corollary whose results will be useful for deriving performance upgrades that will be discussed later.
+
+\begin{corollary}\label{cor:optimization}
+  \emph{Given Theorem \ref{thm:stability}, considering the decision variables $\Sigma, P, T, Z, \bX, \bR, S, Q, W, N^1_i, N^2_i, N^3_i, \beta_i \quad \forall i \in \left\{ 1, \dots, l \right\}$ with $\Sigma$ being a diagonal matrix $\Sigma > 0 \in \RR^{(n_x + 2 n_{\phi} + n_q) \times (n_x + 2 n_{\phi} + n_q)}$, and $\beta_i$ being scalars with the following optimization scheme:}
+  \begin{equation}\label{eqn:optimization}
+  \begin{aligned}
+    &\min \trace{P} + \trace{\Sigma} + \sum_{i=1}^{l} \beta_i\\
+    &\text{subject to } \eqref{eqn:LMI},\\ 
+    &\qquad \qquad \Xi + \Sigma \succeq 0,\\
+    &\qquad \qquad \bmx{-\beta_i \Id & X^i\\
+    {X^i}\tr & -\Id} \preceq 0 \qquad i \in \left\{ 1, \dots, l \right\}
+  \end{aligned}
+  \end{equation}
+  \emph{Then $(x_*, \0)$ is LES and $\cE(P, x_*)$ is an inner approximation of the ROA.}
+\end{corollary}
+
+By adding these constraints, we obtain the following performance upgrades:
+\begin{itemize}
+  \item Minimizing the trace of $P$ maximizes the volume of the ellipsoid $\cE(P, x_*)$, thereby increasing the Region of Attraction (ROA).
+  \item Minimizing the trace of $\Sigma$ reduces the conservatism of the controller by obtaining a constraint matrix $\Xi$ that is directly related to the decay rate of the Lyapunov function, as close to zero as possible but never greater due to the second condition in \eqref{eqn:optimization}.
+  \item Applying the Schur complement to the last condition in \eqref{eqn:optimization}, we obtain:
+  $$
+  || X^i || \preceq \beta_i \Id \qquad \forall i \in \left\{1, \dots, l \right\}
+  $$
+ This means that the smaller the value of $\beta_i$, the smaller the triggering function in the ETM \eqref{eqn:etm-trigger}, thus reducing the number of events.
+\end{itemize}
+
+It is important to comment on the quantity $\alpha$ in \eqref{eqn:inclusion}.
+If $\alpha$ were treated as a variable, the statements of Theorem \ref{thm:stability} would not hold as LMI conditions because bilinear terms would arise.
+To address this, we fix $\alpha$ to a constant value.
+The selection of $\alpha$ is important, as it directly influences the conservatism of the solution.
+To find a good value for $\alpha$, we use a \emph{golden ratio search algorithm}.
+This approach is based on the bisection method, with some basic improvements to enhance efficiency.
+
+\begin{itemize}
+  \item It reduces the number of LMI solutions required by efficiently narrowing the search interval.
+  \item It maintains a constant reduction factor for the interval size, leading to more consistent progress towards the optimal value.
+  \item It is particularly effective for unimodal functions, such as the one we encounter when optimizing the volume of the Region of Attraction (ROA).
+\end{itemize}
 
 \section{Simulations}
+We considered an inverted pendulum system with mass $m = 0.15 kg$, length $l = 0.5 m$, damping coefficient $\mu = 0.05 N ms/rad$, gravitational acceleration $g = 9.81 m/s^2$.
+Additionally, an integrator is added to the system's dynamics.
+The discrete-time (DT) model of the system is given in the form of \eqref{eqn:system-dynamics} as:
+\begin{equation}\label{eqn:pendulum}
+\begin{aligned}
+  &x = \bmx{\theta \\ \dot{\theta} \\ z} \quad 
+  A = \bmx{1 & dt & 0\\
+  \frac{g}{l} dt & 1 - \frac{\mu}{m l^2} dt & 0\\
+  1 & 0 & 1}\quad
+  B = \bmx{0\\ \frac{\bar{\tau}}{m l^2} dt\\ 0} \\
+  &C = \bmx{0 \\ \frac{g}{l} dt \\ 0} \quad
+  D = \bmx{0 \\ 0 \\ -1} \quad
+  E = \bmx{1 & 0 & 0}
+\end{aligned}
+\end{equation}
+With the maximum torque available $\bar{\tau} = 5 N m$, the sampling time $dt = 0.02 s$ and the nonlinearity $\Phi(Ex) = \text{sin}(Ex) - Ex$, proper quadratic abstractions to handle this nonlinearity have are already been explored in the literature, such as in \citep{junnarkar2024synthesizing}, which uses:
+$$
+\bmx{y\\sin(y) - y}\tr \bmx{0 & -1 \\ -1 & -2} \bmx{y\\sin(y)-y} > 0 \qquad \forall y \in \bmx{-\pi, \pi}
+$$
+Considering the system under examination, due to the periodicity of the $sin$ function, the previous conditions hold globally.
+By selecting $n_q = 1, S = 0, Q = -1, W=-2$, we avoid adding condition \eqref{eqn:nonlin-inclusion-constraint} in the LMI problem formulation.
+A stabilizing NN controller was trained using Reinforcement Learning (RL), structured as an MLP of $3$ layers each having $n_{\phi_1} = n_{\phi_2} = n_{\phi_3} = 32$ neurons.
+The addition of the integrator in the system dynamics makes the training less trivial, enabling compensation for the intrinsic equilibrium points of the controller, which are typically different from the origin.
+The constroller's output $\bar{u}$ is a scalar, and the saturation level $\widehat{u}$ of the last ETM (see fig.
+\ref{fig:second_scheme}) is set to $1$, due to the nature of RL training, which always normalizes the final output of the network.
+By imposing $\alpha = 0.053$ we solved the LMI for 4 significant setups:
+\begin{itemize}
+    \item $\cS 1$: Old triggering conditions with static ETM minimizing $\trace{P}$ 
+    \item $\cS 2$: Old triggering conditions with dynamic ETM minimizing $\trace{P} + \trace{\Sigma}$
+    \item $\cS 3$: New triggering conditions minimizing $\trace{P} + \trace{\Sigma}$
+    \item $\cS 4$: New triggering conditions minimizing $\trace{P} + \trace{\Sigma} + \sum_{i=1}^l \beta_i$
+\end{itemize}
+In Table \ref{tab:results}, the novelties and upgrades of the proposed method are highlighted.
+The data comes from the average of simulations starting from $20$ initial configurations such that $(x_0, \boldsymbol{\eta}_0) \in \cE(P_{\cS 2}, x_*)$, which corresponds to the more conservative result.
+The simulations are stopped whenever the Lyapunov function reaches an arbitrary small threshold, $10^{-15}$ in our case.
+The update rates are denoted by $\lambda_{li}$ and $\lambda_{\text{tot}}$, representing the percentage of the $i$-th layer and the overall controller respectively.
+Their expression follows:
+$$
+\lambda_{li} = \frac{n_{\text{events}_i}}{n_{\text{steps}}} \cdot 100; \qquad \lambda_{\text{tot}} = \frac{\sum_{i=1}^l \lambda_{li} \cdot n_{\phi_i}}{l}
+$$
+The volume of the ellipsoid $\cE(P_{\cC_i}, x_*)$ is a direct indication of the conservatism of the solution and is computed as:
+$$
+\VV(\cE(P_{\cC_i}, x_*) = \frac{4}{3} \frac{\pi}{\sqrt{\det P_{\cC_i}}}
+$$
+Finally, we analyze the number of simulation steps required to reach the previously mentioned threshold.
+A considerable decrease in the update rate of the controller is observed with every proposed method, accompanied by a more uniform event pattern across the different layers.
+The increased conservatism observed in $\cS 2$ is counterbalanced by the introduction of $X^i$ matrices in Lemma \ref{lem:finsler}, which leads to a larger ROA approximation and an improved update profile.
+$\cS 4$ represents a performance trade-off compared to $\cS 3$ in terms of computational savings and ellipsoid size.
+Both $\cS 3$ and $\cS 4$ maintain a low value for $\lambda_{\text{tot}}$ while having higher $n_{\text{steps}}$, indicating that the raw number of events is low, leading to more efficient event-triggering and fewer updates.
 
-\section{Conclusions}
+The following plots refer to $\cS 4$ with $$
+x_0 = \pmx{-9.43(deg) & 5.77(rad/s) &  0.0}\tr; \quad \boldsymbol{\eta}_0 = 0.01 \cdot \1$$ 
+as initial configuration.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figures/evolution_plot}
+    \caption{First $600$ steps evolution of $u, x$}
+    \label{fig:evolution}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figures/event_plot}
+    \caption{Events occurring in first $300$ steps}
+    \label{fig:events}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figures/lyapunov_plot}
+    \caption{First $1000$ steps evolution of $V$ and normalized $\Delta V$}
+    \label{fig:lyapunov}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figures/eta_plot}
+    \caption{First $600$ steps evolution of $\eta^i$ dynamic thresholds}
+    \label{fig:eta}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.45\textwidth]{Figures/ellipsoid_plot}
+    \caption{Trajectory of the system within $\cE(P, x_*)$}
+    \label{fig:ellipsoid}
+\end{figure}
+
+{\color{red} TODO: Chiedi misericordia per l'introduzione}
+
+\section{Conclusions and future works}
+In this work, we improved upon the previously published results by addressing a more complex non-linear system controlled by a neural network with more layers, which often leads to a more conservative behavior.
+Our approach effectively reduced this conservatism, improving the system's performance.
+Future work will focus on refining Lemma 1 related to Finsler's lemma, specifically to better handle the inclusion where the parameter $\alpha$ appears.
+Additionally, we plan to explore methods for managing updates close to the equilibrium point, inspired by the concepts introduced in stubborn observer theory.
 

### 2024-12-16 Paper draft/draft.tex
Updated version

@@
 \begin{equation}
-    \bX = \text{diag} \pmx{R_1\tr X^1 R_1 & \dots & R_l\tr X^l & R_l}, \qquad \xi = \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\nu}}
+    \bX = \text{diag} \pmx{{R^1}\tr X^1 R^1 & \dots & {R^l}\tr X^l & R^l}, \qquad \xi = \bmx{\tilde{x} \\ \tilde{\psi} \\ \tilde{\nu}}
 \end{equation}
@@
 \begin{aligned}
-    \bar{A} &= A + B R_{\omega}\\
-    \bar{B} &= -B N_{u \omega} R\\
-    \tilde{x}^+ &= \bar{A}\tilde{x} + \bar{B}\tilde{\psi} + C \tilde{\Phi}\\
-R_{\nu} &= \bmx{
+    &\bar{A} = A + B R_{\omega}\\
+    &\bar{B} = -B N_{u \omega} R\\
+    &\tilde{x}^+ = \bar{A}\tilde{x} + \bar{B}\tilde{\psi} + C \tilde{\Phi}\\
+&R_{\nu} = \bmx{
   \Id & 0 & 0 \\
@@
 } \in \RR^{(n_x + n_{\phi} + n_q) \times (n_x + 2 n_{\phi})}\\
-R_s &\in \RR^{(n_x + n_{\phi} + n_q) \times (n_x + n_q)}
+&R_s \in \RR^{(n_x + n_{\phi} + n_q) \times (n_x + n_q)}
 \end{aligned}
@@
 \star & (T^i_{j,j})^2(\widehat{\nu}^i_j)^2
-} \geq 
+} \succeq 
 \bmx{
@@
 \star & 2 \alpha T^i_{j,j} - \alpha^2 (\widehat{\nu}_j^i)^{-2}
-} \geq 0
+} \succeq 0
 $$
@@
 The following plots refer to $\cS 4$ with $$
-x_0 = \pmx{-9.43(deg) & 5.77(rad/s) &  0.0}\tr; \quad \boldsymbol{\eta}_0 = 0.01 \cdot \1$$ 
+x_0 = \bmx{-9.43(deg) \\ 5.77(rad/s) \\  0.0}; \quad \boldsymbol{\eta}_0 = 0.01 \cdot \1; \quad d = 2.76(deg)$$ 
 as initial configuration.
@@
     \includegraphics[width=0.45\textwidth]{Figures/ellipsoid_plot}
-    \caption{Trajectory of the system within $\cE(P, x_*)$}
+    \caption{Trajectory of the system within $\cE(P, x_*)$ and with respect to its projections}
     \label{fig:ellipsoid}
